### 2020-11-18 sceptre_paper/manuscript/manuscript_v3.tex
added manuscript and started organizing plotting code

%% Template for a preprint Letter or Article for submission
%% to the journal Nature.
%% Written by Peter Czoschke, 26 February 2004
%%

\documentclass{nature}

%% make sure you have the nature.cls and naturemag.bst files where
%% LaTeX can find them

\usepackage{graphicx}	
\usepackage{amssymb, amsmath, bbm}
\usepackage{geometry}
\usepackage[colorlinks=true,citecolor=blue,linkcolor=blue]{hyperref}
\newcommand\independent{\perp \!\!\! \perp}
\usepackage{chngcntr}

\bibliographystyle{naturemag}

%\usepackage[minnames = 3, maxnames=3, url = false, doi = false, isbn = false, sorting = none, backend=bibtex]{biblatex}
%\addbibresource{/home/ekatsevi/Dropbox/Research/Literature/library.bib}

\title{Conditional resampling improves calibration \\ in single cell CRISPR screen analysis}

%% Notice placement of commas and superscripts and use of &
%% in the author list

\author{Eugene Katsevich$^{1}$, Timothy Barry$^{2}$, and Kathryn Roeder$^{2,3}$}

\begin{document}

\maketitle

\addtocounter{page}{-1}
\thispagestyle{empty}

\begin{affiliations}
 \item Department of Statistics, Wharton School, University of Pennsylvania
 \item Department of Statistics and Data Science, Carnegie Mellon University
 \item Computational Biology Department, Carnegie Mellon University
\end{affiliations}

\begin{abstract}
Single cell CRISPR screens are an emerging biotechnology promising unprecedented insights into gene regulation. However, the analysis of these screens presents significant statistical challenges. For example, technical factors like sequencing depth impact not only expression measurement but also perturbation detection, creating a confounding effect. We demonstrate on a recent large-scale single cell CRISPR screen how these challenges cause calibration issues among existing analysis methods. To address these challenges, we propose SCEPTRE: analysis of single cell perturbation screens via conditional resampling. This methodology, designed to avoid calibration issues due to technical confounders and expression model misspecification, infers associations between perturbations and expression by resampling the former according to a working model for perturbation detection probability in each cell. Applying SCEPTRE to the CRISPR screen data yields excellent calibration on negative control perturbations and 217 regulatory relationships not found in the original study, many of which are supported by existing functional data.
\end{abstract}

\clearpage

Single cell CRISPR screens (implemented independently as Perturb-seq\cite{Dixit2016, Adamson2016}, CRISP-seq\cite{Jaitin2016}, and CROP-seq\cite{Datlinger2017}) facilitate unprecedented insights into gene regulation by measuring the impact of precisely targeted perturbations on the whole transcriptome. These pooled screens deliver a library of CRISPR perturbations, targeting either genes or non-coding regulatory elements, to a population of cells and then employ single cell RNA sequencing (scRNA-seq) to identify the perturbations each cell received and measure its transcriptional response. Single cell CRISPR screens have been scaled up through multiplexing to systematically map gene-enhancer relationships\cite{Xie2017,Gasperini2019,Xie2019}, applied using combinatorial perturbation libraries to study genetic interactions\cite{Norman2019}, and extended to single cell protein expression readout\cite{Mimitou2019,Frangieh2020}. Given their scale, resolution, and versatility, these assays are likely to be a major driver of discovery in functional genomics and its applications to the molecular understanding of human disease in the coming decade\cite{Gasperini2020}.

Despite their promise, the data obtained from such assays pose significant statistical challenges. In particular, various difficulties have been encountered in calibrating tests of association between a CRISPR perturbation and the expression of a gene. Gasperini et al.\cite{Gasperini2019} found substantial inflation in their negative binomial regression based $p$-values for negative control perturbations. Similarly, Xie et al.\cite{Xie2019} found an excess of false positive hits in their rank-based ``virtual FACS" analysis. Finally, Yang et al.\cite{Yang2020} found that their permutation-based scMAGeCK-RRA method deems almost all gene-enhancer pairs significant in a reanalysis of the Gasperini et al. data. While these works propose ad hoc fixes to improve calibration, it is clear that new analysis methodology is indispensable to fully realize the potential of single cell CRISPR screens. However, a recent review\cite{Lin2020} noted that few such methods have been proposed so far.

In this work, we explore statistical challenges of single cell CRISPR screens and present a novel analysis methodology to address them. In addition to known scRNA-seq analysis challenges such as expression normalization and discreteness\cite{Townes2019,Hafemeister2019,Svensson2020}, we identify a key challenge that sets CRISPR screens apart from traditional differential expression experiments: the ``treatment"---in this case the presence of a CRISPR perturbation in a given cell---is subject to measurement error\cite{Dixit2016,Hill2018,Replogle2020}. In fact, underlying this measurement error are the same technical factors contributing to errors in the measurement of gene expression, including sequencing depth and batch effects. These technical factors therefore act as confounders, invalidating traditional nonparametric calibration approaches such as those based on ranks or permutations. Indeed, the symmetry among cells assumed by these approaches breaks down in the presence of technical factors varying across cells. On the other hand, parametric modeling of single cell expression data is also fraught with challenges and remains an active area of research.

To address these challenges, we propose SCEPTRE (analysis of Single CEll PerTurbation screens via conditional REsampling). This resampling methodology is based on the conditional randomization test\cite{CetL16}, a recently proposed statistical methodology which, like traditional nonparametric methods, does not rely on correct specification of the outcome (expression) model. Unlike traditional nonparametric methods, SCEPTRE employs a resampling scheme that explicitly accounts for heterogeneity in technical factors among cells. We analyze both synthetic data and that of Gasperini et al.\cite{Gasperini2019} to illustrate the aforementioned analysis challenges and the performance of the proposed methodology. SCEPTRE demonstrates excellent calibration and reveals many novel regulatory relationships supported by a variety of existing sources of functional evidence.

\section*{Results}

\paragraph{Analysis challenges.}

The Gasperini et al.\cite{Gasperini2019} data exemplify several of the analysis challenges in single cell CRISPR screens. A library of CRISPR guide RNAs (gRNAs) targeting 5,779 candidate enhancers, 50 negative controls, and 381 positive controls was transduced via CROP-seq\cite{Datlinger2017,Hill2018} at high multiplicity of infection into a population of 207,324 K562 cells expressing the Cas9-KRAB repressive complex, each cell receiving an average of 28 perturbations. Polyadenylated gRNA barcodes were sequenced alongside the whole transcriptome, and perturbation identities assigned to cells by thresholding the resulting gRNA UMI counts. Initially, $p$-values for each gRNA-gene pair were obtained from a DESeq2\cite{Love2014}-inspired negative binomial regression analysis implemented in Monocle2\cite{Qiu2017}. By examining the distribution of negative binomial $p$-values pairing each gene with each of 50 non-targeting control (NTC) gRNAs, Gasperini et al. found that these $p$-values are inflated (Figure~\ref{fig:analysis-challenges}a). To remedy this issue, Gasperini et al. calibrated the candidate enhancer $p$-values against the distribution of all gene-NTC $p$-values instead of the uniform distribution. The resulting ``empirical" $p$-values were used for determining significance.

First, we illustrate the confounding effect discussed in the introduction. Despite the use of a targeted amplification protocol for gRNA detection\cite{Hill2018}, not all gRNAs are detected. Indeed, the total number of gRNAs detected in a cell increases with the sequencing depth, defined as the total number of mRNA UMIs detected per cell ($\rho = 0.35, p < 10^{-15}$; Figure~\ref{fig:analysis-challenges}b). We observed a similar phenomenon in the Mosaic-seq data\cite{Xie2017} ($\rho = 0.23, p < 10^{-15}$; Figure~\ref{fig:mosaicseq}). Therefore, sequencing depth and other technical factors induce correlations between the measured gRNA presence and gene expression, even in the absence of a regulatory relationship (Figure~\ref{fig:analysis-challenges}c). Parametric regression approaches like those employed by Gasperini et al. are the easiest way to adjust for confounders, but their heavy reliance on correct model specification creates a different set of calibration issues.

We hypothesized that imperfect dispersion estimation may be responsible for the inflation observed by Gasperini et al. Monocle2 fits a raw dispersion for each gene, then fits a mean-dispersion relationship across genes, and finally collapses each raw dispersion estimate onto this fitted line (Figure~\ref{fig:analysis-challenges}d). While the raw dispersion estimates are noisy and some shrinkage is helpful\cite{Hafemeister2019}, the collapsing of the dispersion estimates to the fitted curve is likely to underestimate the dispersions for points significantly above the curve. On the other hand, points near the curve may not have this issue. Indeed, plotting the 50 raw NTC $p$-values for genes near the curve, we found good calibration (e.g., LUC7L2; Figure~\ref{fig:analysis-challenges}e), while genes significantly above the curve showed marked signs of raw $p$-value inflation (e.g., HIST1H1D; Figure~\ref{fig:analysis-challenges}f).

This heterogeneity in inflation among gene-gRNA pairs reveals a weakness of the ``one-size-fits-all" NTC-based calibration approach employed by Gasperini et al.: building a null distribution using all gene-NTC pairs leads to overcorrection for some gene-enhancer pairs (false negatives) and undercorrection for others (false positives). For example, the empirical correction is unnecessary for genes like LUC7L2. However, we pay a price in sensitivity for this empirical correction; it decreases the significance of the $p$-value for a positive control gRNA by three orders of magnitude (Figure~\ref{fig:analysis-challenges}e). On the other hand, the empirical correction is insufficient for high-dispersion genes. We computed the deviation from uniformity of the empirical NTC $p$-values for each gene using the Kolmogorov-Smirnov (KS) test, represented by the color of each point in Figure~\ref{fig:analysis-challenges}b. Circled genes have significantly miscalibrated empirical $p$-values based on a Bonferroni correction at level $\alpha = 0.05$; 21 of these were flagged by Gasperini et al. as outliers but 21 were not. Among the latter group is HIST1H1D, whose empirical $p$-values are still inflated (Figure~\ref{fig:analysis-challenges}f), leading to two potential false discoveries, one of which is deemed ``high confidence" (Figure~\ref{fig:analysis-challenges}g, Figure~\ref{fig:potential-false-positives}).

\clearpage

\addtocounter{page}{-1}
\thispagestyle{empty} 
\begin{figure}[h!]
	\includegraphics[width = \textwidth]{figures/Figure1/Figure1.png}
	\caption{\textbf{CRISPR screen analysis challenges can lead to false positives and false negatives.} \textbf{a}, QQ-plot of Gasperini et al. $p$-values for all gene-gRNA pairs involving either candidate enhancer or negative control gRNAs. Inflation in negative controls makes it harder to interpret associations between genes and candidate enhancers, motivating the empirical $p$-value correction. \textbf{b-c}, Sequencing depth (total mRNA UMIs per cell) impacts gRNA detection (b) and observed expression levels, and therefore acts as a confounder (c). This analysis challenge distinguishes CRISPR screens from traditional differential expression applications. \textbf{d-g}, Collapsing dispersion estimates to fitted mean-variance relationship (d) leads to over-correction for some genes (e) and under-correction for others (f, g). \textbf{d,} Dispersion estimation procedure employed leads to miscalibration for high-dispersion genes, which the empirical correction does not adequately correct for, as measured by KS test applied to empirical $p$-values per gene (point colors). \textbf{e,} Raw $p$-values already well-calibrated for LUC7L2 gene, so empirical correction unnecessarily shrinks the significance of the association with TSS-targeting gRNA, depicted by horizontal lines, by three orders of magnitude. \textbf{f,} Empirical correction not strong enough for HIST1H1D, which is among circled genes in panel d, which have an NTC-based miscalibration $p$-value smaller than the Bonferroni threshold. \textbf{g}, Under-correction leads to two potential false discoveries for HIST1H1D. Dashed horizontal line represents the multiple testing threshold.}
	\label{fig:analysis-challenges}
\end{figure}

\clearpage

\paragraph{Improvements to the negative binomial approach.}

We attempted to alleviate the miscalibration within the negative binomial regression framework (Figure \ref{fig:calibration-comparison-all}). First, we replaced the collapsed dispersion estimates by the raw estimates (recall Figure~\ref{fig:analysis-challenges}d). This simple modification is based on the intuition that not as much dispersion shrinkage is necessary as in bulk RNA-seq, since in scRNA-seq we have many more samples (one for each cell) to estimate this parameter. This dispersion estimate modification yielded greatly improved calibration over all gene-NTC pairs (blue curve in Figure \ref{fig:calibration-comparison-all}a), but separating by NTC revealed substantial remaining gRNA-dependent miscalibration (Figure~\ref{fig:calibration-comparison-all}b). 
This dependency on gRNA further underscores the problem with ``one-size-fits-all" NTC-based calibration approaches, also implemented by Xie et al.\cite{Xie2019}

We hypothesized that this remaining miscalibration stems from inadequate sequencing depth correction. Indeed, we found we could predict the direction and strength of the miscalibration for each gRNA from an adjusted correlation between the sequencing depth and the gRNA presence (Methods). This yielded $z$-values as shown in the legend of Figure~\ref{fig:calibration-comparison-all}b; large positive (negative) $z$-values signal spurious positive (negative) correlations between gRNA presence and sequencing depth, leading to conservative (liberal) bias. Gasperini et al. corrected for sequencing depth using DESeq2-style size factors, whereas recent work on scRNA-seq analysis\cite{Hafemeister2019} advocates for including sequencing depth as a covariate in the negative binomial regression. We therefore included as covariates both the total number of UMIs observed and the number of genes with at least one UMI in a given cell, in addition to the technical factors Gasperini et al. corrected for (see Methods).

To assess the impact of these two improvements, we applied the negative binomial test with and without these improvements (four total possibilities) to all 50 NTC gRNAs paired with all genes  (Figure~\ref{fig:calibration-comparison-all}a,c). The changes to confounder correction and dispersion estimation both markedly improve calibration, especially when used in conjunction. However, the resulting improved negative binomial method still shows clear signs of miscalibration, producing mostly conservative $p$-values. There are many possible reasons for the remaining miscalibration, most likely due to remaining misspecification of the negative binomial model. While more effort could certainly be invested in further improvements of this parametric model for gene expression, we acknowledge that modeling scRNA-seq expression is a challenging problem that remains open. This highlights the appeal of nonparametric approaches; however, as discussed before, traditional nonparametric approaches are not applicable in the presence of technical factors impacting both gRNA detection and gene expression measurement. To resolve this impasse, we propose a resampling-based approach designed to account for such confounders.

\clearpage

\paragraph{SCEPTRE: Analysis of single cell perturbation screens via conditional resampling.}

Instead of treating cells symmetrically, we propose to estimate the probability a given gRNA will be observed in a given cell, based on that cell's technical factors. For example, this probability tends to be higher in cells with larger sequencing depths. We propose to generate a null distribution for a given gene and a given gRNA by independently resampling the gRNA assignments for each cell based on their respective perturbation probabilities (Figure~\ref{fig:method}). This scheme is an instance of the conditional randomization test (CRT), recently proposed by Cand\`{e}s et al.\cite{CetL16} In contrast to standard permutation tests, the CRT explicitly accounts for technical factors that vary from cell to cell.

To quantify the effect of a given gRNA on a given gene, we use the improved negative binomial regression statistic described above. This yields a $z$-value, which would typically be compared to a standard normal null distribution based on the parametric negative binomial model. Instead, we build a null distribution for this statistic via conditional resampling. To this end, we first fit a logistic regression model for the observation of the gRNA in a cell based on its technical factors, yielding a fitted probability for each cell. Then, we generate a large number of resampled datasets, where the expression and the technical factors stay the same, while the gRNA assignment is redrawn independently for each cell based on its fitted probability. The negative binomial $z$-value is then recomputed for each of these datasets, which comprise a null distribution (depicted as a gray histogram in Figure~\ref{fig:method}). The SCEPTRE $p$-value is defined as the left tail probability of the original $z$-value under this null distribution, for sensitivity to perturbations that decrease expression (the right tail could be used for screens targeting repressive elements). The conditional resampling null distribution can differ substantially from that based on the negative binomial model---for the same test statistic---depending on the extent of model misspecification (Figure~\ref{fig:NB-vs-CP}).

To mitigate the extra computational cost of resampling, we implemented several accelerations. We found that the skew-$t$ distribution, used by CRISPhieRmix\cite{Daley2018} for a different purpose, provided a good fit to the null histograms. We employed this skew-$t$ approximation to obtain precise $p$-values based on a limited number of resamples (500 in the current implementation). Furthermore, we implemented computational accelerations that reduced the cost of each resample by a factor of about 100 (see Methods). The original negative binomial regression takes about 3 seconds per gene-gRNA pair, while recomputing the test statistic for 500 resamples takes a total of 16 seconds. Therefore, SCEPTRE takes about 19 seconds per pair, compared to 3 seconds for the original method.

\clearpage
\thispagestyle{empty} 
\addtocounter{page}{-1}
\begin{figure}[h!]
	\includegraphics[width = \textwidth]{figures/Figure2/Figure2.png}
	\caption{\textbf{SCEPTRE: Analysis of single cell perturbation screens via conditional resampling.} A schematic and outline of the SCEPTRE methodology for one gene and one gRNA is shown, entailing the application of the conditional randomization test\cite{CetL16} to single cell CRISPR screens. The idea is to assign to every cell the probability of gRNA observation based on its technical factors, and then to build a null distribution for the negative binomial $z$-value by independently resampling the gRNA indicator for each cell according to these probabilities to form ``negative control" data sets. A skew-$t$ distribution is fit to the resulting histogram to obtain precise $p$-values based on a limited number of resamples, against which the original NB $z$-value is compared. The dashed line shows the standard normal distribution, against which the NB $z$-value would normally be compared.}
	\label{fig:method}
\end{figure}
\clearpage

\paragraph{SCEPTRE is well calibrated despite sequencing depth confounding and expression model misspecification.}

First, we demonstrated SCEPTRE's calibration in a small proof-of-concept simulation with 1,000 cells, one gene, one NTC gRNA, and one confounder (Figure~\ref{fig:calibration}a). We considered the one-sided $z$ test statistic based on a negative binomial regression with dispersion 1. We then generated the data from this model as well as from three others: one with dispersion 0.2, one with dispersion 5, and one with dispersion 1 but with 25\% zero-inflation. We compared three ways of building a null distribution for the test statistic based on the possibly misspecified negative binomial model: the standard parametric approach, the permutation approach (based on permuting gRNA assignments), and conditional resampling. The parametric approach works as expected when the negative binomial model is correctly specified, but breaks down in all three cases of model misspecification. The permutation approach is systematically conservative because of inadequate confounder correction. Finally, SCEPTRE is well-calibrated regardless of expression model misspecification.

Next, we applied SCEPTRE to test the association between all NTC gRNAs and all genes in the Gasperini et al. data (Figure~\ref{fig:calibration}b-c). For comparison, we also applied the original negative binomial method, the improved negative binomial method, and a permutation-based calibration of the latter. The simplest calibration assessment is to compare the resulting 538,560 $p$-values to the uniform distribution (Figure~\ref{fig:calibration}b). SCEPTRE shows excellent calibration, substantially improving on the parametric approach based on the same test statistic. Breaking the pairs down by NTC (Figure~\ref{fig:calibration}c), SCEPTRE again shows nearly perfect calibration, as evidenced by its colored confidence region coinciding with the gray shaded region representing the 95\% confidence band for the uniform distribution.

To assess the sensitivity of SCEPTRE, we applied it to the 381 positive control gRNA / gene pairs assayed in Gasperini et al (Figure~\ref{fig:calibration}d), where each gRNA targeted the transcription start site of the corresponding gene. Comparing to the original empirical $p$-values, SCEPTRE $p$-values for these positive controls are generally more significant, suggesting greater sensitivity. We note that the empirical correction employed by Gasperini et al. limits the accuracy of $p$-values to about $10^{-6}$, explaining at least part of the difference.

\clearpage
\thispagestyle{empty} 
\addtocounter{page}{-1}
\begin{figure}[h!]
	\includegraphics[width = \textwidth]{figures/Figure3/Figure3.png}
	\caption{\textbf{SCEPTRE improves on the sensitivity and specificity of the parametric approach by correcting for expression model misspecification.} \textbf{a,} Numerical simulation comparing three ways of calibrating a negative binomial test statistic (parametric, permutation, conditional resampling), when the test statistic is based on the correct model or contains expression model misspecifications. Only the conditional resampling approach maintains calibration despite model misspecification. \textbf{b,c,} Application of SCEPTRE to Gasperini et al negative control gRNAs, comparing all gene-NTC pairs to the uniform distribution (b) or breaking down by NTC (c); details as in Figure~\ref{fig:calibration-comparison-all}a,c. The matching of SCEPTRE's colored region with the gray one in panel c, together with the overall near-uniformity in panel b, demonstrates its excellent calibration on real negative control data. \textbf{d,} SCEPTRE $p$-values for TSS-targeting controls generally more significant than the original empirical $p$-values; dashed line is the $45^\circ$ line. Note: $p$-values truncated for visualization, and empirical $p$-values in panel d at least 1/538,560 $\approx$ 2$\times 10^{-6}$ by construction.}
	\label{fig:calibration}
\end{figure}
\clearpage

\paragraph{Analysis of candidate gene-enhancer regulatory relationships.}

We applied SCEPTRE to the 84,595 gene-enhancer pairs considered in Gasperini et al., encompassing 10,560 genes and 5,779 candidate enhancers (Figure~\ref{fig:application}a). We applied the Benjamini-Hochberg correction at level 0.1 (the same level used in the original analysis) to the $p$-values obtained for all of these candidate pairs, obtaining a total of 585 gene-enhancer pairs. By comparison, Gasperini et al. found 470 high-confidence pairs. Comparing the SCEPTRE $p$-values against the original empirical $p$-values (Figure~\ref{fig:application}b), we see that the two often diverge substantially. Our analysis found 217 gene-enhancer pairs that the original analysis did not, while 102 were found only by the original analysis. Many of the discoveries found only in the original analysis show signs of the $p$-value inflation observed in Figure~\ref{fig:analysis-challenges} (Figure~\ref{fig:potential-false-positives}).

Among the 217 new gene-enhancer pairs discovered, several are supported by evidence from orthogonal functional assays. In particular, we highlight five of these pairs (Figure~\ref{fig:application}c) involving genes not paired to any enhancers in the original analysis, which are supported by GTEx\cite{GTExPilot} eQTL $p$-values in whole blood or enhancer RNA correlation $p$-values across tissues from the FANTOM project\cite{Andersson2014}. These pairs are listed in the GeneHancer database\cite{Fishilevich2017}, which aggregates eQTL, eRNA, and other sources of evidence of gene-enhancer interactions. The SCEPTRE $p$-values for these promising pairs are generally 1-2 orders of magnitude more significant than the original empirical $p$-values. Also among the 217 new gene-enhancer pairs are 9 pairs involving 7 genes that were deemed outliers in the original analysis (blue triangles in Figure~\ref{fig:application}b), underscoring SCEPTRE's ability to handle genes with arbitrary distributions that may not fit into traditional parametric models.

We also found that the total set of gene-enhancer pairs discovered was better enriched for regulatory biological signals, including HI-C and ChIP-seq. Gene-enhancer pairs discovered by SCEPTRE tended to be closer to each other on average that those discovered in the original analysis (Figure~\ref{fig:application}d). Furthermore, 76\% of SCEPTRE's 585 gene-enhancer pairs fell in the same topologically associating domain (TAD), compared to 71\% of the original 470 pairs. We also repeated Gasperini et al.'s contact frequency enrichment analysis for those pairs falling in the same TAD (Figure~\ref{fig:application}e). We found similar levels of contact frequency enrichment, despite the fact that we had 108 more gene-enhancer pairs in the same TADs. Finally, we repeated the ChIP-seq enrichment analysis of Gasperini et al. for seven cell-type relevant transcription factors and H3K27ac. We quantified enrichment of ChIP-seq signal in paired enhancers as the odds ratio that a candidate enhancer is paired to a gene, comparing those falling in the top quintile of candidate enhancers by ChIP-seq signal and those not overlapping a ChIP-seq peak at all. We find improved enrichment for each of the eight ChIP-seq targets (Figures~\ref{fig:application}f and \ref{fig:chipseq-enrichment-details}).

\clearpage
\newgeometry{top=0.6in}
\thispagestyle{empty} 
\addtocounter{page}{-1}
\vspace{-2in}

\begin{figure}
	\includegraphics[width = \textwidth]{figures/Figure4/Figure4.png}
	\caption{\textbf{Application to Gasperini et al. data yields biologically plausible gene-enhancer links.}  \textbf{a,} Well-calibrated NTC $p$-values give confidence in SCEPTRE $p$-values for candidate enhancers (compare to Figure~\ref{fig:analysis-challenges}a). \textbf{b}, Comparing the original empirical $p$-values to those obtained from SCEPTRE. The two analysis methods differ substantially, with 217 gene-enhancer links discovered only by SCEPTRE and 102 discovered only by the original. \textbf{c}, Five gene-enhancer pairs discovered by SCEPTRE but not the original analysis, each supported by a whole blood GTEx eQTL or FANTOM enhancer RNA correlation $p$-value. \textbf{d,} Distribution of distances from TSS to upstream paired enhancers. Compared to the original analysis, SCEPTRE pairs genes with nearer enhancers on average. \textbf{e,} For those gene-enhancer pairs falling in the same TAD, the cumulative distribution of the fractional rank of the HI-C interaction frequency compared to other distance-matched loci pairs within the same TAD. SCEPTRE shows similar enrichment despite finding 32\% more within-TAD pairs. Inset table shows gene-enhancer pairs falling in the same TAD. SCEPTRE finds 115 more total pairs, and a higher percentage of pairs fall in the same TAD. \textbf{f,} Enrichment of ChIP-seq signal from seven cell-type relevant transcription factors and one histone mark among paired enhancers. SCEPTRE exhibits greater enrichment across all ChIP-seq targets.}
	\label{fig:application}
\end{figure}

\clearpage

\restoregeometry

\section*{Discussion}

Here we illustrate a variety of statistical challenges arising in the analysis of single cell CRISPR screens, leaving existing methods (based on parametric expression models, permutations, or negative control data) vulnerable to calibration issues. To address these challenges, we develop SCEPTRE, a novel resampling method based on modeling the probability a given gRNA will be observed in a given cell, based on that cell's technical factors. SCEPTRE exhibits excellent calibration despite the presence of confounding technical factors and despite misspecification of single cell gene expression models. We implement computational accelerations to bring the cost of our resampling-based methodology down to well within an order of magnitude of the traditional approach, making it quite feasible to apply for large-scale data. We use SCEPTRE to reanalyze the Gasperini et al. data, yielding many new biologically plausible gene-enhancer relationships supported by evidence from eQTL, enhancer RNA co-expression, ChIP-seq, and HI-C data.

The main assumption behind SCEPTRE's validity is the accuracy of the model for gRNA observation. Our calibration results suggest that this is a reasonable assumption for the simulated and negative control data considered here. However, more work is necessary to assess the quality of gRNA measurement models across a broader range of contexts. Furthermore, we conjecture that misspecifications in the gRNA measurement model can be compensated by improved gene expression models, a statistical phenomenon referred to as ``double robustness"\cite{Robins2001,VanderLaan2003}. 
%While SCEPTRE greatly reduces the burden of modeling single cell expression data, there is therefore still good reason to seek improved expression models. 
Better expression models can also improve the sensitivity of the conditional randomization test\cite{Katsevich2020a}.

There are several directions for improvement in the analysis of single cell CRISPR screens, which are not addressed by SCEPTRE. Many of these have been considered before, especially in the context of non-single-cell CRISPR screens. Such remaining challenges include variable effectiveness of gRNAs\cite{Li2015,Daley2018,Xie2019}, interactions among perturbed elements\cite{Xie2017,Zamanighomi2019,Norman2019}, and the limited resolution of CRISPR interference\cite{Hsu2018}. Additionally, single cell CRISPR screens would benefit from the adoption of techniques from the increasingly rich literature on scRNA-seq analysis.

The conditional resampling methodology underlying SCEPTRE is extremely flexible, paving the way for its applicability to future iterations of CRISPR screens and beyond. For example, direct capture Perturb-seq\cite{Replogle2020} has recently been proposed in order to improve the efficiency of gRNA detection. However, the capture rate can vary substantially across gRNAs. We look forward to adapting the SCEPTRE framework to the analysis of such new CRISPR screen technologies. Finally, confounding technical factors like those pictured in Figure~\ref{fig:analysis-challenges}c are ubiquitous across genomics; we anticipate the conditional randomization test will complement traditional parametric and nonparametric methods in such contexts as a valuable analysis tool.

\bibliography{library}

\clearpage

\begin{addendum}
 \item We are indebted to Molly Gasperini and Jacob Tome for clarifying several aspects of their data analysis\cite{Gasperini2019}, and to the Shendure lab and Timothy Barry for providing extensive feedback on an earlier draft of this paper. We also thank Tom Norman, Atray Dixit, and Wesley Tansey for useful discussions on single cell CRISPR screens. This work was supported, in part, by National Institute of Mental Health (NIMH) grants R01MH123184 and R37MH057881 as well as SFARI Grant 575547. Part of the data analysis used the Extreme Science and Engineering Discovery Environment (XSEDE)\cite{XSEDE}, which is supported by National Science Foundation grant number ACI-1548562. Specifically, it used the Bridges system\cite{Bridges}, which is supported by NSF award number ACI-1445606, at the Pittsburgh Supercomputing Center (PSC). \item[Competing Interests] The authors declare that they have no
competing financial interests.
 \item[Correspondence] Correspondence and requests for materials
should be addressed to E.K. \\ (email: ekatsevi@wharton.upenn.edu).
\end{addendum}

\section*{Methods}

\paragraph{Reproducing the negative binomial approach of Gasperini et al.}

Consider a particular gene/gRNA pair. For each cell $i = 1, \dots, n$, let $X_i \in \{0,1\}$ indicate whether the gRNA was present in the cell, let $Y_i \in \{0,1,2,\dots\}$ be the gene expression in the cell, defined as the number of unique molecular identifiers (UMIs) from this gene, and let $Z_i \in \mathbb R^d$ be a list of cell-level technical factors.

Gasperini et al. used a negative binomial regression of $Y_i$ on $X_i$ and $Z_i$: 
\begin{equation}
Y_i \overset{\text{ind}} \sim \text{NegBin}(\mu_i, \alpha); \quad \log(\mu_i) = \beta_0 + X_i \beta + Z_i^T \gamma,
\label{negative-binomial}
\end{equation}
with $\alpha$ being the dispersion and $Z_i$ consisting of three cell-level technical factors: the total number of gRNAs in the cell, the percentage of observed transcripts that are mitochondrial, and the sequencing batch. To correct for sequencing depth, Gasperini et al. replaced $Y_i$ in the above regression by rounding $Y_i/s_i$ to the nearest integer, where $s_i$ are DESeq2-style size factors. Raw dispersions $\alpha_{\text{raw}}$ were estimated for each gene based on its sample mean and variance. Gasperini et al fit a mean-dispersion relationship to these raw dispersions, and finally obtained a shrunk estimate of $\alpha$ by projecting $\alpha_{\text{raw}}$ onto this curve (Figure~\ref{fig:analysis-challenges}b). The significance of $X_i$ in the above regression was determined with a likelihood ratio test for $H_0: \beta = 0$. Since enhancer perturbation decreases gene expression, gene-enhancer pairs with $\widehat \beta > 0$ were removed from the analysis.

%Our analysis was based on the conditional permutation test outlined above, using a test statistic $T(X,Y,Z)$ similar to that used by Gasperini et al., but with three modifications: improved dispersion estimation, improved confounder correction, and $z$-statistic instead of likelihood ratio statistic. We discuss each of these next.
%
%Gasperini et al. computed raw dispersion estimates via the method of moments, then fit a mean-dispersion relationship, and then projected the raw dispersions onto this curve (recall Figure~\ref{fig:analysis-challenges}b). We simply removed this projection step, using the raw dispersions instead. We could have used a more sophisticated dispersion estimation scheme (which, if successful, could have boosted our power), but the key feature of our procedure is that its validity does not depend on the expression model---or the dispersion in particular---being correctly specified. 
%
%In addition to the three cell-level covariates $Z_i$ corrected for by Gasperini et al., we also included two others: the total number of UMIs observed and the number of genes with at least one UMI in a given cell. We use these additional covariates to correct for sequencing depth, instead of the size factors used in the original analysis. This strategy is in line with that advocated by a recent work on scRNA-seq analysis\cite{Hafemeister2019}.
%
%The likelihood ratio test employed by Gasperini et al. is a two-sided test; it does not distinguish between positive and negative associations of the gRNA with gene expression. Instead, we used the standard GLM $z$-statistic for the coefficient $\beta$, and measured one-sided departures from the null in the negative direction.

\paragraph{Exploration of sequencing depth confounding.}
%To focus on the issue of sequencing depth confounding, we used raw dispersion estimates (compare the red and blue curves in Figure~\ref{fig:calibration-comparison-all}) as well as one-sided $z$-tests. We then repeated the negative binomial regression analysis underlying Monocle2 with these updated dispersion estimates. These two modifications yielded new negative binomial regression $p$-values for each gene/gRNA pair (Figure~\ref{fig:analysis-challenges}g).

After improving the dispersion estimates, we sought to understand the direction and strength of the remaining miscalibration for each gRNA, reasoning it may relate to sequencing depth confounding. To this end, we tested the residual association between each gRNA and the sequencing depth, after accounting for the total number of gRNAs per cell. In particular, for each gRNA, we ran a logistic regression of the gRNA indicator against the total number of gRNAs and the total number of UMIs detected per cell. We then extracted the $z$-score of the coefficient of the total number of UMIs. We found these agreed well with the direction and strength of the confounding effect, as shown in Figure~\ref{fig:calibration-comparison-all}b.

\paragraph{Conditional randomization test and accelerations.}

Letting $(X, Y, Z) = \{(X_i, Y_i, Z_i)\}_{i = 1}^{n}$, consider any test statistic $T(X,Y,Z)$ measuring the effect of the gRNA on the expression of the gene. The conditional randomization test\cite{CetL16} is based on resampling the gRNA indicators independently for each cell. Letting $\pi_i = \mathbb P[X_i = 1 | Z_i]$, define random variables
\begin{equation}
\widetilde X_i \overset{\text{ind}}\sim \text{Ber}(\pi_i). 
\label{X-tilde-dist}
\end{equation}
Then, the CRT $p$-value is given by
\begin{equation}
p_{\text{CRT}} = \mathbb P[T(\widetilde X, Y, Z) \geq T(X,Y,Z)\mid X,Y,Z].
\label{conditional-randomization}
\end{equation}
This translates to repeatedly sampling $\widetilde X$ from the distribution~\eqref{X-tilde-dist}, recomputing the test statistic with $X$ replaced by $\widetilde X$, and defining the $p$-value as the probability the resampled test statistic exceeds the original. Under the null hypothesis that the gRNA perturbation does not impact the cell (adjusting for technical factors), i.e. $Y \independent X \mid Z$, we obtain a valid $p$-value~\eqref{conditional-randomization}, \textit{regardless of the expression distribution $Y|X,Z$ and regardless of the test statistic $T$}. We choose a test statistic $T$ based on the improved negative binomial regression discussed in the main text, with two computational accelerations.

First, we employed the recently proposed\cite{Liu2020} \textit{distillation} technique to accelerate the recomputation of the negative binomial regression for each resample. The idea is to use a slightly modified test statistic, consisting of two steps:
\begin{enumerate}
	\item Fit $(\widehat \beta_0, \widehat \gamma)$ from the negative binomial regression~\eqref{negative-binomial} except without the gRNA term:
	\begin{equation}
	Y_i \overset{\text{ind}}\sim \textrm{NegBin}(\mu_i, \alpha); \quad \log(\mu_i) = \beta_0 + Z_i^T \gamma.
	\label{negative-binomial-distillation}
	\end{equation}
	\item Fit $\widehat \beta$ from a negative binomial regression with the estimated contributions of $Z_i$ from step 1 as offsets:
	\begin{equation}
	Y_i \overset{\text{ind}}\sim \textrm{NegBin}(\mu_i, \alpha); \quad \log(\mu_i) = X_i \beta + \widehat \beta_0 + Z_i^T \widehat \gamma.
	\label{distilled-negative-binomial}
	\end{equation}
\end{enumerate}
Conditional randomization testing with this two step test statistic, which is nearly identical to the full negative binomial regression~\eqref{negative-binomial}, is much faster. Indeed, since the first step is not a function of $X_i$, it remains the same for each resampled triple $(\widetilde X, Y, Z)$. Therefore, only the second step must be recomputed with each resample, and this step is faster because it involves only a univariate regression.

Next, we accelerated the second step above using the sparsity of the binary vector $(X_1, \dots, X_n)$ (or a resample of it). To do so, we wrote the log-likelihood of the reduced negative binomial regression~\eqref{distilled-negative-binomial} as follows, denoting by $\ell(Y_i, \log(\mu_i))$ the negative binomial log-likelihood:
\begin{equation*}
\begin{split}
\sum_{i = 1}^n \ell(Y_i, X_i \beta + \widehat \beta_0 + Z_i^T \widehat \gamma) &= \sum_{i: X_i = 0} \ell(Y_i, \widehat \beta_0 + Z_i^T \widehat \gamma) + \sum_{i: X_i = 1} \ell(Y_i, \beta + \widehat \beta_0 + Z_i^T \widehat \gamma) \\
&= C + \sum_{i: X_i = 1} \ell(Y_i, \beta + \widehat \beta_0 + Z_i^T \widehat \gamma).
\end{split}
\end{equation*}
This simple calculation shows that, up to a constant that does not depend on $\beta$, the negative binomial log-likelihood corresponding to the model~\eqref{distilled-negative-binomial} is the same as that corresponding to the model with only intercept and offset term for those cells with a gRNA:
\begin{equation}
Y_i \overset{\text{ind}}\sim \textrm{NegBin}(\mu_i, \alpha); \quad \log(\mu_i) = \beta + \widehat \beta_0 + Z_i^T \widehat \gamma, \quad \text{for } i \text{ such that } X_i = 1.
\label{distilled-negative-binomial-accelerated}
\end{equation}
The above negative binomial regression is therefore equivalent to equation~\eqref{distilled-negative-binomial}, but much faster to compute, because it involves only the thousand or so cells containing the gRNA instead of the 200,000 total cells.

\paragraph{SCEPTRE methodology.}

In practice, we must estimate the gRNA probabilities $\pi_i$ as well as the $p$-value $p_{\text{CRT}}$. This is because usually we do not know the distribution $X|Z$, and cannot compute the conditional probability in equation~\eqref{conditional-randomization} exactly. We propose to estimate $\pi_i$ via logistic regression of $X$ on $Z$, and to estimate $p_{\text{CRT}}$ by resampling $\widetilde X$ a large number of times and then fitting a skew-$t$ distribution to the resampling null distribution $T(\widetilde X, Y, Z)|X,Y,Z$. We outline SCEPTRE below: 
\begin{enumerate}
	\item Fit technical factor effects $(\widehat \beta_0,\widehat \gamma)$ on gene expression using the negative binomial regression~\eqref{negative-binomial-distillation}.
	\item Extract a $z$-score $z(X,Y,Z)$ from the reduced negative binomial regression~\eqref{distilled-negative-binomial-accelerated}.
	\item Assume that 
	\begin{equation}
	X_i \overset{\text{ind}}\sim \text{Ber}(\pi_i); \quad \log\left(\frac{\pi_i}{1-\pi_i}\right) = \tau_0 + Z_i^T \tau
	\label{logistic-regression}
	\end{equation}
	for $\tau_0 \in \mathbb R$ and $\tau \in \mathbb R^d$, and fit $(\widehat \tau_0, \widehat \tau)$ via logistic regression of $X$ on $Z$. Then, extract the fitted probabilities $\widehat \pi_i = (1+\exp(-(\widehat \tau_0 + Z_i^T \widehat \tau)))^{-1}$.	
	\item For $b = 1, \dots, B$,
	\begin{itemize}
		\item Resample the gRNA assignments based on the probabilities $\widehat \pi_i$ to obtain $\widetilde X^b$~\eqref{X-tilde-dist}.
		\item Extract a $z$-score $z(\widetilde X^b, Y, Z)$ from the reduced negative binomial regression~\eqref{distilled-negative-binomial-accelerated}.
	\end{itemize}
	\item Fit a skew-$t$ distribution $\widehat F_{\text{null}}$ to the resampled $z$-scores $\{z(\widetilde X^b, Y, Z)\}_{b = 1}^B$. 
	\item Return the $p$-value $\widehat p_{\text{SCEPTRE}} = \mathbb P[\widehat F_{\text{null}} \leq z(X,Y,Z)]$.
\end{enumerate}
In our data analysis, we used $B = 500$ resamples. Following Gasperini et al, we based our analysis on the top two gRNAs targeting each enhancer. Some enhancers were also targeted with two additional gRNAs, but we excluded these from the analysis.

\paragraph{Numerical simulation to assess calibration.}

We simulated one gene $Y_i$, one gRNA $X_i$, and one confounder $Z_i$ in $n = 1000$ cells. We drew $Z_i \overset{\text{i.i.d.}}\sim N(0,1)$. Then, we drew $X_i$ based on the logistic model~\eqref{logistic-regression}, with $\tau = 4$ and $\tau_0 = \log \frac{0.05}{0.95}$, the latter to make the probability of gRNA occurrence about $0.05$ on average across cells. Finally, we drew the gene expression $Y_i$ from the following zero-inflated negative binomial model:
\begin{equation*}
Y_i \overset{\text{ind}}\sim \lambda \delta_0 + (1-\lambda)\text{NegBin}(\mu_i, \alpha), \quad \log(\mu_i) = \beta_0 + 4Z_i.
\end{equation*}
Note that for any $\beta_0,\lambda,\alpha$, the gRNA does not impact the gene expression in this setup. We chose $\beta_0 = \log(5)$ to make the average gene expression about 5. The four settings shown in Figure~\ref{fig:calibration}a correspond to 
\begin{equation*}
(\lambda_1, \alpha_1) = (0,1); \quad (\lambda_2, \alpha_2) = (0,5); \quad (\lambda_3, \alpha_3) = (0,0.2); \quad (\lambda_4, \alpha_4) = (0.25,1).
\end{equation*}
For the first, the negative binomial model is correctly specified. For the second and third, the dispersion estimate of 1 is too small and too large, respectively. The last setting exhibits zero inflation.

We applied three methods to these four problem settings, each with 500 repetitions. The negative binomial method was based on the $z$ statistic from the standard negative binomial regression~\eqref{negative-binomial} with $\alpha = 1$. The permutation method was implemented the same as SPECTRE, except skipping step 3 and defining $\widetilde X^b$ to be a random permutation of $X$. Both the permutation method and SPECTRE used $B = 250$ resamples.

\paragraph{Definition of Gasperini et al. discovery set.}

Gasperini et al. reported a total of 664 gene-enhancer pairs, identifying 470 of these as ``high-confidence." We chose to use the latter set, rather than the former, for all our comparisons. Gasperini et al. carried out their ChIP-seq and HI-C enrichment analyses only on the high-confidence discoveries, so for those comparisons we do the same. Furthermore, the 664 total gene-enhancer pairs reported in the original analysis were the result of a Benjamini-Hochberg FDR correction that included not only the candidate enhancers but also hundreds of positive controls. While Bonferroni corrections can only become more conservative when including more hypotheses, BH corrections are known to become anticonservative when extra positive controls are included\cite{FR01}. To avoid this extra risk of false positives, we chose to use the ``high-confidence" set throughout.

\paragraph{ChIP-seq, HI-C enrichment analyses.}

These analyses (see Figures~\ref{fig:application}e-f and \ref{fig:chipseq-enrichment-details}) were carried out almost exactly following Gasperini et al. The only change we made is in our quantification of the ChIP-seq enrichment (Figure~\ref{fig:application}f). We use the odds ratio of a candidate enhancer being paired to a gene, comparing the top and bottom ChIP-seq quintiles.

%\paragraph{Methods used in calibration comparison.}
%
%In Figure~\ref{fig:calibration-comparison-all}, we compared the calibration of the proposed approach to three alternatives. The first, ``Original NB", is the same as Gasperini et al.'s negative binomial regression, except with a one-sided $z$-statistic instead of a likelihood ratio statistic. The second ``Improved NB", is the same except with improved confounder correction and dispersion estimation. The third, ``Marginal permutation", is the same as conditional permutation, except the gRNAs are permuted uniformly among cells rather than conditionally on the their covariates.

\section*{Data availability}

Analysis results are available online at \url{https://bit.ly/SCEPTRE}. All analysis was performed on publicly available data. The CRISPR screen data\cite{Gasperini2019} is available at \url{https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE120861}. The ChIP-seq data are drawn from the ENCODE project\cite{Dunham2012} and are available at \url{https://www.encodeproject.org/}. The HI-C enrichment analysis is based on the data from Rao et al\cite{Rao2014}, available at \url{https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE63525}. The eQTL and eRNA co-expression $p$-values are taken from the GeneHancer database\cite{Fishilevich2017}, available as part of GeneCards (\url{https://www.genecards.org/}).

\section*{Code availability}

Code to reproduce all data analysis is available on Github at \url{https://github.com/ekatsevi/SCEPTRE}.

\clearpage

\appendix
%\newgeometry{top=1in}
\thispagestyle{empty} 
\addtocounter{page}{-1}

\counterwithin{figure}{section}
\section{Supplementary Figures}
\renewcommand{\thesection}{S}

\vspace{0.25in}

%\clearpage

%\makeatletter 
%\renewcommand{\thefigure}{S\@arabic\c@figure}
%\makeatother

%\section*{Supplementary Figures}
\thispagestyle{empty}
%\vspace{-0.1in}
%\setcounter{figure}{0}    
%\renewcommand{\thefigure}{S\arabic{figure}}

\begin{figure}[h!]
	\centering
	\includegraphics[width = 0.75\textwidth]{figures/FigureS1/FigureS1.png}
	\caption{Sequencing depth (total mRNA UMIs per cell) impacts gRNA detection for Mosaic-seq data\cite{Xie2017}, as was demonstrated for Gasperini et al. data in Figure~\ref{fig:analysis-challenges}b.}
	\label{fig:mosaicseq}
\end{figure}

\clearpage

\thispagestyle{empty} 
\begin{figure}[h!]
	\includegraphics[width = \textwidth]{figures/FigureS2/FigureS2.png}
	\caption{\textbf{Comparing negative binomial and conditional resampling $p$-values based on the same test statistic.} \textbf{a,} The standard parametric negative binomial $p$-value versus that obtained from the same test statistic by conditional resampling, for each gene / candidate enhancer pair (both truncated at $10^{-10}$ for visualization). The two can diverge fairly substantially. \textbf{b-d,} Parametric and conditional resampling null distributions for the negative binomial $z$-value in three cases: the parametric $p$-value is more significant (b), the conditional resampling $p$-value is more significant (c), the two $p$-values are about the same (d).}
	\label{fig:NB-vs-CP}
\end{figure}

%\restoregeometry

\clearpage
\thispagestyle{empty} 
\begin{figure}[h!]
	\includegraphics[width = \textwidth]{figures/FigureS3/FigureS3.pdf}
	\caption{\textbf{Details on ChIP-seq enrichment analysis.} Fraction of candidate enhancers paired with a gene, broken down by quintile of ChIP-seq signal (0 means the candidate enhancer did not overlap a ChIP-seq peak), based on which the odds ratios in Figure~\ref{fig:application}f were computed. Both methods generally pair candidate enhancers in higher ChIP-seq quintiles more frequently, but this enrichment is more pronounced in SCEPTRE across all eight ChIP-seq targets.}
	\label{fig:chipseq-enrichment-details}
\end{figure}

\end{document}

### 2020-11-18 sceptre_paper/plotting/Figure1.R

@@ -1,7 +1,6 @@
 source("load_data_for_plotting.R")
 source("../katsevich2020/R/plotting_functs.R")
 
-
 # QQ plot of Gasperini NTC p-values
 
 ci = 0.95
@@ -65,6 +64,7 @@ p = df %>% arrange(desc(KS_pval)) %>% mutate(KS_pval = ifelse(KS_pval < 1e-5, 1e
                      legend.background = element_rect(fill = "transparent", colour = NA))
 plot(p)
 
+covariates_xie$guide_count = rowSums(grna_indicator_matrix_xie)
 p = covariates_xie %>% 
   mutate(total_umis = 10^(log_n_umis)) %>%
   ggplot(aes(x = total_umis, y = guide_count)) +

### 2020-11-18 sceptre_paper/plotting/Figure2.R

@@ -1,0 +1,12 @@
+#######################################################
+#
+# Reproduce Figure 2c from Katsevich and Roeder (2020).
+#
+#######################################################
+
+gene_to_plot = "ENSG00000135390"
+grna_to_plot = "chr12.1893_top_two"
+p = plot_skew_t(gene_to_plot, grna_to_plot) + theme(plot.title = element_blank())
+plot(p)
+ggsave(filename = sprintf("%s/Figure2c.png", figures_dir), plot = p, device = "png", 
+       width = 3.5, height = 2.5)

### 2020-11-18 sceptre_paper/plotting/Figure3.R

@@ -1,0 +1,275 @@
+#######################################################
+#
+# Reproduce Figure 3 from Katsevich and Roeder (2020).
+#
+#######################################################
+
+# collate the NTC p-values of four methods
+df_NTC = rbind(resampling_results %>% 
+                 filter(site_type == "NTC", method == "conditional_randomization") %>%
+                 select(gene_id, grna_group, corrected_pvalue_st) %>% 
+                 rename(pvalue = corrected_pvalue_st) %>%
+                 mutate(method = "Conditional randomization"),
+               resampling_results %>% 
+                 filter(site_type == "NTC", method == "marginal_permutation") %>%
+                 select(gene_id, grna_group, corrected_pvalue_st) %>% 
+                 rename(pvalue = corrected_pvalue_st) %>%
+                 mutate(method = "Marginal permutation"),
+               likelihood_results %>%
+                 filter(site_type == "NTC", method == "raw_2") %>%
+                 select(gene_id, grna_group, pvalue) %>%
+                 mutate(method = "Improved NB"),
+               likelihood_results %>%
+                 filter(site_type == "NTC", method == "shrunk_1") %>%
+                 select(gene_id, grna_group, pvalue) %>%
+                 mutate(method = "Original NB")) %>%
+  mutate(method = factor(method, levels = c("Original NB", "Improved NB", 
+                                            "Marginal permutation", "Conditional randomization")))
+
+ci = 0.95
+# compute information for overall QQ plot
+df1 = df_NTC %>% 
+  group_by(method) %>% 
+  mutate(r = rank(pvalue), expected = ppoints(n())[r], 
+         clower = qbeta(p=(1-ci)/2, shape1 = r, shape2 = n()+1-r),
+         cupper = qbeta(p=(1+ci)/2, shape1 = r, shape2 = n()+1-r)) %>%
+  ungroup()
+
+# compute information for QQ plot by gRNA
+df2 = df_NTC %>%   
+  group_by(grna_group, method) %>% 
+  mutate(r = rank(pvalue), expected = ppoints(n())[r], 
+         clower = qbeta(p=(1-ci)/2, shape1 = r, shape2 = n()+1-r),
+         cupper = qbeta(p=(1+ci)/2, shape1 = r, shape2 = n()+1-r)) %>% 
+  ungroup() %>%  
+  group_by(r, expected, clower, cupper, method) %>% 
[…]

### 2020-11-18 sceptre_paper/plotting/Figure4.R

@@ -1,0 +1,211 @@
+#######################################################
+#
+# Reproduce Figure 4 from Katsevich and Roeder (2020).
+#
+#######################################################
+
+# a: DHS and NTC p-values
+ci = 0.95
+subsampling_factor = 250
+p = 
+  resampling_results %>%
+  filter(site_type %in% c("DHS", "NTC"), method == "conditional_randomization") %>%
+  rename(pvalue = corrected_pvalue_st) %>%
+  mutate(site_type = factor(site_type, levels = c("DHS", "NTC"), labels = c("Candidate enhancer", "Negative control"))) %>%
+  group_by(site_type) %>% 
+  mutate(r = rank(pvalue), expected = ppoints(n())[r], 
+         clower = qbeta(p=(1-ci)/2, shape1 = r, shape2 = n()+1-r),
+         cupper = qbeta(p=(1+ci)/2, shape1 = r, shape2 = n()+1-r)) %>% 
+  ungroup() %>%
+  mutate(clower = ifelse(site_type == "Candidate enhancer", NA, clower), 
+         cupper = ifelse(site_type == "Candidate enhancer", NA, cupper)) %>%
+  mutate(pvalue = ifelse(pvalue < 1e-20, 0, pvalue)) %>%
+  filter(-log10(expected) > 2 | (site_type == "Candidate enhancer" & row_number() %% 20 == 0) | row_number() %% subsampling_factor == 0) %>%
+  ggplot(aes(x = expected, y = pvalue, group = site_type, ymin = clower, ymax = cupper)) + 
+  geom_point(aes(colour = site_type), size = 3, alpha = 0.5) + 
+  geom_ribbon(alpha = 0.2) + 
+  geom_abline(intercept = 0, slope = 1) +
+  scale_colour_manual(values = c("firebrick3", "slategray"), name = "Perturbation target") + 
+  scale_x_continuous(trans = revlog_trans(base = 10)) + scale_y_continuous(trans = revlog_trans(base = 10)) +
+  xlab(expression(paste("Expected null p-value for gene-enhancer pair"))) + 
+  ylab(expression(paste("SCEPTRE p-value"))) + 
+  theme_bw() + theme(legend.position = c(0.225,0.8),
+                     legend.key = element_rect(colour = "transparent", fill = "transparent"),
+                     legend.background = element_rect(colour = "transparent", fill = "transparent"),
+                     panel.grid = element_blank(), 
+                     panel.border = element_blank(),
+                     axis.line = element_line())
+plot(p)
[…]

### 2020-11-18 sceptre_paper/plotting/Figure5.R

@@ -1,0 +1,93 @@
+df = original_results %>% 
+  rename(old_rejected = rejected, old_pvalue = pvalue.empirical) %>%
+  filter(site_type == "DHS", quality_rank_grna == "top_two") %>%
+  left_join(resampling_results %>% 
+              filter(method == "conditional_randomization") %>%
+              select(gene_id, grna_group, rejected) %>% 
+              rename(new_rejected = rejected), 
+            by = c("gene_id", "grna_group")) %>%
+  mutate(strand = ifelse(target_gene.start == TSS, "+", "-"),
+         enhancer_location = 0.5*(target_site.start + target_site.stop),
+         TSS_dist = ifelse(strand == "+", enhancer_location - TSS, TSS - enhancer_location))
+
+df1 = df %>% filter(old_rejected) %>% mutate(group = "old_rejections")
+df2 = df %>% filter(new_rejected) %>% mutate(group = "new_rejections")
+
+p = rbind(df1, df2) %>%
+  filter(TSS_dist <= 0) %>%
+  mutate(group = factor(group, 
+                        levels = c("new_rejections", "old_rejections"),
+                        labels = c("SCEPTRE", "Original"))) %>%
+  mutate(TSS_dist = TSS_dist/1000) %>%
+  ggplot(aes(x = TSS_dist, group = group, fill = group)) + 
+  geom_histogram(position = position_identity(), alpha = 0.5) + 
+  scale_fill_manual(values = c("blue", "red")) +
+  scale_x_continuous(expand = c(0,0)) + scale_y_continuous(expand = c(0,0)) +
+  xlab("Enhancer to TSS distance (kb)") +
+  theme_bw() + theme(panel.grid = element_blank(), 
+                     panel.border = element_blank(),
+                     axis.line = element_line(),
+                     legend.title = element_blank(),
+                     legend.position = c(0.2,0.8)) 
+plot(p)
+ggsave(filename = sprintf("%s/Figure5/TSS_dist_histogram.pdf", figures_dir), plot = p, device = "pdf",
+       width = 4, height = 3)
+
+rbind(
+  df1 %>% 
+    filter(TSS_dist <= 0) %>% 
+    summarise(`Mean distance (kb)` = mean(TSS_dist)/1000, 
[…]

### 2020-11-18 sceptre_paper/plotting/FigureS5.R

@@ -1,0 +1,206 @@
+#######################################################
+#
+# Reproduce Figure S5 from Katsevich and Roeder (2020).
+#
+#######################################################
+
+# collate NTC p-values from four negative binomial variants
+df_NTC = likelihood_results %>% 
+  filter(site_type == "NTC") %>% 
+  mutate(method = ifelse(method == "raw_1", 
+                         "NB (disp)",
+                         ifelse(method == "raw_2",
+                                "NB (both)",
+                                ifelse(method == "shrunk_1",
+                                       "NB (orig)", "NB (conf)")))) %>%
+  select(gene_id, grna_group, pvalue, method)
+
+# compute information for overall QQ plot
+ci = 0.95
+df1 = df_NTC %>% 
+  group_by(method) %>% 
+  mutate(r = rank(pvalue), expected = ppoints(n())[r], 
+         clower = qbeta(p=(1-ci)/2, shape1 = r, shape2 = n()+1-r),
+         cupper = qbeta(p=(1+ci)/2, shape1 = r, shape2 = n()+1-r)) %>%
+  ungroup()
+
+# compute information for QQ plot by gene
+df2 = df_NTC %>%   
+  group_by(gene_id, method) %>% 
+  mutate(r = rank(pvalue), expected = ppoints(n())[r], 
+         clower = qbeta(p=(1-ci)/2, shape1 = r, shape2 = n()+1-r),
+         cupper = qbeta(p=(1+ci)/2, shape1 = r, shape2 = n()+1-r)) %>% 
+  ungroup() %>%  
+  group_by(r, expected, clower, cupper, method) %>% 
+  summarise(pmedian = median(pvalue), plower = quantile(pvalue, 0.025), pupper = quantile(pvalue, 0.975)) %>% 
+  ungroup()
+
+# compute information for QQ plot by gRNA
+df3 = df_NTC %>%   
+  group_by(grna_group, method) %>% 
+  mutate(r = rank(pvalue), expected = ppoints(n())[r], 
+         clower = qbeta(p=(1-ci)/2, shape1 = r, shape2 = n()+1-r),
+         cupper = qbeta(p=(1+ci)/2, shape1 = r, shape2 = n()+1-r)) %>% 
+  ungroup() %>%  
+  group_by(r, expected, clower, cupper, method) %>% 
[…]

### 2020-11-18 sceptre_paper/plotting/FigureS6.R

@@ -1,0 +1,58 @@
+#######################################################
+#
+# Reproduce Figure S6 from Katsevich and Roeder (2020).
+#
+#######################################################
+
+# Collate candidate enhancer p-values from negative binomial and 
+# conditional randomization approaches
+df = likelihood_results %>% 
+  filter(site_type == "DHS", method == "raw_2") %>% 
+  select(gene_id, grna_group, pvalue, zvalue) %>% 
+  dplyr::rename(old_pvalue = pvalue, old_zvalue = zvalue) %>%
+  left_join(resampling_results %>% 
+              filter(site_type == "DHS", method == "conditional_randomization") %>% 
+              select(gene_id, grna_group, corrected_pvalue_st, original_zvalue, xi, alpha, omega, nu) %>% 
+              dplyr::rename(new_pvalue = corrected_pvalue_st, new_zvalue = original_zvalue), 
+            by = c("grna_group", "gene_id"))
+
+# three gene/gRNA pairs to illustrate the differences in the two approaches
+genes_to_plot = c("ENSG00000124657","ENSG00000145592","ENSG00000065268")
+grnas_to_plot = c("chr6.1215_top_two", "chr5.1161_top_two","chr19.272_top_two")
+
+# panel a: comparing the two ways of computing p-values
+p0 = df %>% 
+  mutate(highlighted = 
+           (gene_id == genes_to_plot[1] & grna_group == grnas_to_plot[1]) |
+           (gene_id == genes_to_plot[2] & grna_group == grnas_to_plot[2]) |
+           (gene_id == genes_to_plot[3] & grna_group == grnas_to_plot[3])) %>%
+  arrange(highlighted) %>%
+  filter(old_pvalue > 1e-10, new_pvalue > 1e-10, !is.na(alpha), alpha < 4) %>% 
+  ggplot(aes(x = old_pvalue, y = new_pvalue, colour = highlighted)) + 
+  geom_point() + geom_abline(slope = 1, linetype = "dashed") +
+  scale_x_log10() + scale_y_log10() + 
+  xlab("Negative binomial p-value") + ylab("SCEPTRE p-value") +
+  scale_colour_manual(values = c("dodgerblue", "red")) + 
+  theme_bw() + theme(legend.position = "none",
+                     panel.grid = element_blank(), 
+                     panel.border = element_blank(),
[…]

### 2020-11-18 sceptre_paper/plotting/FigureS7.R

@@ -1,0 +1,78 @@
+#######################################################
+#
+# Reproduce Figure S7 from Katsevich and Roeder (2020).
+#
+#######################################################
+
+# compute KS p-values
+KS_pvals = original_results %>% 
+  filter(site_type == "NTC") %>% 
+  group_by(gene_id) %>% 
+  summarise(KS_pval = ks.test(pvalue.empirical, "punif")$p.value) %>% ungroup() 
+
+# collect several pieces of information about each gene-gRNA pair
+thresh_old = original_results %>% filter(rejected) %>% summarise(max(pvalue.empirical)) %>% pull()
+df = original_results %>% 
+  dplyr::rename(old_rejected = rejected, old_pvalue = pvalue.empirical) %>%
+  left_join(grna_confounding, by = "grna_group") %>% 
+  left_join(KS_pvals, by = "gene_id") %>% 
+  filter(site_type == "DHS", quality_rank_grna == "top_two") %>%
+  left_join(resampling_results %>% 
+              filter(method == "conditional_randomization") %>%
+              select(gene_id, grna_group, rejected) %>% 
+              dplyr::rename(new_rejected = rejected), 
+            by = c("gene_id", "grna_group")) %>%
+  left_join(rejected_pairs_HIC %>% 
+              mutate(different_TADs = is.na(TAD_left)) %>% 
+              select(gene_id, target_site, different_TADs), 
+            by = c("gene_id", "target_site")) %>% 
+  arrange((old_rejected + new_rejected) %% 2, old_rejected + new_rejected) %>%
+  mutate(reason_not_rejected = ifelse(outlier_gene, 
+                                      "Outlier gene", 
+                                      ifelse(beta > 0, 
+                                             "Positive effect", 
+                                             ifelse(!old_rejected & old_pvalue <= thresh_old, "Low confidence", "none"))),
+         reason_not_rejected = factor(reason_not_rejected, levels = c("none", "Positive effect", "Outlier gene", "Low confidence")),
+         rejected_by = ifelse(old_rejected, 
+                              ifelse(new_rejected, "Both methods", "Original only"), 
+                              ifelse(new_rejected, "SCEPTRE only", "Neither method")),
+         rejected_by = factor(rejected_by, levels = c("Both methods", "SCEPTRE only",
[…]

### 2020-11-18 sceptre_paper/plotting/load_data_for_plotting.R

@@ -1,11 +1,25 @@
-offsite_dir = "/Users/ekatsevi/Dropbox (Penn)/project-files/SCEPTRE"
+offsite_dir = "/Users/ekatsevi/Dropbox (Penn)/project-files/sceptre_files"
 
-original_results = read_tsv(sprintf("%s/data/raw/CRISPR/GSE120861_all_deg_results.at_scale.txt", offsite_dir))
-resampling_results = read.fst(sprintf("%s/resampling_results.fst", offsite_dir)) %>% as_tibble()
-likelihood_results = read.fst(sprintf("%s/likelihood_results.fst", offsite_dir)) %>% as_tibble()
-
-covariates_xie = read.fst(sprintf("%s/cell_covariate_model_matrix_xie.fst", offsite_dir)) %>% as_tibble()
+# Gasperini results
+original_results_gasp = read_tsv(sprintf("%s/data/gasperini/raw/GSE120861_all_deg_results.at_scale.txt", offsite_dir))
+resampling_results_gasp = read.fst(sprintf("%s/results/gasperini/sceptre/resampling_results.fst", offsite_dir)) %>% as_tibble()
+likelihood_results_gasp = read.fst(sprintf("%s/results/gasperini/negative_binomial/likelihood_results.fst", offsite_dir)) %>% as_tibble()
 covariates_gasp = read.fst(sprintf("%s/cell_covariate_model_matrix_gasp.fst", offsite_dir)) %>% as_tibble()
 
+# Xie results
+original_results_xie = readRDS(sprintf("%s/data/xie/processed/xie_p_values.rds", offsite_dir))
+resampling_results_xie = read.fst(sprintf("%s/results/xie/sceptre/all_results_with_names.fst", offsite_dir)) %>% as_tibble()
+resampling_results_xie = read.fst(sprintf("%s/results/xie/sceptre/all_results_with_names.fst", offsite_dir)) %>% as_tibble()
+covariates_xie = read.fst(sprintf("%s/cell_covariate_model_matrix_xie.fst", offsite_dir)) %>% as_tibble()
 grna_indicator_matrix_xie = read.fst(sprintf("%s/grna_indicator_matrix_xie.fst", offsite_dir)) %>% as_tibble()
-covariates_xie$guide_count = rowSums(grna_indicator_matrix_xie)
+
+# simulation results
+simulation_results = read.fst(sprintf("%s/results/simulations/all_results.fst", offsite_dir)) %>% as_tibble()
+
+# enrichment results
[…]

### 2020-11-18 sceptre_paper/analysis_drivers/analysis_drivers_gasp/check_directory_structure_1.R
reset the run_sceptre_at_scale.bash file

@@ -8,7 +8,7 @@ for (package in packages) suppressPackageStartupMessages(library(package, charac
 # Hardcode the directories to create.
 sub_dirs <- c(create_parent_directories("data/gasperini/raw"), create_parent_directories("data/gasperini/precomp/gRNA"), "data/gasperini/precomp/gene", "data/gasperini/processed",
   create_parent_directories("data/functional"), "data/functional/HIC", "data/functional/ChIP-seq", "data/functional/GeneHancer",
-  create_parent_directories("results/gasperini/sceptre"), "results/gasperini/negative_binomial",
+  create_parent_directories("results/gasperini/sceptre"), "results/gasperini/negative_binomial", "results/gasperini/enrichment",
   "figures", create_parent_directories("logs/gasperini")) %>% unique()
 
 dirs_to_create <- paste0(offsite_dir, "/", sub_dirs)

### 2020-11-18 sceptre_paper/plotting/Figure3.R
updates to plotting functions

@@ -5,7 +5,7 @@
 #######################################################
 
 # collate the NTC p-values of four methods
-df_NTC = rbind(resampling_results %>% 
+df_NTC = rbind(resampling_results %>%
                  filter(site_type == "NTC", method == "conditional_randomization") %>%
                  select(gene_id, grna_group, corrected_pvalue_st) %>% 
                  rename(pvalue = corrected_pvalue_st) %>%

### 2020-11-18 sceptre_paper/plotting/Figure5.R

@@ -1,93 +1,0 @@
-df = original_results %>% 
-  rename(old_rejected = rejected, old_pvalue = pvalue.empirical) %>%
-  filter(site_type == "DHS", quality_rank_grna == "top_two") %>%
-  left_join(resampling_results %>% 
-              filter(method == "conditional_randomization") %>%
-              select(gene_id, grna_group, rejected) %>% 
-              rename(new_rejected = rejected), 
-            by = c("gene_id", "grna_group")) %>%
-  mutate(strand = ifelse(target_gene.start == TSS, "+", "-"),
-         enhancer_location = 0.5*(target_site.start + target_site.stop),
-         TSS_dist = ifelse(strand == "+", enhancer_location - TSS, TSS - enhancer_location))
-
-df1 = df %>% filter(old_rejected) %>% mutate(group = "old_rejections")
-df2 = df %>% filter(new_rejected) %>% mutate(group = "new_rejections")
-
-p = rbind(df1, df2) %>%
-  filter(TSS_dist <= 0) %>%
-  mutate(group = factor(group, 
-                        levels = c("new_rejections", "old_rejections"),
-                        labels = c("SCEPTRE", "Original"))) %>%
-  mutate(TSS_dist = TSS_dist/1000) %>%
-  ggplot(aes(x = TSS_dist, group = group, fill = group)) + 
-  geom_histogram(position = position_identity(), alpha = 0.5) + 
-  scale_fill_manual(values = c("blue", "red")) +
-  scale_x_continuous(expand = c(0,0)) + scale_y_continuous(expand = c(0,0)) +
-  xlab("Enhancer to TSS distance (kb)") +
-  theme_bw() + theme(panel.grid = element_blank(), 
-                     panel.border = element_blank(),
-                     axis.line = element_line(),
-                     legend.title = element_blank(),
-                     legend.position = c(0.2,0.8)) 
-plot(p)
-ggsave(filename = sprintf("%s/Figure5/TSS_dist_histogram.pdf", figures_dir), plot = p, device = "pdf",
-       width = 4, height = 3)
-
-rbind(
-  df1 %>% 
-    filter(TSS_dist <= 0) %>% 
-    summarise(`Mean distance (kb)` = mean(TSS_dist)/1000, 
[…]

### 2020-11-18 sceptre_paper/plotting/FigureS3.R

@@ -1,6 +1,6 @@
 #######################################################
 #
-# Reproduce Figure S8 from Katsevich and Roeder (2020).
+# Reproduce Figure S3 from Katsevich and Roeder (2020).
 #
 #######################################################
 
@@ -21,4 +21,4 @@ p = paired_fractions %>%
         axis.line = element_line()
   )
 plot(p)
-ggsave(plot = p, filename = sprintf("%s/figures/FigureS8.pdf", base_dir), width = 7, height = 4)
+ggsave(plot = p, filename = sprintf("%s/figures/FigureS3/FigureS3.pdf", base_dir), width = 7, height = 4)

### 2020-11-18 sceptre_paper/plotting/FigureS5.R

@@ -1,206 +1,0 @@
-#######################################################
-#
-# Reproduce Figure S5 from Katsevich and Roeder (2020).
-#
-#######################################################
-
-# collate NTC p-values from four negative binomial variants
-df_NTC = likelihood_results %>% 
-  filter(site_type == "NTC") %>% 
-  mutate(method = ifelse(method == "raw_1", 
-                         "NB (disp)",
-                         ifelse(method == "raw_2",
-                                "NB (both)",
-                                ifelse(method == "shrunk_1",
-                                       "NB (orig)", "NB (conf)")))) %>%
-  select(gene_id, grna_group, pvalue, method)
-
-# compute information for overall QQ plot
-ci = 0.95
-df1 = df_NTC %>% 
-  group_by(method) %>% 
-  mutate(r = rank(pvalue), expected = ppoints(n())[r], 
-         clower = qbeta(p=(1-ci)/2, shape1 = r, shape2 = n()+1-r),
-         cupper = qbeta(p=(1+ci)/2, shape1 = r, shape2 = n()+1-r)) %>%
-  ungroup()
-
-# compute information for QQ plot by gene
-df2 = df_NTC %>%   
-  group_by(gene_id, method) %>% 
-  mutate(r = rank(pvalue), expected = ppoints(n())[r], 
-         clower = qbeta(p=(1-ci)/2, shape1 = r, shape2 = n()+1-r),
-         cupper = qbeta(p=(1+ci)/2, shape1 = r, shape2 = n()+1-r)) %>% 
-  ungroup() %>%  
-  group_by(r, expected, clower, cupper, method) %>% 
-  summarise(pmedian = median(pvalue), plower = quantile(pvalue, 0.025), pupper = quantile(pvalue, 0.975)) %>% 
-  ungroup()
-
-# compute information for QQ plot by gRNA
-df3 = df_NTC %>%   
-  group_by(grna_group, method) %>% 
-  mutate(r = rank(pvalue), expected = ppoints(n())[r], 
-         clower = qbeta(p=(1-ci)/2, shape1 = r, shape2 = n()+1-r),
-         cupper = qbeta(p=(1+ci)/2, shape1 = r, shape2 = n()+1-r)) %>% 
-  ungroup() %>%  
-  group_by(r, expected, clower, cupper, method) %>% 
[…]

### 2020-11-18 sceptre_paper/plotting/load_data_for_plotting.R

@@ -4,14 +4,14 @@ offsite_dir = "/Users/ekatsevi/Dropbox (Penn)/project-files/sceptre_files"
 original_results_gasp = read_tsv(sprintf("%s/data/gasperini/raw/GSE120861_all_deg_results.at_scale.txt", offsite_dir))
 resampling_results_gasp = read.fst(sprintf("%s/results/gasperini/sceptre/resampling_results.fst", offsite_dir)) %>% as_tibble()
 likelihood_results_gasp = read.fst(sprintf("%s/results/gasperini/negative_binomial/likelihood_results.fst", offsite_dir)) %>% as_tibble()
-covariates_gasp = read.fst(sprintf("%s/cell_covariate_model_matrix_gasp.fst", offsite_dir)) %>% as_tibble()
+covariates_gasp = read.fst(sprintf("%s/data/gasperini/processed/cell_covariate_model_matrix.fst", offsite_dir)) %>% as_tibble()
 
 # Xie results
 original_results_xie = readRDS(sprintf("%s/data/xie/processed/xie_p_values.rds", offsite_dir))
 resampling_results_xie = read.fst(sprintf("%s/results/xie/sceptre/all_results_with_names.fst", offsite_dir)) %>% as_tibble()
-resampling_results_xie = read.fst(sprintf("%s/results/xie/sceptre/all_results_with_names.fst", offsite_dir)) %>% as_tibble()
-covariates_xie = read.fst(sprintf("%s/cell_covariate_model_matrix_xie.fst", offsite_dir)) %>% as_tibble()
-grna_indicator_matrix_xie = read.fst(sprintf("%s/grna_indicator_matrix_xie.fst", offsite_dir)) %>% as_tibble()
+likelihood_results_xie = read.fst(sprintf("%s/results/xie/negative_binomial/all_results.fst", offsite_dir)) %>% as_tibble()
+covariates_xie = read.fst(sprintf("%s/data/xie/processed/covariate_model_matrix.fst", offsite_dir)) %>% as_tibble()
+grna_indicator_matrix_xie = read.fst(sprintf("%s/data/xie/processed/grna_indicator_matrix.fst", offsite_dir)) %>% as_tibble()
 
 # simulation results
 simulation_results = read.fst(sprintf("%s/results/simulations/all_results.fst", offsite_dir)) %>% as_tibble()
@@ -23,3 +23,9 @@ TF_enrichments = read_tsv(sprintf("%s/results/gasperini/enrichment/TF_enrichment
[…]

### 2020-11-18 sceptre_paper/plotting/FigureS1.R
updates to plotting functions

@@ -1,0 +1,138 @@
+#######################################################
+#
+# Reproduce Figure S1 from Katsevich and Roeder (2020).
+#
+#######################################################
+
+# a: dispersion estimation
+KS_pvals = original_results %>% 
+  filter(site_type == "NTC") %>% 
+  group_by(gene_id, outlier_gene) %>% 
+  summarise(KS_pval = ks.test(pvalue.empirical, "punif")$p.value) %>%
+  ungroup()
+df = disp_table %>% 
+  rename(disp_raw = disp) %>% 
+  mutate(disp_shrunk = disp_coeffs[1] + disp_coeffs[2]/mu) %>% 
+  inner_join(KS_pvals, by = "gene_id") %>%
+  mutate(KS_pval_adj = p.adjust(KS_pval, "fdr"))
+problematic_genes = df %>% filter(KS_pval_adj < 0.1) %>% pull(gene_id)
+p = df %>% arrange(desc(KS_pval)) %>% mutate(KS_pval = ifelse(KS_pval < 1e-5, 1e-5, KS_pval)) %>% 
+  mutate(outlier_gene = factor(outlier_gene, levels = c(FALSE, TRUE), 
+                               labels = c("Non-\"outlier\" gene", "\"Outlier\" gene"))) %>%
+  ggplot() + geom_point(aes(x = mu, y = disp_raw, colour = KS_pval, shape = outlier_gene)) + 
+  geom_point(aes(x = 1.29, y = 3.98), colour = "red") + 
+  geom_point(aes(x = 0.964, y = 0.304), colour = "red") + 
+  geom_point(aes(x = mu, y = disp_raw), size = 4, shape = 21, data = df %>% filter(KS_pval < 0.05/num_genes)) + 
+  geom_line(aes(x = mu, y = disp_coeffs[1] + disp_coeffs[2]/mu), colour = "black", linetype = "dashed") + 
+  scale_colour_continuous(trans = "log10", name = "Empirical p-value\nmiscalibration, per gene") +
+  scale_x_log10() + scale_y_log10() + 
+  xlab("Mean gene expression") + ylab("Dispersion of gene expression") + 
+  guides(colour = guide_colorbar(title.position = "top", direction = "horizontal", order = 1)) +
+  guides(shape = guide_legend(title = NULL, direction = "vertical", order = 2)) +
+  theme_bw() + theme(panel.grid = element_blank(), legend.position = c(0.65, 0.175),
+                     panel.border = element_blank(), axis.line = element_line(),
[…]

### 2020-11-18 sceptre_paper/plotting/Figure1.R
committing to merge

@@ -31,6 +31,7 @@ plot(p)
 # QQ plot of Xie ARL15 p-values
 
 
+
 # Dispersion estimation
 
 KS_pvals = original_results %>% 

### 2020-11-19 sceptre_paper/plotting/Figure1.R
update plotting files

@@ -1,5 +1,6 @@
-source("load_data_for_plotting.R")
-source("../katsevich2020/R/plotting_functs.R")
+args <- commandArgs(trailingOnly = TRUE)
+code_dir <- if (is.na(args[1])) "/Users/timbarry/Box/SCEPTRE/sceptre_paper" else args[1]
+source(paste0(code_dir, "/sceptre_paper/plotting/load_data_for_plotting.R"))
 
 # QQ plot of Gasperini NTC p-values
 ci = 0.95

### 2020-11-19 sceptre_paper/plotting/Figure3.R

@@ -1,9 +1,92 @@
-#######################################################
-#
 # Reproduce Figure 3 from Katsevich and Roeder (2020).
-#
-#######################################################
+args <- commandArgs(trailingOnly = TRUE)
+code_dir <- if (is.na(args[1])) "/Users/timbarry/Box/SCEPTRE/sceptre_paper" else args[1]
+require(katsevich2020)
+source(paste0(code_dir, "/sceptre_paper/plotting/load_data_for_plotting.R"))
 
+# subfigure a: simulation results
+truncate_thresh <- 1e-7
+qq_data <- simulation_results %>%
+  group_by(method, dataset_id) %>%
+  mutate(r = rank(p_value, ties.method = "first"), expected = ppoints(n())[r],
+         clower = qbeta(p=(1-ci)/2, shape1 = r, shape2 = n()+1-r),
+         cupper = qbeta(p=(1+ci)/2, shape1 = r, shape2 = n()+1-r)) %>%
+  ungroup() %>%
+  mutate(pvalue = ifelse(p_value < truncate_thresh, truncate_thresh, p_value),
+         facet_label = factor(x = as.character(dataset_id), levels = c("2", "1", "3", "4"), labels = c("Correct model", "Dispersion too large", "Dispersion too small", "Zero inflation")),
+         method = factor(x = as.character(method), levels = c("sceptre", "negative_binomial", "scMAGeCK"), labels = c("SCEPTRE", "Fixed dispersion NB", "scMAGeCK")))
+
+p <- qq_data %>%
+  ggplot(aes(x = expected, col = method, y = pvalue, ymin = clower, ymax = cupper)) +
+  geom_ribbon(alpha = 0.2, color = NA) +
+  geom_abline(intercept = 0, slope = 1) +
+  geom_point(size = 1, alpha = 0.5) +
+  xlab("Expected null p-value") +
+  ylab("Observed p-value") +
+  scale_colour_manual(values = setNames(plot_colors[c("sceptre", "fixed_dispersion_nb", "scMAGeCK")], NULL), name = "Method") +
+  guides(color = guide_legend(override.aes = list(alpha = 1))) +
+  scale_x_continuous(trans = revlog_trans(base = 10)) +
+  scale_y_continuous(trans = revlog_trans(base = 10)) +
+  facet_wrap(.~facet_label, nrow = 1) +
+  theme_bw() + theme(
+    panel.spacing.x = unit(1.25, "lines"),
[…]

### 2020-11-19 sceptre_paper/plotting/Tims_plotting/arl15_enh_qqplots_1.R

@@ -10,7 +10,6 @@ p_vals_hypergeo <- paste0(processed_dir, "/xie_p_values.rds") %>% readRDS() %>% 
 p_vals_hypergeo <- p_vals_hypergeo[names(p_vals_hypergeo) %in% p_vals_sceptre$gene_id]
 to_plot <- tibble(method = rep(x = c("SCEPTRE", "NB regression", "Hypergeometric"), each = length(p_vals_hypergeo)) %>% factor(), p_value = c(p_vals_sceptre$p_value, p_vals_nb$p_value, set_names(p_vals_hypergeo, NULL) ), gene = c(as.character(p_vals_sceptre$gene_id), as.character(p_vals_nb$gene_id), names(p_vals_hypergeo)))
 
-ci <- 0.95
 truncate_thresh <- 1e-9
 qq_data <- to_plot %>%
   rename(pvalue = p_value) %>%

### 2020-11-19 sceptre_paper/plotting/Tims_plotting/simulation_qqplot_4.R

@@ -4,10 +4,8 @@ offsite_dir <- if (is.na(args[2])) "/Volumes/tims_new_drive/research/sceptre_fil
 require(katsevich2020)
 source(paste0(code_dir, "/sceptre_paper/plotting/aux_objects.R"))
 
-sim_res <- paste0(offsite_dir, "/results/simulations/all_results.fst") %>% read.fst()
-ci <- 0.95
-truncate_thresh <- 1e-9
-qq_data <- sim_res %>%
+truncate_thresh <- 1e-7
+qq_data <- simulation_results %>%
   group_by(method, dataset_id) %>%
   mutate(r = rank(p_value), expected = ppoints(n())[r],
          clower = qbeta(p=(1-ci)/2, shape1 = r, shape2 = n()+1-r),
@@ -15,17 +13,16 @@ qq_data <- sim_res %>%
   ungroup() %>%
   mutate(pvalue = ifelse(p_value < truncate_thresh, truncate_thresh, p_value),
          facet_label = factor(x = as.character(dataset_id), levels = c("2", "1", "3", "4"), labels = c("Correct model", "Dispersion too large", "Dispersion too small", "Zero inflation")),
-         method = factor(x = as.character(method), levels = c("sceptre", "negative_binomial", "scMAGeCK"), labels = c("SCEPTRE", "Original NB", "scMAGeCK")))
-
+         method = factor(x = as.character(method), levels = c("sceptre", "negative_binomial", "scMAGeCK"), labels = c("SCEPTRE", "Fixed dispersion NB", "scMAGeCK")))
 
 p <- qq_data %>%
   ggplot(aes(x = expected, col = method, y = pvalue, ymin = clower, ymax = cupper)) +
   geom_ribbon(alpha = 0.2, color = NA) +
-  geom_point(size = 1, alpha = 0.5) +
   geom_abline(intercept = 0, slope = 1) +
+  geom_point(size = 1, alpha = 0.5) +
   xlab("Expected null p-value") +
   ylab("Observed p-value") +
-  scale_colour_manual(values = setNames(plot_colors[c("original_nb", "sceptre", "scMAGeCK")], NULL), name = "Method") +
+  scale_colour_manual(values = setNames(plot_colors[c("sceptre", "fixed_dispersion_nb", "scMAGeCK")], NULL), name = "Method") +
   guides(color = guide_legend(override.aes = list(alpha = 1))) +
[…]

### 2020-11-19 sceptre_paper/plotting/load_data_for_plotting.R

@@ -1,4 +1,5 @@
-offsite_dir = "/Users/ekatsevi/Dropbox (Penn)/project-files/sceptre_files"
+offsite_dir <- if (is.na(args[2])) "/Volumes/tims_new_drive/research/sceptre_files" else args[2]
+manuscript_figure_dir <- paste0(code_dir, "/sceptre_paper/manuscript/figures")
 
 # Gasperini results
 original_results_gasp = read_tsv(sprintf("%s/data/gasperini/raw/GSE120861_all_deg_results.at_scale.txt", offsite_dir))
@@ -29,3 +30,7 @@ disp_coeffs = as.numeric(readRDS(sprintf("%s/data/gasperini/processed/disp_coeff
 
 # dispersion table (mean expressions and dispersion estimates for each gene)
 disp_table = read.fst(sprintf("%s/data/gasperini/processed/disp_table.fst", offsite_dir)) %>% as_tibble()
+
+# additional items
+ci <- 0.95
+plot_colors <- c(gasperini_nb = "red", hf_nb = "violet", fixed_dispersion_nb = "darkslategray4", sceptre = "royalblue4", hypergeometric = "turquoise3", scMAGeCK = "magenta4")

### 2020-11-19 sceptre_paper/katsevich2020/R/scMAGeCK_functions.R
increase B for scMAGeCK to 1000

@@ -178,11 +178,10 @@ run_scMaGECK_analysis <- function(expression_matrix, gRNA_indic_matrix) {
   Xmat = mat_for_single_reg[[1]]
   Ymat = mat_for_single_reg[[2]]
   LAMBDA <- 0.01
-  Amat_pm_lst = getsolvedmatrix_with_permutation_cell_label(Xmat, Ymat, lambda = LAMBDA, npermutation = 500)
+  Amat_pm_lst = getsolvedmatrix_with_permutation_cell_label(Xmat, Ymat, lambda = LAMBDA, npermutation = 1000)
   Amat_pval = Amat_pm_lst[[2]]
   return(Amat_pval)
 }
-
 
 
 #' Run scMAGeCK simulation

### 2020-11-19 sceptre_paper/plotting/Figure3.R
fix plotting scripts

@@ -16,7 +16,7 @@ qq_data <- simulation_results %>%
          facet_label = factor(x = as.character(dataset_id), levels = c("2", "1", "3", "4"), labels = c("Correct model", "Dispersion too large", "Dispersion too small", "Zero inflation")),
          method = factor(x = as.character(method), levels = c("sceptre", "negative_binomial", "scMAGeCK"), labels = c("SCEPTRE", "Fixed dispersion NB", "scMAGeCK")))
 
-p <- qq_data %>%
+p_a <- qq_data %>%
   ggplot(aes(x = expected, col = method, y = pvalue, ymin = clower, ymax = cupper)) +
   geom_ribbon(alpha = 0.2, color = NA) +
   geom_abline(intercept = 0, slope = 1) +
@@ -66,7 +66,7 @@ for (i in 1:length(sorted_methods_rightmost_pnt)) {
 annotation_df <- filter(qq_data, gene == filter(resampling_results_xie, gene_names == "ARL15") %>% pull(gene_id) %>% as.character())
 arrow_coords <- tibble(x1 = 4e-4, x2 = annotation_df$expected + 1e-5, y1 = 1e-8, y2 = annotation_df$pvalue)
 
-p <- qq_data %>%
+p_e <- qq_data %>%
   ggplot(aes(x = expected, y = pvalue, group = method, ymin = clower, ymax = cupper)) +
   geom_point(aes(color = method), size = 1, alpha = 0.5) +
   geom_ribbon(alpha = 0.2) +
@@ -85,6 +85,7 @@ p <- qq_data %>%
   annotate(geom = "text", x = 10e-4, y = 1e-8, label = "ARL15", col = "firebrick3") +
   geom_segment(aes(x = x1, y = y1, xend = x2, yend = y2), lwd = 0.2, data = arrow_coords, arrow = arrow(length=unit(0.2,"cm")), inherit.aes = FALSE, col = "grey40")
 
+# 
 
 if (FALSE) {
 # collate the NTC p-values of four methods

### 2020-11-19 sceptre_paper/plotting/load_data_for_plotting.R

@@ -29,12 +29,8 @@ paired_fractions = read_tsv(sprintf("%s/results/gasperini/enrichment/TF_paired_e
 disp_coeffs = as.numeric(readRDS(sprintf("%s/data/gasperini/processed/disp_coefficients.rds", offsite_dir)))
 
 # dispersion table (mean expressions and dispersion estimates for each gene)
-<<<<<<< HEAD
 disp_table = read.fst(sprintf("%s/data/gasperini/processed/disp_table.fst", offsite_dir)) %>% as_tibble()
 
 # additional items
 ci <- 0.95
 plot_colors <- c(gasperini_nb = "red", hf_nb = "violet", fixed_dispersion_nb = "darkslategray4", sceptre = "royalblue4", hypergeometric = "turquoise3", scMAGeCK = "magenta4")
-=======
-disp_table = read.fst(sprintf("%s/data/gasperini/processed/disp_table.fst", offsite_dir)) %>% as_tibble()
->>>>>>> 0eb5fe180f8149fdbfeac19983d2f2596dceac86

### 2020-11-22 sceptre_package/sceptre/R/deprecated.R
figures 1 - 3

@@ -12,7 +12,6 @@
 #' @param log_dir file path to the directory in which to sink the log output
 #'
 #' @return NULL
-#' @export
 #'
 #' @examples
 #' pod_id <- 1
@@ -50,3 +49,71 @@ run_gene_precomputation_at_scale <- function(pod_id, gene_precomp_dir, cell_gene
   saveRDS(out_sizes, size_save_fp)
   if (!is.null(log_dir)) deactivate_sink()
 }
+
+#' Run sceptre using precomputations for gRNAs and genes (old version)
+#'
+#' This function is the workhorse function of the sceptre package. It runs a distilled CRT using a negative binomial test statistic based on an expression vector, a gRNA indicator vector, an offset vector (from the distillation step), gRNA conditional probabilities, an estimate of the negative binomial size parameter, and the number of resampling replicates.
+#'
+#' This currently is a one-tailed, left-sided test. Thus, it is suitable for up-regulatory elements like enhancers and promoters but not down-regulatory elements like silencers.
+#'
+#' @param expressions a vector of gene expressions (in UMI counts)
+#' @param gRNA_indicators a vector of gRNA indicators
+#' @param gRNA_precomp a vector of conditional probabilities for gRNA assignments
+#' @param gene_precomp_size the pre-computed size
+#' @param gene_precomp_offsets the pre-computed distillation offsets
+#' @param B the number of resamples to make (default 500)
+#' @param seed an arguement to set.seed; if null, no seed is set
+#'
+#' @return a p-value of the null hypothesis of no gRNA effect on gene expression
+run_sceptre_using_precomp_old <- function(expressions, gRNA_indicators, gRNA_precomp, gene_precomp_size, gene_precomp_offsets, B, seed, reduced_output = TRUE) {
+  if (!is.null(seed)) set.seed(seed)
+
+  # compute the test statistic on the real data
+  fit_star <- vglm(formula = expressions[gRNA_indicators == 1] ~ 1, family = negbinomial.size(gene_precomp_size), offset = gene_precomp_offsets[gRNA_indicators == 1])
+  t_star <- summaryvglm(fit_star)@coef3["(Intercept)", "z value"]
+
+  # resample B times
+  t_nulls <- sapply(1:B, function(i) {
[…]

### 2020-11-22 sceptre_package/sceptre/R/resampling_method_functions.R

@@ -105,75 +105,6 @@ run_sceptre_using_precomp <- function(expressions, gRNA_indicators, gRNA_precomp
 }
 
 
-#' Run sceptre using precomputations for gRNAs and genes (old version)
-#'
-#' This function is the workhorse function of the sceptre package. It runs a distilled CRT using a negative binomial test statistic based on an expression vector, a gRNA indicator vector, an offset vector (from the distillation step), gRNA conditional probabilities, an estimate of the negative binomial size parameter, and the number of resampling replicates.
-#'
-#' This currently is a one-tailed, left-sided test. Thus, it is suitable for up-regulatory elements like enhancers and promoters but not down-regulatory elements like silencers.
-#'
-#' @param expressions a vector of gene expressions (in UMI counts)
-#' @param gRNA_indicators a vector of gRNA indicators
-#' @param gRNA_precomp a vector of conditional probabilities for gRNA assignments
-#' @param gene_precomp_size the pre-computed size
-#' @param gene_precomp_offsets the pre-computed distillation offsets
-#' @param B the number of resamples to make (default 500)
-#' @param seed an arguement to set.seed; if null, no seed is set
-#'
-#' @export
-#' @return a p-value of the null hypothesis of no gRNA effect on gene expression
-run_sceptre_using_precomp_old <- function(expressions, gRNA_indicators, gRNA_precomp, gene_precomp_size, gene_precomp_offsets, B, seed, reduced_output = TRUE) {
-  if (!is.null(seed)) set.seed(seed)
-
-  # compute the test statistic on the real data
-  fit_star <- vglm(formula = expressions[gRNA_indicators == 1] ~ 1, family = negbinomial.size(gene_precomp_size), offset = gene_precomp_offsets[gRNA_indicators == 1])
-  t_star <- summaryvglm(fit_star)@coef3["(Intercept)", "z value"]
-
-  # resample B times
-  t_nulls <- sapply(1:B, function(i) {
-    if (i %% 100 == 0) cat(paste0("Running resample ", i ,"/", B, ".\n"))
-    gRNA_indicators_null <- rbinom(n = length(gRNA_precomp), size = 1, prob = gRNA_precomp)
-    tryCatch({
[…]

### 2020-11-22 sceptre_paper/analysis_drivers/analysis_drivers_gasp/run_enrichment_analysis_6.R

@@ -2,7 +2,7 @@ args <- commandArgs(trailingOnly = TRUE)
 code_dir <- if (is.na(args[1])) "/Users/timbarry/Box/SCEPTRE/sceptre_paper/" else args[1]
 require(plyranges)
 require(GenomicRanges)
-require(katsevich2020)
+require(sceptre)
 source(paste0(code_dir, "/sceptre_paper/analysis_drivers/analysis_drivers_gasp/file_paths_to_dirs.R"))
 
 # Define a couple directories
@@ -10,9 +10,8 @@ results_dir_enrichment <- paste0(offsite_dir, "/results/gasperini/enrichment")
 functional_data_dir <- paste0(offsite_dir, "/data/functional/")
 
 # Read in the association results
-original_results = paste0(processed_dir, "/original_results.fst") %>% read.fst()
-resampling_results = paste0(results_dir, "/resampling_results.fst") %>% read.fst()
-original_results <- filter(original_results, pair_id %in% resampling_results$pair_id)
+original_results <- paste0(processed_dir, "/original_results.fst") %>% read.fst()
+resampling_results <- paste0(results_dir, "/resampling_results.fst") %>% read.fst()
 
 # ChIP-seq enrichment analysis
 
@@ -134,12 +133,12 @@ if (!file.exists(sprintf("%s/rejected_pairs_HIC.tsv", results_dir_enrichment))) 
     filter(site_type == "DHS", quality_rank_grna == "top_two") %>%
     select(chr, gene_id, target_gene.start, target_gene.stop, TSS,
            target_site, target_site.start, target_site.stop, rejected) %>%
-    rename(rejected_old = rejected) %>%
+    dplyr::rename(rejected_old = rejected) %>%
     left_join(resampling_results %>%
                 filter(site_type == "DHS",
                        quality_rank_grna == "top_two") %>%
                 select(gene_id,  target_site, rejected) %>%
-                rename(rejected_new = rejected),
+                dplyr::rename(rejected_new = rejected),
               by = c("gene_id", "target_site"))
 
   all_enhancers = all_pairs %>% select(target_site, chr, target_site.start, target_site.stop) %>% unique()

### 2020-11-22 sceptre_paper/katsevich2020/R/plotting_functs.R

@@ -19,3 +19,91 @@ revlog_trans <- function(base = exp(1)) {
             domain = c(1e-100, Inf) ## The domain over which the transformation is valued
   )
 }
+
+
+#' Plot skew-t given gene and gRNA.
+#'
+#' @param gene_id id of gene
+#' @param gRNA_id id of gRNA
+#' @param sceptre_function_args path to the sceptre_function_args.R file. Defaults to the location on Tim's machine.
+#' @param offsite_dirpath to the offsite dir. Defaults to the location on Tim's machine.
+#'
+#' @return a ggplot containing the plot
+#' @export
+#'
+#' @examples
+#' gene_id <- "ENSG00000008256"
+#' gRNA_id <- "ACTB_TSS"
+#' plot_skew_t_gene_gRNA(gene_id, gRNA_id)
+plot_skew_t_gene_gRNA <- function(gene_id, gRNA_id, sceptre_function_args = "/Users/timbarry/Box/SCEPTRE/sceptre_paper/sceptre_paper/analysis_drivers/analysis_drivers_gasp/sceptre_function_args.R", offsite_dir = "/Volumes/tims_new_drive/research/sceptre_files") {
+  info_pack <- get_all_data_for_gene_gRNA(gene_id, gRNA_id, sceptre_function_args, offsite_dir)
+  sceptre_res <- run_sceptre_using_precomp(expressions = info_pack$expressions, gRNA_indicators = info_pack$gRNA_indicators, gRNA_precomp = info_pack$gRNA_precomp, gene_precomp_size = info_pack$gene_precomp_size, gene_precomp_offsets = info_pack$gene_precomp_offsets, B = 500,seed = 1234, reduced_output = FALSE)
+  p <- plot_skew_t(resampled_zvalues = sceptre_res$resampled_z_values, original_zvalue = sceptre_res$z_value, dp = sceptre_res$skew_t_mle)
+  return(list(plot = p, p_val = sceptre_res$p_value, skew_t_fit_success = sceptre_res$skew_t_fit_success))
+}
+
+
+#' Plot skew-t
+#'
+#' @param resampled_zvalues the set of resampled z-values
+#' @param original_zvalue the z value of the original negative binomial fit
+#' @param dpv the skew-t fit MLE
+#'
+#' @return a ggplot object containing the plot
+plot_skew_t <- function(resampled_zvalues, original_zvalue, dp) {
+  z = seq(-4, 4, length.out = 1000)
[…]

### 2020-11-22 sceptre_paper/plotting/Figure1.R

@@ -1,87 +1,143 @@
+# Reproduce Figure 1 from Katsevich, Barry, and Roeder (2020).
 args <- commandArgs(trailingOnly = TRUE)
 code_dir <- if (is.na(args[1])) "/Users/timbarry/Box/SCEPTRE/sceptre_paper" else args[1]
+require(katsevich2020)
+require(ggpubr)
+require(cowplot)
 source(paste0(code_dir, "/sceptre_paper/plotting/load_data_for_plotting.R"))
+fig1_dir <- paste0(manuscript_figure_dir, "/Figure1")
 
-# QQ plot of Gasperini NTC p-values
-ci = 0.95
-subsampling_factor = 250
-p = original_results %>%
+# Subfigure a: Gasperini NTC p-values and Xie ARL15 p-value qqplots
+p_thresh <- 1e-21
+pa.1 <- original_results_gasp %>%
   rename(pvalue = pvalue.raw) %>%
   filter(site_type == "NTC", beta < 0) %>%
-  mutate(r = rank(pvalue), expected = ppoints(n())[r], 
+  mutate(r = rank(pvalue), expected = ppoints(n())[r],
          clower = qbeta(p=(1-ci)/2, shape1 = r, shape2 = n()+1-r),
-         cupper = qbeta(p=(1+ci)/2, shape1 = r, shape2 = n()+1-r)) %>% 
-  # mutate(pvalue = ifelse(pvalue < 1e-50, 0, pvalue)) %>%
-  filter(-log10(expected) > 2 | row_number() %% subsampling_factor == 0) %>%
-  ggplot(aes(x = expected, y = pvalue, ymin = clower, ymax = cupper)) + 
-  geom_point(aes(colour = site_type), size = 3, alpha = 0.5) + 
-  geom_ribbon(alpha = 0.2) + 
+         cupper = qbeta(p=(1+ci)/2, shape1 = r, shape2 = n()+1-r)) %>%
+  filter(-log10(expected) > 2 | row_number() %% subsampling_factor == subsampling_factor) %>%
+  mutate(pvalue = ifelse(pvalue < p_thresh, p_thresh, pvalue)) %>%
+  ggplot(aes(x = expected, y = pvalue, ymin = clower, ymax = cupper)) +
+  geom_point(col = plot_colors[["gasperini_nb"]], size = 2, alpha = 0.5) +
+  geom_ribbon(alpha = 0.2) +
   geom_abline(intercept = 0, slope = 1) +
[…]

### 2020-11-22 sceptre_paper/plotting/Figure2.R

@@ -1,12 +1,14 @@
-#######################################################
-#
-# Reproduce Figure 2c from Katsevich and Roeder (2020).
-#
-#######################################################
+# Reproduce Figure 2 from Katsevich, Barry, and Roeder (2020).
+args <- commandArgs(trailingOnly = TRUE)
+code_dir <- if (is.na(args[1])) "/Users/timbarry/Box/SCEPTRE/sceptre_paper" else args[1]
+offsite_dir <- if (is.na(args[2])) "/Volumes/tims_new_drive/research/sceptre_files" else args[2]
+fig2_dir <- paste0(code_dir, "/sceptre_paper/manuscript/figures/Figure2")
+require(katsevich2020)
 
-gene_to_plot = "ENSG00000135390"
-grna_to_plot = "chr12.1893_top_two"
-p = plot_skew_t(gene_to_plot, grna_to_plot) + theme(plot.title = element_blank())
-plot(p)
-ggsave(filename = sprintf("%s/Figure2c.png", figures_dir), plot = p, device = "png", 
-       width = 3.5, height = 2.5)
+# Select the gene and gRNA to use in the plot
+gene_to_plot <- "ENSG00000135390"
+grna_to_plot <- "chr12.1893_top_two"
+sceptre_function_args <- paste0(code_dir, "/sceptre_paper/analysis_drivers/analysis_drivers_gasp/sceptre_function_args.R")
+p <- plot_skew_t_gene_gRNA(gene_id = gene_to_plot, gRNA_id = grna_to_plot, sceptre_function_args = sceptre_function_args, offsite_dir = offsite_dir)
+plot(p$plot)
+ggsave(filename = paste0(fig2_dir, "/Figure2c.png"), plot = p$plot, device = "png", width = 3.5, height = 2.5, dpi = "retina")

### 2020-11-22 sceptre_paper/plotting/Figure3.R

@@ -2,17 +2,19 @@
 args <- commandArgs(trailingOnly = TRUE)
 code_dir <- if (is.na(args[1])) "/Users/timbarry/Box/SCEPTRE/sceptre_paper" else args[1]
 require(katsevich2020)
+require(cowplot)
 source(paste0(code_dir, "/sceptre_paper/plotting/load_data_for_plotting.R"))
+fig3_dir <- paste0(manuscript_figure_dir, "/Figure3")
 
 # subfigure a: simulation results
-truncate_thresh <- 1e-7
+p_thresh <- 1e-8
 qq_data <- simulation_results %>%
   group_by(method, dataset_id) %>%
   mutate(r = rank(p_value, ties.method = "first"), expected = ppoints(n())[r],
          clower = qbeta(p=(1-ci)/2, shape1 = r, shape2 = n()+1-r),
          cupper = qbeta(p=(1+ci)/2, shape1 = r, shape2 = n()+1-r)) %>%
   ungroup() %>%
-  mutate(pvalue = ifelse(p_value < truncate_thresh, truncate_thresh, p_value),
+  mutate(pvalue = ifelse(p_value < p_thresh, p_thresh, p_value),
          facet_label = factor(x = as.character(dataset_id), levels = c("2", "1", "3", "4"), labels = c("Correct model", "Dispersion too large", "Dispersion too small", "Zero inflation")),
          method = factor(x = as.character(method), levels = c("sceptre", "negative_binomial", "scMAGeCK"), labels = c("SCEPTRE", "Fixed dispersion NB", "scMAGeCK")))
 
@@ -23,7 +25,8 @@ p_a <- qq_data %>%
   geom_point(size = 1, alpha = 0.5) +
   xlab("Expected null p-value") +
   ylab("Observed p-value") +
-  scale_colour_manual(values = setNames(plot_colors[c("sceptre", "fixed_dispersion_nb", "scMAGeCK")], NULL), name = "Method") +
+  ggtitle("Simulated negative control pair") +
+  scale_colour_manual(values = setNames(plot_colors[c("sceptre", "hf_nb", "scMAGeCK")], NULL), name = "Method") +
   guides(color = guide_legend(override.aes = list(alpha = 1))) +
   scale_x_continuous(trans = revlog_trans(base = 10)) +
   scale_y_continuous(trans = revlog_trans(base = 10)) +
@@ -32,20 +35,53 @@ p_a <- qq_data %>%
[…]

### 2020-11-22 sceptre_paper/plotting/Figure4.R

@@ -1,47 +1,44 @@
-#######################################################
-#
-# Reproduce Figure 4 from Katsevich and Roeder (2020).
-#
-#######################################################
+# Reproduce Figure 4 from Katsevich, Barry, and Roeder (2020).
+
 
 # a: DHS and NTC p-values
 ci = 0.95
 subsampling_factor = 250
-p = 
+p =
   resampling_results %>%
   filter(site_type %in% c("DHS", "NTC"), method == "conditional_randomization") %>%
   rename(pvalue = corrected_pvalue_st) %>%
   mutate(site_type = factor(site_type, levels = c("DHS", "NTC"), labels = c("Candidate enhancer", "Negative control"))) %>%
-  group_by(site_type) %>% 
-  mutate(r = rank(pvalue), expected = ppoints(n())[r], 
+  group_by(site_type) %>%
+  mutate(r = rank(pvalue), expected = ppoints(n())[r],
          clower = qbeta(p=(1-ci)/2, shape1 = r, shape2 = n()+1-r),
-         cupper = qbeta(p=(1+ci)/2, shape1 = r, shape2 = n()+1-r)) %>% 
+         cupper = qbeta(p=(1+ci)/2, shape1 = r, shape2 = n()+1-r)) %>%
   ungroup() %>%
-  mutate(clower = ifelse(site_type == "Candidate enhancer", NA, clower), 
+  mutate(clower = ifelse(site_type == "Candidate enhancer", NA, clower),
          cupper = ifelse(site_type == "Candidate enhancer", NA, cupper)) %>%
   mutate(pvalue = ifelse(pvalue < 1e-20, 0, pvalue)) %>%
   filter(-log10(expected) > 2 | (site_type == "Candidate enhancer" & row_number() %% 20 == 0) | row_number() %% subsampling_factor == 0) %>%
-  ggplot(aes(x = expected, y = pvalue, group = site_type, ymin = clower, ymax = cupper)) + 
-  geom_point(aes(colour = site_type), size = 3, alpha = 0.5) + 
-  geom_ribbon(alpha = 0.2) + 
+  ggplot(aes(x = expected, y = pvalue, group = site_type, ymin = clower, ymax = cupper)) +
+  geom_point(aes(colour = site_type), size = 3, alpha = 0.5) +
+  geom_ribbon(alpha = 0.2) +
   geom_abline(intercept = 0, slope = 1) +
[…]

### 2020-11-22 sceptre_paper/plotting/load_data_for_plotting.R

@@ -2,7 +2,7 @@ offsite_dir <- if (is.na(args[2])) "/Volumes/tims_new_drive/research/sceptre_fil
 manuscript_figure_dir <- paste0(code_dir, "/sceptre_paper/manuscript/figures")
 
 # Gasperini results
-original_results_gasp = read_tsv(sprintf("%s/data/gasperini/raw/GSE120861_all_deg_results.at_scale.txt", offsite_dir))
+original_results_gasp = paste0(offsite_dir, "/data/gasperini/processed/original_results.fst") %>% read.fst()
 resampling_results_gasp = read.fst(sprintf("%s/results/gasperini/sceptre/resampling_results.fst", offsite_dir)) %>% as_tibble()
 likelihood_results_gasp = read.fst(sprintf("%s/results/gasperini/negative_binomial/likelihood_results.fst", offsite_dir)) %>% as_tibble()
 covariates_gasp = read.fst(sprintf("%s/data/gasperini/processed/cell_covariate_model_matrix.fst", offsite_dir)) %>% as_tibble()
@@ -11,6 +11,7 @@ covariates_gasp = read.fst(sprintf("%s/data/gasperini/processed/cell_covariate_m
 original_results_xie = readRDS(sprintf("%s/data/xie/processed/xie_p_values.rds", offsite_dir))
 resampling_results_xie = read.fst(sprintf("%s/results/xie/sceptre/all_results_with_names.fst", offsite_dir)) %>% as_tibble()
 likelihood_results_xie = read.fst(sprintf("%s/results/xie/negative_binomial/all_results.fst", offsite_dir)) %>% as_tibble()
+p_vals_bulk <- paste0(offsite_dir, "/results/xie/bulk_rna_seq/pvals_arl15_enh.rds") %>% readRDS() %>% as_tibble() %>% rename(gene_names = gene_id)
 covariates_xie = read.fst(sprintf("%s/data/xie/processed/covariate_model_matrix.fst", offsite_dir)) %>% as_tibble()
 grna_indicator_matrix_xie = read.fst(sprintf("%s/data/xie/processed/grna_indicator_matrix.fst", offsite_dir)) %>% as_tibble()
 
@@ -33,4 +34,5 @@ disp_table = read.fst(sprintf("%s/data/gasperini/processed/disp_table.fst", offs
 
 # additional items
 ci <- 0.95
[…]

### 2020-11-22 sceptre_paper/manuscript/manuscript_v3.tex

@@
 This analysis challenge distinguishes CRISPR screens from traditional differential expression applications.
-\textbf{d-g}, Collapsing dispersion estimates to fitted mean-variance relationship (d) leads to over-correction for some genes (e) and under-correction for others (f, g).
-\textbf{d,} Dispersion estimation procedure employed leads to miscalibration for high-dispersion genes, which the empirical correction does not adequately correct for, as measured by KS test applied to empirical $p$-values per gene (point colors).
-\textbf{e,} Raw $p$-values already well-calibrated for LUC7L2 gene, so empirical correction unnecessarily shrinks the significance of the association with TSS-targeting gRNA, depicted by horizontal lines, by three orders of magnitude.
-\textbf{f,} Empirical correction not strong enough for HIST1H1D, which is among circled genes in panel d, which have an NTC-based miscalibration $p$-value smaller than the Bonferroni threshold.
-\textbf{g}, Under-correction leads to two potential false discoveries for HIST1H1D.
-Dashed horizontal line represents the multiple testing threshold.}
+}
 	\label{fig:analysis-challenges}
@@
 	\centering
-	\includegraphics[width = 0.75\textwidth]{figures/FigureS1/FigureS1.png}
-	\caption{Sequencing depth (total mRNA UMIs per cell) impacts gRNA detection for Mosaic-seq data\cite{Xie2017}, as was demonstrated for Gasperini et al. data in Figure~\ref{fig:analysis-challenges}b.}
+	\includegraphics[width = \textwidth]{figures/FigureS1/FigureS1.png}
+	\caption{Gasperini et al.'s empirical correction is insufficient to correct for miscalibration.
+\textbf{a,} Dispersion estimation procedure employed leads to miscalibration for high-dispersion genes, which the empirical correction does not adequately correct for, as measured by KS test applied to empirical $p$-values per gene (point colors).
+\textbf{b,} Raw $p$-values already well-calibrated for LUC7L2 gene, so empirical correction unnecessarily shrinks the significance of the association with TSS-targeting gRNA, depicted by horizontal lines, by three orders of magnitude.
+\textbf{c,} Empirical correction not strong enough for HIST1H1D, which is among circled genes in panel a, which have an NTC-based miscalibration $p$-value smaller than the Bonferroni threshold.
+\textbf{d}, Under-correction leads to two potential false discoveries for HIST1H1D.
+Dashed horizontal line represents the multiple testing threshold.}
 	\label{fig:mosaicseq}

### 2020-11-28 sceptre_paper/manuscript/manuscript_v3.tex
update manuscript

@@
 \newcommand\independent{\perp \!\!\! \perp}
-\usepackage{chngcntr}
+% \usepackage{chngcntr}
 
@@
 For example, technical factors like sequencing depth impact not only expression measurement but also perturbation detection, creating a confounding effect.
-We demonstrate on a recent large-scale single cell CRISPR screen how these challenges cause calibration issues among existing analysis methods.
+We demonstrate on two recent large-scale single cell CRISPR screens how these challenges cause calibration issues among existing analysis methods.
 To address these challenges, we propose SCEPTRE: analysis of single cell perturbation screens via conditional resampling.
 This methodology, designed to avoid calibration issues due to technical confounders and expression model misspecification, infers associations between perturbations and expression by resampling the former according to a working model for perturbation detection probability in each cell.
-Applying SCEPTRE to the CRISPR screen data yields excellent calibration on negative control perturbations and 217 regulatory relationships not found in the original study, many of which are supported by existing functional data.
+Applying SCEPTRE to the CRISPR screen data yields excellent calibration on negative control perturbations and 200 regulatory relationships not found in the original study, many of which are supported by existing functional data.
 \end{abstract}
@@
 
-To address these challenges, we propose SCEPTRE (analysis of Single CEll PerTurbation screens via conditional REsampling).
+To address these challenges, we propose SCEPTRE (analysis of Single CEll PerTurbation screens via conditional REsampling; pronounced ``scepter'').
 This resampling methodology is based on the conditional randomization test\cite{CetL16}, a recently proposed statistical methodology which, like traditional nonparametric methods, does not rely on correct specification of the outcome (expression) model.
 Unlike traditional nonparametric methods, SCEPTRE employs a resampling scheme that explicitly accounts for heterogeneity in technical factors among cells.
-We analyze both synthetic data and that of Gasperini et al.\cite{Gasperini2019} to illustrate the aforementioned analysis challenges and the performance of the proposed methodology.
+We analyze three datasets -- that of Gasperini et al.\cite{Gasperini2019}, that of Xie et al.\cite{Xie2019}, and a synthetic dataset -- to illustrate the aforementioned analysis challenges and the performance of the proposed methodology.
 SCEPTRE demonstrates excellent calibration and reveals many novel regulatory relationships supported by a variety of existing sources of functional evidence.
@@
 
-The Gasperini et al.\cite{Gasperini2019} data exemplify several of the analysis challenges in single cell CRISPR screens.
-A library of CRISPR guide RNAs (gRNAs) targeting 5,779 candidate enhancers, 50 negative controls, and 381 positive controls was transduced via CROP-seq\cite{Datlinger2017,Hill2018} at high multiplicity of infection into a population of 207,324 K562 cells expressing the Cas9-KRAB repressive complex, each cell receiving an average of 28 perturbations.
-Polyadenylated gRNA barcodes were sequenced alongside the whole transcriptome, and perturbation identities assigned to cells by thresholding the resulting gRNA UMI counts.
-Initially, $p$-values for each gRNA-gene pair were obtained from a DESeq2\cite{Love2014}-inspired negative binomial regression analysis implemented in Monocle2\cite{Qiu2017}.
-By examining the distribution of negative binomial $p$-values pairing each gene with each of 50 non-targeting control (NTC) gRNAs, Gasperini et al.
-found that these $p$-values are inflated (Figure~\ref{fig:analysis-challenges}a).
-To remedy this issue, Gasperini et al.
-calibrated the candidate enhancer $p$-values against the distribution of all gene-NTC $p$-values instead of the uniform distribution.
-The resulting ``empirical" $p$-values were used for determining significance.
+The Gasperini et al.\cite{Gasperini2019} and Xie et al.\cite{Xie2019} data exemplify several of the analysis challenges in single cell CRISPR screens.
+Gasperini et al.\ transduced via CROP-seq\cite{Datlinger2017,Hill2018}  at a high multiplicity of infection a library of CRISPR guide RNAs (gRNAs) into a population of 207,324 K562 cells expressing the Cas9-KRAB repressive complex, each cell receiving an average of 28 perturbations.
+The gRNA library targeted 5,779 candidate enhancers, 50 negative controls, and 381 positive controls.
+Xie et al.\ used Mosaic-seq\cite{Xie2019,Xie2017} to perturb at a high multiplicity of infection 518 putative enhancers in a population of 106,670 Cas9-KRAB-expressing K562 cells.
+Each putative enhancer was perturbed in an average of 1,276 cells.
+Both Gasperini et al.\ and Xie et al.\ sequenced polyadenylated gRNA barcodes alongside the whole transcriptome and assigned perturbation identities to cells by thresholding the resulting gRNA UMI counts.
 
-First, we illustrate the confounding effect discussed in the introduction.
-Despite the use of a targeted amplification protocol for gRNA detection\cite{Hill2018}, not all gRNAs are detected.
-Indeed, the total number of gRNAs detected in a cell increases with the sequencing depth, defined as the total number of mRNA UMIs detected per cell ($\rho = 0.35, p < 10^{-15}$; Figure~\ref{fig:analysis-challenges}b).
-We observed a similar phenomenon in the Mosaic-seq data\cite{Xie2017} ($\rho = 0.23, p < 10^{-15}$; Figure~\ref{fig:mosaicseq}).
-Therefore, sequencing depth and other technical factors induce correlations between the measured gRNA presence and gene expression, even in the absence of a regulatory relationship (Figure~\ref{fig:analysis-challenges}c).
-Parametric regression approaches like those employed by Gasperini et al.
-are the easiest way to adjust for confounders, but their heavy reliance on correct model specification creates a different set of calibration issues.
+Gasperini et al.\ and Xie et al.\ computed $p$-values for gRNA-gene pairs
+to test for associations between candidate enhancers and genes.
+Gasperini et al.\ obtained $p$-values from a DESeq2\cite{Love2014}-inspired negative binomial regression analysis implemented in Monocle2.\cite{Qiu2017} Xie et al., in contrast, obtained $p$-values from Virtual FACS, a nonparametric rank-based method proposed by the authors.\cite{Xie2017} Gasperini et al.\ and Xie et al.\  observed miscalibration in their respective sets of $p$-values.
+Gasperini et al.\ examined the distribution of negative binomial $p$-values obtained from pairing each protein-coding gene with each of 50 non-targeting control gRNAs.
+These ``null'' $p$-values exhibited inflation, deviating substantially from the expected uniform distribution (Figure \ref{fig:analysis-challenges}a, top panel).
+Xie et al.\ examined the distribution of Virtual FACS $p$-values obtained from pairing gRNAs that targeted an enhancer of gene \textit{ARL15}, ARL15-enh, with all protein-coding genes and long noncoding RNAs (lncRNAs) genome-wide.
+As ARL15-enh is expected to modulate the expression of ARL15 and neighboring lncRNAs only, these $p$-values can be used to assess model calibration.
+Xie et al.\
+ found these $p$-values to be miscalibrated, with those corresponding to protein-coding genes exhibiting conservative bias (Figure \ref{fig:analysis-challenges}a, bottom panel) and those corresponding to lncRNAs exhibiting liberal bias (not depicted).
 
-We hypothesized that imperfect dispersion estimation may be responsible for the inflation observed by Gasperini et al.
-Monocle2 fits a raw dispersion for each gene, then fits a mean-dispersion relationship across genes, and finally collapses each raw dispersion estimate onto this fitted line (Figure~\ref{fig:analysis-challenges}d).
-While the raw dispersion estimates are noisy and some shrinkage is helpful\cite{Hafemeister2019}, the collapsing of the dispersion estimates to the fitted curve is likely to underestimate the dispersions for points significantly above the curve.
-On the other hand, points near the curve may not have this issue.
-Indeed, plotting the 50 raw NTC $p$-values for genes near the curve, we found good calibration (e.g., LUC7L2; Figure~\ref{fig:analysis-challenges}e), while genes significantly above the curve showed marked signs of raw $p$-value inflation (e.g., HIST1H1D; Figure~\ref{fig:analysis-challenges}f).
+Confounding likely is a cause of the miscalibration noted by Gasperini et al.\ and Xie et al.
+Despite the use of targeted amplification protocols for gRNA detection\cite{Hill2018}, not all transduced gRNAs are detected in single cell CRISPR screen experiments.
+Indeed, the total number of gRNAs detected in a cell increases with the sequencing depth, defined as the total number of mRNA UMIs detected per cell ($\rho = 0.35, p < 10^{-15}$ in Gasperini et al.\ data; $\rho = 0.25, p < 10^{-15}$ in Xie et al.\ data; Figures \ref{fig:analysis-challenges}b-c).
+Therefore, sequencing depth and other technical factors induce correlations between the measured gRNA presence and gene expression, even in the absence of a regulatory relationship (Figure \ref{fig:analysis-challenges}d).
+This confounding effect, if not explicitly accounted for, results in test miscalibration.
+Confounding is especially problematic for nonparametric methods like Virtual FACS, which implicitly (and incorrectly) assume that there are no technical factors that impact both gRNA detection and gene expression measurement.
 
-This heterogeneity in inflation among gene-gRNA pairs reveals a weakness of the ``one-size-fits-all" NTC-based calibration approach employed by Gasperini et al.: building a null distribution using all gene-NTC pairs leads to overcorrection for some gene-enhancer pairs (false negatives) and undercorrection for others (false positives).
-For example, the empirical correction is unnecessary for genes like LUC7L2.
-However, we pay a price in sensitivity for this empirical correction; it decreases the significance of the $p$-value for a positive control gRNA by three orders of magnitude (Figure~\ref{fig:analysis-challenges}e).
-On the other hand, the empirical correction is insufficient for high-dispersion genes.
-We computed the deviation from uniformity of the empirical NTC $p$-values for each gene using the Kolmogorov-Smirnov (KS) test, represented by the color of each point in Figure~\ref{fig:analysis-challenges}b.
-Circled genes have significantly miscalibrated empirical $p$-values based on a Bonferroni correction at level $\alpha = 0.05$; 21 of these were flagged by Gasperini et al.
-as outliers but 21 were not.
-Among the latter group is HIST1H1D, whose empirical $p$-values are still inflated (Figure~\ref{fig:analysis-challenges}f), leading to two potential false discoveries, one of which is deemed ``high confidence" (Figure~\ref{fig:analysis-challenges}g, Figure~\ref{fig:potential-false-positives}).
+Parametric regression approaches, like those employed by Gasperini et al., are the easiest way to adjust for confounders.
+However, parametric methods rely heavily on correct model specification, a challenge in single cell analysis given the heterogeneity, complexity, and occasional multimodality of the data.
+We hypothesized that inaccurate estimation of the negative binomial dispersion parameter was (in part) responsible for the $p$-value inflation observed by Gasperini et al.
+Monocle2 fits a raw dispersion for each gene, then fits a parametric mean-dispersion relationship across genes, and finally collapses each raw dispersion estimate onto this fitted line (Figure \ref{fig:analysis-challenges}e).
+We computed the deviation from uniformity of the negative control $p$-values for each gene using the Kolmogorov-Smirnov (KS) test, represented by the color of each point in Figure \ref{fig:analysis-challenges}e.
+Circled genes have significantly miscalibrated $p$-values based on a Bonferroni correction at level $\alpha = 0.05$.
+Genes significantly above the curve showed marked signs of $p$-value inflation.
 
-\clearpage
+Gasperini et al.\ and Xie et al.\ adjusted their analyses to remedy the observed calibration problems.
+Gasperini et al.\ calibrated the candidate enhancer $p$-values against the distribution of all gene-negative control $p$-values instead of the uniform distribution.
+The resulting ``empirical $p$-values'' were used for determining significance.
+Along similar lines, Xie et al.\ calibrated the Virtual FACS $p$-values against a simulated set of gene-specific ``null'' $p$-values.
+These analysis adjustments, unfortunately, did not satisfactorily improve calibration.
+The adjustment of Gasperini et al., for example, lead to overcorrection for some gene-enhancer pairs (false negatives) and undercorrection for others (false positives) (Figure 1s).
+Methods that avoid reliance on ad hoc calibration fixes might produce more consistently calibrated $p$-values across entire CRISPR screen datasets.
 
-\addtocounter{page}{-1}
-\thispagestyle{empty} 
 \begin{figure}[h!]
-	\includegraphics[width = \textwidth]{figures/Figure1/Figure1.png}
-	\caption{\textbf{CRISPR screen analysis challenges can lead to false positives and false negatives.} \textbf{a}, QQ-plot of Gasperini et al.
-$p$-values for all gene-gRNA pairs involving either candidate enhancer or negative control gRNAs.
-Inflation in negative controls makes it harder to interpret associations between genes and candidate enhancers, motivating the empirical $p$-value correction.
-\textbf{b-c}, Sequencing depth (total mRNA UMIs per cell) impacts gRNA detection (b) and observed expression levels, and therefore acts as a confounder (c).
+\includegraphics[width = \textwidth]{figures/Figure1/Figure1_cropped.pdf}
+\caption{\textbf{CRISPR screen analysis challenges can lead to false positives and false negatives}.
+\textbf{a}, (Top) QQ-plot of Gasperini et al.\ p-values for all pairs of genes and negative control gRNAs.
+(Bottom) QQ-plot of Xie et al.\ p-values for all pairs of genes and ARL15-enh-targeting gRNAs (\textit{ARL15} gene omitted).
+These sets of $p$-values deviate from the expected uniform distribution, indicating test miscalibration.
+\textbf{b-d}, Sequencing depth (total mRNA UMIs per cell) impacts gRNA detection and observed gene expression levels in both Gasperini et al.\ (\textbf{b}) and Xie et al.\ (\textbf{c}) data.
+Thus, sequencing depth acts as a confounder (\textbf{d}).
 This analysis challenge distinguishes CRISPR screens from traditional differential expression applications.
+\textbf{e}, Monocle2 estimates the dispersion of each gene by projecting each gene's raw dispersion estimate onto the fitted raw dispersion-mean expression curve.
+This estimation procedure leads to miscalibration for high-dispersion genes.
+% \textbf{b-c}, Sequencing depth (total mRNA UMIs per cell) impacts gRNA detection (b) and observed expression levels, and therefore acts as a confounder (c).
+This analysis challenge distinguishes CRISPR screens from traditional differential expression applications.
+}
 }
@@
 \end{figure}
-
 \clearpage
@@
 
-We attempted to alleviate the miscalibration within the negative binomial regression framework (Figure \ref{fig:calibration-comparison-all}).
-First, we replaced the collapsed dispersion estimates by the raw estimates (recall Figure~\ref{fig:analysis-challenges}d).
-This simple modification is based on the intuition that not as much dispersion shrinkage is necessary as in bulk RNA-seq, since in scRNA-seq we have many more samples (one for each cell) to estimate this parameter.
-This dispersion estimate modification yielded greatly improved calibration over all gene-NTC pairs (blue curve in Figure \ref{fig:calibration-comparison-all}a), but separating by NTC revealed substantial remaining gRNA-dependent miscalibration (Figure~\ref{fig:calibration-comparison-all}b).
-This dependency on gRNA further underscores the problem with ``one-size-fits-all" NTC-based calibration approaches, also implemented by Xie et al.\cite{Xie2019}
+We attempted to alleviate the miscalibration within the negative binomial regression framework.
+To this end, we followed the recommendations of Hafmeister and Satija, who recently proposed a strategy for parametric modeling of single cell RNA-seq data that, to the best of our knowledge, is state-of-the-art.\cite{Hafemeister2019}  First, we abandoned the DESeq2-style size factors of Monocle2 and instead corrected for sequencing depth by including sequencing depth as a covariate in the negative binomial regression model.
+Second, we adopted a more flexible dispersion estimation procedure.
+Following Hafmeister and Satija, we (i) computed raw dispersion estimates for each gene, (ii) regressed the raw dispersion estimates onto the mean gene expressions via kernel regression, and (iii) projected the raw dispersion estimates onto the fitted nonparametric regression curve.
 
-We hypothesized that this remaining miscalibration stems from inadequate sequencing depth correction.
-Indeed, we found we could predict the direction and strength of the miscalibration for each gRNA from an adjusted correlation between the sequencing depth and the gRNA presence (Methods).
-This yielded $z$-values as shown in the legend of Figure~\ref{fig:calibration-comparison-all}b; large positive (negative) $z$-values signal spurious positive (negative) correlations between gRNA presence and sequencing depth, leading to conservative (liberal) bias.
-Gasperini et al.
-corrected for sequencing depth using DESeq2-style size factors, whereas recent work on scRNA-seq analysis\cite{Hafemeister2019} advocates for including sequencing depth as a covariate in the negative binomial regression.
-We therefore included as covariates both the total number of UMIs observed and the number of genes with at least one UMI in a given cell, in addition to the technical factors Gasperini et al.
-corrected for (see Methods).
+We reanalyzed the Gasperini et al.\ data using the improved negative binomial regression approach.
+In addition to sequencing depth, we included as covariates in the regression model the total number of expressed genes per cell, as well as the technical factors that Gasperini et al.\ accounted for in their analysis (namely, total number of gRNAs detected per cell, percentage of transcripts mapped to mitochondrial genes, and sequencing batch).
+We also applied the improved negative binomial approach to the Xie et al.\ ARL15-enh data.
+In the latter analysis we included sequencing depth, number of detected gRNA UMIs per cell, and sequencing batch as covariates.
+We found that the improved negative binomial model was better calibrated than the Monocle2 negative binomial model on the Gasperini data (Figure 3c; improved negative binomial denoted ``Hafemeister NB'').
+However, on both the Gasperini et al.\ data and Xie et al.\ data, the improved negative binomial model showed clear signs of p-value inflation (Figure 3c-d).
 
-To assess the impact of these two improvements, we applied the negative binomial test with and without these improvements (four total possibilities) to all 50 NTC gRNAs paired with all genes  (Figure~\ref{fig:calibration-comparison-all}a,c).
-The changes to confounder correction and dispersion estimation both markedly improve calibration, especially when used in conjunction.
-However, the resulting improved negative binomial method still shows clear signs of miscalibration, producing mostly conservative $p$-values.
-There are many possible reasons for the remaining miscalibration, most likely due to remaining misspecification of the negative binomial model.
-While more effort could certainly be invested in further improvements of this parametric model for gene expression, we acknowledge that modeling scRNA-seq expression is a challenging problem that remains open.
-This highlights the appeal of nonparametric approaches; however, as discussed before, traditional nonparametric approaches are not applicable in the presence of technical factors impacting both gRNA detection and gene expression measurement.
+There are many possible explanations for the remaining miscalibration, mostly due to misspecification of the negative binomial model.
+While more effort could be invested into further improvements of a parametric model for gene expression, modeling single cell data remains a challenging problem outside the scope of the current work.
+This difficulty highlights the appeal of nonparametric methods for singe cell analysis.\cite{Zhu2019} However, as discussed in the previous section, traditional nonparametric approaches are not applicable in the presence of technical factors that impact both gRNA detection and gene expression measurement.
 To resolve this impasse, we propose a resampling-based approach designed to account for such confounders.
 
+% The difficulty of applying parametric models to single cell data highlights the appeal of nonparametric approaches for this purpose.
 \clearpage
-
 \paragraph{SCEPTRE: Analysis of single cell perturbation screens via conditional resampling.}
@@
 \thispagestyle{empty}
-\addtocounter{page}{-1}
 \begin{figure}[h!]
@@
 
-\paragraph{SCEPTRE is well calibrated despite sequencing depth confounding and expression model misspecification.}
+\paragraph{SCEPTRE produces correct results when the ground truth is known.}
 
-First, we demonstrated SCEPTRE's calibration in a small proof-of-concept simulation with 1,000 cells, one gene, one NTC gRNA, and one confounder (Figure~\ref{fig:calibration}a).
-We considered the one-sided $z$ test statistic based on a negative binomial regression with dispersion 1.
-We then generated the data from this model as well as from three others: one with dispersion 0.2, one with dispersion 5, and one with dispersion 1 but with 25\% zero-inflation.
-We compared three ways of building a null distribution for the test statistic based on the possibly misspecified negative binomial model: the standard parametric approach, the permutation approach (based on permuting gRNA assignments), and conditional resampling.
-The parametric approach works as expected when the negative binomial model is correctly specified, but breaks down in all three cases of model misspecification.
-The permutation approach is systematically conservative because of inadequate confounder correction.
-Finally, SCEPTRE is well-calibrated regardless of expression model misspecification.
+We assessed the calibration, sensitivity, and specificity of SCEPTRE in several settings in which the ground truth was known.
+First, we demonstrated the calibration of SCEPTRE in a small, proof-of-concept simulation study (Figure \ref{fig:ground_truth}a).
+We considered a class of negative binomial regression models with dispersion $\alpha$ and two technical covariates (sequencing depth and batch).
+We simulated expression data for a single gene in 1000 cells using four models selected from this class: the first with $\alpha = 1$, the second with $\alpha = 0.25$, the third with $\alpha = 5$, and the last with $\alpha = 1,$ but with 25$\%$ zero-inflation.
+We also simulated negative control gRNA data using a logistic regression model with the same covariates as in the gene expression model.
+We assessed the calibration of three methods across the four simulated datasets: SCEPTRE, improved negative binomial regression, and scMAGeCK-LR,\cite{Yang2020} a recently-proposed, permutation-based nonparametric method.
+To assess the impact of model misspecification on SCEPTRE and the improved negative binomial method (on which SCEPTRE relies), we fixed the dispersion of the negative binomial method to 1 across all four simulated datasets.
+We found that the negative binomial method works as expected when the model is correctly specified but breaks down in all three cases of model misspecification.
+scMAGeCK-LR exhibits poor calibration across all simulated datasets, likely because, as a traditional nonparametric method, it fails to adequately account for confounders.
+Finally, SCEPTRE is well-calibrated regardless of expression model misspecification and confounder presence.
 
-Next, we applied SCEPTRE to test the association between all NTC gRNAs and all genes in the Gasperini et al.
-data (Figure~\ref{fig:calibration}b-c).
-For comparison, we also applied the original negative binomial method, the improved negative binomial method, and a permutation-based calibration of the latter.
-The simplest calibration assessment is to compare the resulting 538,560 $p$-values to the uniform distribution (Figure~\ref{fig:calibration}b).
-SCEPTRE shows excellent calibration, substantially improving on the parametric approach based on the same test statistic.
-Breaking the pairs down by NTC (Figure~\ref{fig:calibration}c), SCEPTRE again shows nearly perfect calibration, as evidenced by its colored confidence region coinciding with the gray shaded region representing the 95\% confidence band for the uniform distribution.
+Next, to assess the calibration of SCEPTRE on real data, we applied SCEPTRE to test the association between all negative control gRNAs and all genes in the Gasperini et al.\ data (Figure \ref{fig:ground_truth}b) and all ARL15-enh-targeting gRNAs and all genes in the Xie et al.\ data (Figure \ref{fig:ground_truth}c).
+We compared SCEPTRE to the improved negative binomial method, as well as to the original analysis methods (i.e., Monocle regression on the Gasperini et al.\ data and Virtual FACS on the Xie et al.\ data).
+We did not apply scMAGeCK-LR to these data given its poor calibration on the simulated data.
+SCEPTRE shows excellent calibration on the Gasperini et al.\ negative control data; Monocle regression and improved negative binomial regression, in contrast, demonstrate signs of severe $p$-value inflation.
+Similarly, on the Xie et al.\ ARL15-enh data, SCEPTRE exhibits nearly perfect calibration, whereas Virtual FACS and the improved negative binomial method show conservative and liberal bias, respectively.
 
-To assess the sensitivity of SCEPTRE, we applied it to the 381 positive control gRNA / gene pairs assayed in Gasperini et al (Figure~\ref{fig:calibration}d), where each gRNA targeted the transcription start site of the corresponding gene.
-Comparing to the original empirical $p$-values, SCEPTRE $p$-values for these positive controls are generally more significant, suggesting greater sensitivity.
-We note that the empirical correction employed by Gasperini et al.
-limits the accuracy of $p$-values to about $10^{-6}$, explaining at least part of the difference.
+Last, we assessed the sensitivity of SCEPTRE using the Gasperini et al.\ positive control data and confirmatory bulk RNA-seq data produced by Xie et al.
+We applied SCEPTRE to the 381 positive control gRNA-gene pairs assayed in Gasperini et al.\ (Figure~\ref{fig:ground_truth}d), where each gRNA targeted the transcription start site of the corresponding gene.
+The SCEPTRE $p$-values for these positive controls are highly significant, and in particular, more significant than the original empirical $p$-values, indicating greater sensitivity.
+We note that the empirical correction employed by Gasperini et al.\ limits the accuracy of $p$-values to about $10^{-6}$, explaining at least part of the difference.
+Finally, we compared the SCEPTRE results for ARL15-enh to the results of an arrayed CRISPR screen of ARL15-enh in which populations of perturbed cells were compared to populations of unperturbed cells via bulk RNA-seq.\cite{Xie2019} Both SCEPTRE and the bulk RNA-seq differential expression analysis rejected gene \textit{ARL15} (and only gene \textit{ARL15}) at an FDR of 0.1 after a Benjamini-Hochberg correction, increasing our confidence in sensitivity (and specificity) of SCEPTRE (Figure \ref{fig:ground_truth}e).
 
+\newpage
+\begin{figure}[h!]
+	\includegraphics[width = \textwidth]{figures/Figure3/Figure3_cropped.pdf}
+	\caption{\textbf{SCEPTRE demonstrates excellent calibration, specificity, and sensitivity when the ground truth is known.}}
+	\label{fig:ground_truth}
+ \end{figure}
 \clearpage
-\thispagestyle{empty} 
-\addtocounter{page}{-1}
-\begin{figure}[h!]
-	\includegraphics[width = \textwidth]{figures/Figure3/Figure3.png}
-	\caption{\textbf{SCEPTRE improves on the sensitivity and specificity of the parametric approach by correcting for expression model misspecification.} \textbf{a,} Numerical simulation comparing three ways of calibrating a negative binomial test statistic (parametric, permutation, conditional resampling), when the test statistic is based on the correct model or contains expression model misspecifications.
+
+% \clearpage
+% \thispagestyle{empty} 
+% \addtocounter{page}{-1}
+%\begin{figure}[h!]
+%	\includegraphics[width = \textwidth]{figures/Figure3/Figure3.png}
+%	\caption{\textbf{SCEPTRE improves on the sensitivity and specificity of the parametric approach by correcting for expression model misspecification.} \textbf{a,} Numerical simulation comparing three ways of calibrating a negative binomial test statistic (parametric, permutation, conditional resampling), when the test statistic is based on the correct model or contains expression model misspecifications.
 Only the conditional resampling approach maintains calibration despite model misspecification.
@@
 Note: $p$-values truncated for visualization, and empirical $p$-values in panel d at least 1/538,560 $\approx$ 2$\times 10^{-6}$ by construction.}
-	\label{fig:calibration}
-\end{figure}
+%	\label{fig:calibration}
+% \end{figure}
 \clearpage
@@
 
-We applied SCEPTRE to the 84,595 gene-enhancer pairs considered in Gasperini et al., encompassing 10,560 genes and 5,779 candidate enhancers (Figure~\ref{fig:application}a).
-We applied the Benjamini-Hochberg correction at level 0.1 (the same level used in the original analysis) to the $p$-values obtained for all of these candidate pairs, obtaining a total of 585 gene-enhancer pairs.
+We applied SCEPTRE to the 84,595 gene-enhancer pairs considered in Gasperini et al., encompassing 10,560 genes and 5,779 candidate enhancers.
+We applied the Benjamini-Hochberg correction at level 0.1 (the same level used in the original analysis) to the $p$-values obtained for all of these candidate pairs, obtaining a total of 563 gene-enhancer pairs.
 By comparison, Gasperini et al.
 found 470 high-confidence pairs.
-Comparing the SCEPTRE $p$-values against the original empirical $p$-values (Figure~\ref{fig:application}b), we see that the two often diverge substantially.
-Our analysis found 217 gene-enhancer pairs that the original analysis did not, while 102 were found only by the original analysis.
-Many of the discoveries found only in the original analysis show signs of the $p$-value inflation observed in Figure~\ref{fig:analysis-challenges} (Figure~\ref{fig:potential-false-positives}).
+Comparing the SCEPTRE $p$-values against the original empirical $p$-values (Figure~\ref{fig:application}a), we see that the two often diverge substantially.
+Our analysis found 200 gene-enhancer pairs that the original analysis did not, while 107 were found only by the original analysis.
+Many of the discoveries found only in the original analysis show signs of the $p$-value inflation observed in Figure~\ref{fig:analysis-challenges}.
 
-Among the 217 new gene-enhancer pairs discovered, several are supported by evidence from orthogonal functional assays.
-In particular, we highlight five of these pairs (Figure~\ref{fig:application}c) involving genes not paired to any enhancers in the original analysis, which are supported by GTEx\cite{GTExPilot} eQTL $p$-values in whole blood or enhancer RNA correlation $p$-values across tissues from the FANTOM project\cite{Andersson2014}.
+Among the 200 new gene-enhancer pairs discovered, several are supported by evidence from orthogonal functional assays.
+In particular, we highlight five of these pairs (Figure~\ref{fig:application}b) involving genes not paired to any enhancers in the original analysis, which are supported by GTEx\cite{GTExPilot} eQTL $p$-values in whole blood or enhancer RNA correlation $p$-values across tissues from the FANTOM project\cite{Andersson2014}.
 These pairs are listed in the GeneHancer database\cite{Fishilevich2017}, which aggregates eQTL, eRNA, and other sources of evidence of gene-enhancer interactions.
 The SCEPTRE $p$-values for these promising pairs are generally 1-2 orders of magnitude more significant than the original empirical $p$-values.
-Also among the 217 new gene-enhancer pairs are 9 pairs involving 7 genes that were deemed outliers in the original analysis (blue triangles in Figure~\ref{fig:application}b), underscoring SCEPTRE's ability to handle genes with arbitrary distributions that may not fit into traditional parametric models.
+Also among the 200 new gene-enhancer pairs are 6 pairs involving 5 genes that were deemed outliers in the original analysis (blue triangles in Figure~\ref{fig:application}a), underscoring SCEPTRE's ability to handle genes with arbitrary distributions that may not fit into traditional parametric models.
 
 We also found that the total set of gene-enhancer pairs discovered was better enriched for regulatory biological signals, including HI-C and ChIP-seq.
-Gene-enhancer pairs discovered by SCEPTRE tended to be closer to each other on average that those discovered in the original analysis (Figure~\ref{fig:application}d).
-Furthermore, 76\% of SCEPTRE's 585 gene-enhancer pairs fell in the same topologically associating domain (TAD), compared to 71\% of the original 470 pairs.
-We also repeated Gasperini et al.'s contact frequency enrichment analysis for those pairs falling in the same TAD (Figure~\ref{fig:application}e).
-We found similar levels of contact frequency enrichment, despite the fact that we had 108 more gene-enhancer pairs in the same TADs.
+Gene-enhancer pairs discovered by SCEPTRE tended to be physically closer to each other on average that those discovered in the original analysis (Figure~\ref{fig:application}c).
+Furthermore, 74.4\% of SCEPTRE's 563 gene-enhancer pairs fell in the same topologically associating domain (TAD), compared to 71.1\% of the original 470 pairs.
+We also repeated Gasperini et al.'s contact frequency enrichment analysis for those pairs falling in the same TAD (Figure~\ref{fig:application}d).
+We found similar levels of contact frequency enrichment, despite the fact that SCEPTRE discovered 85 more gene-enhancer pairs in the same TADs.
 Finally, we repeated the ChIP-seq enrichment analysis of Gasperini et al.
@@
 We quantified enrichment of ChIP-seq signal in paired enhancers as the odds ratio that a candidate enhancer is paired to a gene, comparing those falling in the top quintile of candidate enhancers by ChIP-seq signal and those not overlapping a ChIP-seq peak at all.
-We find improved enrichment for each of the eight ChIP-seq targets (Figures~\ref{fig:application}f and \ref{fig:chipseq-enrichment-details}).
+We find improved enrichment for each of the eight ChIP-seq targets (Figures~\ref{fig:application}e and \ref{fig:chipseq-enrichment-details}).
 
 \clearpage
-\newgeometry{top=0.6in}
-\thispagestyle{empty} 
-\addtocounter{page}{-1}
-\vspace{-2in}
-
-\begin{figure}
-	\includegraphics[width = \textwidth]{figures/Figure4/Figure4.png}
-	\caption{\textbf{Application to Gasperini et al.
+% \newgeometry{top=0.6in}
+% \thispagestyle{empty} 
+% \addtocounter{page}{-1}
+% \vspace{-2in}
+% \begin{figure}
+%	\includegraphics[width = \textwidth]{figures/Figure4/Figure4.png}
+%	\caption{\textbf{Application to Gasperini et al.
 data yields biologically plausible gene-enhancer links.}  \textbf{a,} Well-calibrated NTC $p$-values give confidence in SCEPTRE $p$-values for candidate enhancers (compare to Figure~\ref{fig:analysis-challenges}a).
@@
 SCEPTRE exhibits greater enrichment across all ChIP-seq targets.}
+%	\label{fig:application}
+% \end{figure}
+
+\begin{figure}
+\includegraphics[width = \textwidth]{figures/Figure4/Figure4_cropped.pdf}
+	\caption{\textbf{Application to Gasperini et al. data yields biologically plausible gene-enhancer links.} }
 	\label{fig:application}
@@
 \clearpage
-
-\restoregeometry
+% \restoregeometry
 
@@
 We use SCEPTRE to reanalyze the Gasperini et al.
+data and a subset of the Xie et al.
 data, yielding many new biologically plausible gene-enhancer relationships supported by evidence from eQTL, enhancer RNA co-expression, ChIP-seq, and HI-C data.
@@
 Furthermore, we conjecture that misspecifications in the gRNA measurement model can be compensated by improved gene expression models, a statistical phenomenon referred to as ``double robustness"\cite{Robins2001,VanderLaan2003}.
-%While SCEPTRE greatly reduces the burden of modeling single cell expression data, there is therefore still good reason to seek improved expression models. 
 Better expression models can also improve the sensitivity of the conditional randomization test\cite{Katsevich2020a}.
@@
 \begin{addendum}
- \item We are indebted to Molly Gasperini and Jacob Tome for clarifying several aspects of their data analysis\cite{Gasperini2019}, and to the Shendure lab and Timothy Barry for providing extensive feedback on an earlier draft of this paper.
-We also thank Tom Norman, Atray Dixit, and Wesley Tansey for useful discussions on single cell CRISPR screens.
+ \item We are indebted to Molly Gasperini, Jacob Tome, and Andrew Hill for clarifying several aspects of their data analysis\cite{Gasperini2019} and to the Shendure lab for providing extensive feedback on an earlier draft of this paper.
+We thank Shiqi Xie for providing guidance on using the Xie et al.\ 2019 data, Wie Li for a helpful discussion on scMAGeCK, and John Morris for providing feedback on the SCEPTRE code.
+Finally, we thank Tom Norman, Atray Dixit, and Wesley Tansey for useful discussions on single cell CRISPR screens.
 This work was supported, in part, by National Institute of Mental Health (NIMH) grants R01MH123184 and R37MH057881 as well as SFARI Grant 575547.
@@
 
-\paragraph{Reproducing the negative binomial approach of Gasperini et al.}
-
+\paragraph{Conditional randomization test.}
 Consider a particular gene/gRNA pair.
 For each cell $i = 1, \dots, n$, let $X_i \in \{0,1\}$ indicate whether the gRNA was present in the cell, let $Y_i \in \{0,1,2,\dots\}$ be the gene expression in the cell, defined as the number of unique molecular identifiers (UMIs) from this gene, and let $Z_i \in \mathbb R^d$ be a list of cell-level technical factors.
-
-Gasperini et al. used a negative binomial regression of $Y_i$ on $X_i$ and $Z_i$: 
-\begin{equation}
-Y_i \overset{\text{ind}} \sim \text{NegBin}(\mu_i, \alpha); \quad \log(\mu_i) = \beta_0 + X_i \beta + Z_i^T \gamma,
-\label{negative-binomial}
-\end{equation}
-with $\alpha$ being the dispersion and $Z_i$ consisting of three cell-level technical factors: the total number of gRNAs in the cell, the percentage of observed transcripts that are mitochondrial, and the sequencing batch.
-To correct for sequencing depth, Gasperini et al.
-replaced $Y_i$ in the above regression by rounding $Y_i/s_i$ to the nearest integer, where $s_i$ are DESeq2-style size factors.
-Raw dispersions $\alpha_{\text{raw}}$ were estimated for each gene based on its sample mean and variance.
-Gasperini et al fit a mean-dispersion relationship to these raw dispersions, and finally obtained a shrunk estimate of $\alpha$ by projecting $\alpha_{\text{raw}}$ onto this curve (Figure~\ref{fig:analysis-challenges}b).
-The significance of $X_i$ in the above regression was determined with a likelihood ratio test for $H_0: \beta = 0$.
-Since enhancer perturbation decreases gene expression, gene-enhancer pairs with $\widehat \beta > 0$ were removed from the analysis.
-
-%Our analysis was based on the conditional permutation test outlined above, using a test statistic $T(X,Y,Z)$ similar to that used by Gasperini et al., but with three modifications: improved dispersion estimation, improved confounder correction, and $z$-statistic instead of likelihood ratio statistic.
-We discuss each of these next.
-%
-%Gasperini et al.
-computed raw dispersion estimates via the method of moments, then fit a mean-dispersion relationship, and then projected the raw dispersions onto this curve (recall Figure~\ref{fig:analysis-challenges}b).
-We simply removed this projection step, using the raw dispersions instead.
-We could have used a more sophisticated dispersion estimation scheme (which, if successful, could have boosted our power), but the key feature of our procedure is that its validity does not depend on the expression model---or the dispersion in particular---being correctly specified.
-%
-%In addition to the three cell-level covariates $Z_i$ corrected for by Gasperini et al., we also included two others: the total number of UMIs observed and the number of genes with at least one UMI in a given cell.
-We use these additional covariates to correct for sequencing depth, instead of the size factors used in the original analysis.
-This strategy is in line with that advocated by a recent work on scRNA-seq analysis\cite{Hafemeister2019}.
-%
-%The likelihood ratio test employed by Gasperini et al.
-is a two-sided test; it does not distinguish between positive and negative associations of the gRNA with gene expression.
-Instead, we used the standard GLM $z$-statistic for the coefficient $\beta$, and measured one-sided departures from the null in the negative direction.
-
-\paragraph{Exploration of sequencing depth confounding.}
-%To focus on the issue of sequencing depth confounding, we used raw dispersion estimates (compare the red and blue curves in Figure~\ref{fig:calibration-comparison-all}) as well as one-sided $z$-tests.
-We then repeated the negative binomial regression analysis underlying Monocle2 with these updated dispersion estimates.
-These two modifications yielded new negative binomial regression $p$-values for each gene/gRNA pair (Figure~\ref{fig:analysis-challenges}g).
-
-After improving the dispersion estimates, we sought to understand the direction and strength of the remaining miscalibration for each gRNA, reasoning it may relate to sequencing depth confounding.
-To this end, we tested the residual association between each gRNA and the sequencing depth, after accounting for the total number of gRNAs per cell.
-In particular, for each gRNA, we ran a logistic regression of the gRNA indicator against the total number of gRNAs and the total number of UMIs detected per cell.
-We then extracted the $z$-score of the coefficient of the total number of UMIs.
-We found these agreed well with the direction and strength of the confounding effect, as shown in Figure~\ref{fig:calibration-comparison-all}b.
-
-\paragraph{Conditional randomization test and accelerations.}
-
 Letting $(X, Y, Z) = \{(X_i, Y_i, Z_i)\}_{i = 1}^{n}$, consider any test statistic $T(X,Y,Z)$ measuring the effect of the gRNA on the expression of the gene.
@@
 $Y \independent X \mid Z$, we obtain a valid $p$-value~\eqref{conditional-randomization}, \textit{regardless of the expression distribution $Y|X,Z$ and regardless of the test statistic $T$}.
-We choose a test statistic $T$ based on the improved negative binomial regression discussed in the main text, with two computational accelerations.
+We choose as a test statistic $T$ the $z$-score of $X_i$ obtained from a negative binomial regression of $Y_i$ on $X_i$ and $Z_i$:
 
+\begin{equation}
+Y_i \overset{\text{ind}} \sim \text{NegBin}(\mu_i, \alpha); \quad \log(\mu_i) = \beta_0 + X_i \beta + Z_i^T \gamma
+\label{negative-binomial},
+\end{equation}
+where $\alpha$ is the dispersion.
+Following Hafemeister and Satjia\cite{Hafemeister2019}, we estimate $\alpha$ by pooling dispersion information across genes, and we include sequencing depth as an entry in the vector of technical factors  $Z_i$ (see section \textit{Improvements to the negative binomial approach}).
+
+\paragraph{Accelerations to the conditional randomization test.}
+We implemented computational accelerations to the conditional randomization test.
 First, we employed the recently proposed\cite{Liu2020} \textit{distillation} technique to accelerate the recomputation of the negative binomial regression for each resample.
@@
 In our data analysis, we used $B = 500$ resamples.
-Following Gasperini et al, we based our analysis on the top two gRNAs targeting each enhancer.
+Following Gasperini et al.\, we based our analysis on the top two gRNAs targeting each enhancer.
 Some enhancers were also targeted with two additional gRNAs, but we excluded these from the analysis.
@@
 
-We simulated one gene $Y_i$, one gRNA $X_i$, and one confounder $Z_i$ in $n = 1000$ cells.
-We drew $Z_i \overset{\text{i.i.d.}}\sim N(0,1)$.
-Then, we drew $X_i$ based on the logistic model~\eqref{logistic-regression}, with $\tau = 4$ and $\tau_0 = \log \frac{0.05}{0.95}$, the latter to make the probability of gRNA occurrence about $0.05$ on average across cells.
+We simulated one gene $Y_i,$ five gRNAs $X_{i1}, X_{i2}, \dots, X_{i5},$ and two confounders $Z_{i1}, Z_{i2}$ in $n = 1000$ cells.
+We generated the confounders $Z_{i1}$ and $Z_{i2}$ by sampling with replacement the sequence batch IDs and log-transformed sequencing depths, respectively, of the cells in the Gasperini dataset.
+The batch ID confounder $Z_{i1}$ was a binary variable, as the Gasperni data included two batches.
+Next, we drew the gRNA indicators $X_{i1}, X_{i2}, \dots, X_{i5}$ i.i.d.\ from the logistic regression model (7), with $\tau_0 = -7, \tau_1 = -2,$ and $\tau_2 = 0.5$.
+We selected these parameters to make the probability of gRNA occurrence about 0.04 across cells.
 Finally, we drew the gene expression $Y_i$ from the following zero-inflated negative binomial model:
-\begin{equation*}
-Y_i \overset{\text{ind}}\sim \lambda \delta_0 + (1-\lambda)\text{NegBin}(\mu_i, \alpha), \quad \log(\mu_i) = \beta_0 + 4Z_i.
-\end{equation*}
-Note that for any $\beta_0,\lambda,\alpha$, the gRNA does not impact the gene expression in this setup.
-We chose $\beta_0 = \log(5)$ to make the average gene expression about 5.
-The four settings shown in Figure~\ref{fig:calibration}a correspond to
-\begin{equation*}
-(\lambda_1, \alpha_1) = (0,1); \quad (\lambda_2, \alpha_2) = (0,5); \quad (\lambda_3, \alpha_3) = (0,0.2); \quad (\lambda_4, \alpha_4) = (0.25,1).
-\end{equation*}
-For the first, the negative binomial model is correctly specified.
+$$ Y_i \sim \lambda \delta_0 + (1 - \lambda) \textrm{NegBin}(\mu_i, \alpha), \textrm{ } \log(\mu_i) = \beta_0 + Z_i^T \beta.$$ Note that gRNA presence or absence does not impact gene expression in this model.
+We set $\beta_0 = -2.5, \beta_1 = -2, \beta_2 = 0.5$ to make the average gene expression about $4$ across cells.
+We generated the four datasets shown in Figure \ref{fig:ground_truth}a by setting the dispersion parameter $\alpha$ and the zero inflation rate parameter $\lambda$ equal to the following values:
+$$ (\lambda_1, \alpha_1) = (0,1) ; \textrm{    } (\lambda_2, \alpha_2) = (0, 5); \textrm{     } (\lambda_3, \alpha_3) = (0, 0.2); \textrm{     } (\lambda_4, \alpha_4) = (0.25, 1).$$ For the first, the negative binomial model is correctly specified.
 For the second and third, the dispersion estimate of 1 is too small and too large, respectively.
-The last setting exhibits zero inflation.
+The last setting exhibits zero inflation
 
-We applied three methods to these four problem settings, each with 500 repetitions.
-The negative binomial method was based on the $z$ statistic from the standard negative binomial regression~\eqref{negative-binomial} with $\alpha = 1$.
-The permutation method was implemented the same as SPECTRE, except skipping step 3 and defining $\widetilde X^b$ to be a random permutation of $X$.
-Both the permutation method and SPECTRE used $B = 250$ resamples.
+We applied SCEPTRE, negative binomial regression, and scMAGeCK-LR\cite{Yang2020} to the four problem settings, each with $n_{sim} = 500$ repetitions.
+The negative binomial method was based on the $z$ statistic from the Hafemeister-inspired negative binomial model (\ref{negative-binomial}) with $\alpha = 1$.
+scMAGeCK-LR differs from SCEPTRE and the negative binomial method in that scMAGeCK-LR computes $p$-values for all enhancers simultaneously.
+Thus, to facilitate comparisons across methods, we plotted $p$-values corresponding to enhancer $X_{i1}$ only.
+We used $B = 500$ resamples for SCEPTRE and $B = 1000$ permutations for scMAGeCK-LR, the default choices for these methods.
 
@@
 
-These analyses (see Figures~\ref{fig:application}e-f and \ref{fig:chipseq-enrichment-details}) were carried out almost exactly following Gasperini et al.
+ChIP-seq and HI-C enrichment analyses analyses (see Figures~\ref{fig:application}e-f and \ref{fig:chipseq-enrichment-details}) were carried out almost exactly following Gasperini et al.
 The only change we made is in our quantification of the ChIP-seq enrichment (Figure~\ref{fig:application}f).
 We use the odds ratio of a candidate enhancer being paired to a gene, comparing the top and bottom ChIP-seq quintiles.
-
-%\paragraph{Methods used in calibration comparison.}
-%
-%In Figure~\ref{fig:calibration-comparison-all}, we compared the calibration of the proposed approach to three alternatives.
-The first, ``Original NB", is the same as Gasperini et al.'s negative binomial regression, except with a one-sided $z$-statistic instead of a likelihood ratio statistic.
-The second ``Improved NB", is the same except with improved confounder correction and dispersion estimation.
-The third, ``Marginal permutation", is the same as conditional permutation, except the gRNAs are permuted uniformly among cells rather than conditionally on the their covariates.
 
@@
 %\newgeometry{top=1in}
-\thispagestyle{empty} 
-\addtocounter{page}{-1}
+% \thispagestyle{empty} 
+% \addtocounter{page}{-1}
 
-\counterwithin{figure}{section}
+% \counterwithin{figure}{section}
 \section{Supplementary Figures}
@@
 
-\begin{figure}[h!]
-	\centering
-	\includegraphics[width = \textwidth]{figures/FigureS1/FigureS1.png}
-	\caption{Gasperini et al.'s empirical correction is insufficient to correct for miscalibration.
+%\begin{figure}[h!]
+%	\centering
+%	\includegraphics[width = \textwidth]{figures/FigureS1/FigureS1.png}
+%	\caption{Gasperini et al.'s empirical correction is insufficient to correct for miscalibration.
 \textbf{a,} Dispersion estimation procedure employed leads to miscalibration for high-dispersion genes, which the empirical correction does not adequately correct for, as measured by KS test applied to empirical $p$-values per gene (point colors).
@@
 Dashed horizontal line represents the multiple testing threshold.}
-	\label{fig:mosaicseq}
-\end{figure}
+%	\label{fig:mosaicseq}
+% \end{figure}
 
@@
 
-\thispagestyle{empty} 
-\begin{figure}[h!]
-	\includegraphics[width = \textwidth]{figures/FigureS2/FigureS2.png}
-	\caption{\textbf{Comparing negative binomial and conditional resampling $p$-values based on the same test statistic.} \textbf{a,} The standard parametric negative binomial $p$-value versus that obtained from the same test statistic by conditional resampling, for each gene / candidate enhancer pair (both truncated at $10^{-10}$ for visualization).
+% \thispagestyle{empty} 
+% \begin{figure}[h!]
+%	\includegraphics[width = \textwidth]{figures/FigureS2/FigureS2.png}
+%	\caption{\textbf{Comparing negative binomial and conditional resampling $p$-values based on the same test statistic.} \textbf{a,} The standard parametric negative binomial $p$-value versus that obtained from the same test statistic by conditional resampling, for each gene / candidate enhancer pair (both truncated at $10^{-10}$ for visualization).
 The two can diverge fairly substantially.
 \textbf{b-d,} Parametric and conditional resampling null distributions for the negative binomial $z$-value in three cases: the parametric $p$-value is more significant (b), the conditional resampling $p$-value is more significant (c), the two $p$-values are about the same (d).}
-	\label{fig:NB-vs-CP}
-\end{figure}
+%	\label{fig:NB-vs-CP}
+% \end{figure}
 
@@
 
-\clearpage
-\thispagestyle{empty} 
-\begin{figure}[h!]
-	\includegraphics[width = \textwidth]{figures/FigureS3/FigureS3.pdf}
-	\caption{\textbf{Details on ChIP-seq enrichment analysis.} Fraction of candidate enhancers paired with a gene, broken down by quintile of ChIP-seq signal (0 means the candidate enhancer did not overlap a ChIP-seq peak), based on which the odds ratios in Figure~\ref{fig:application}f were computed.
+% \clearpage
+% \thispagestyle{empty} 
+% \begin{figure}[h!]
+%	\includegraphics[width = \textwidth]{figures/FigureS3/FigureS3.pdf}
+%	\caption{\textbf{Details on ChIP-seq enrichment analysis.} Fraction of candidate enhancers paired with a gene, broken down by quintile of ChIP-seq signal (0 means the candidate enhancer did not overlap a ChIP-seq peak), based on which the odds ratios in Figure~\ref{fig:application}f were computed.
 Both methods generally pair candidate enhancers in higher ChIP-seq quintiles more frequently, but this enrichment is more pronounced in SCEPTRE across all eight ChIP-seq targets.}
-	\label{fig:chipseq-enrichment-details}
-\end{figure}
+%	\label{fig:chipseq-enrichment-details}
+% \end{figure}
 

### 2020-11-28 sceptre_paper/plotting/Figure1.R

@@ -65,6 +65,8 @@ vjust <- 3
 # Subfigure b
 gasp_confounding_df <- covariates_gasp %>% summarize(total_umis = ifelse(is.na(lg_total_umis), 0, exp(lg_total_umis)), guide_count = ifelse(is.na(lg_guide_count), 0, exp(lg_guide_count)))
 rho <- cor(log(gasp_confounding_df$total_umis), gasp_confounding_df$guide_count) %>% signif(2)
+cor.test(log(gasp_confounding_df$total_umis), gasp_confounding_df$guide_count)
+
 text_to_add <- paste0("~ rho == ", rho)
 
 p_b <- gasp_confounding_df %>% filter(guide_count > 0, row_number() %% 10 == 0) %>%
@@ -89,6 +91,7 @@ guide_count[is.na(guide_count)] <- 0
 umi_count <- exp(covariates_xie$log_n_umis)
 xie_confounding_df <- tibble(total_umis = umi_count, guide_count = guide_count) %>% filter(guide_count <= 50)
 rho <- cor(xie_confounding_df)[1,2] %>% signif(2)
+cor.test(xie_confounding_df$total_umis, xie_confounding_df$guide_count)
 text_to_add <- paste0("~ rho == ", rho)
 
 p_c <- xie_confounding_df %>% filter(guide_count > 0, row_number() %% 10 == 0) %>%

### 2020-11-28 sceptre_paper/plotting/Figure4.R

@@ -58,6 +58,8 @@ rejected_by_annotated <- n_rejected_annotated[matches]
 my_labs <- n_rejected_annotated[match(c("Neither method", "Both methods", "SCEPTRE only", "Gasperini only"), names(n_rejected))]
 df_a$rejected_by_annotated <- factor(x = rejected_by_annotated, levels = my_labs, labels = my_labs)
 df_a$outlier_gene <- as.logical(df_a$outlier_gene)
+filter(df_a, outlier_gene, new_rejected) %>% pull(gene_id) %>% unique() %>% length()
+filter(df_a, outlier_gene, new_rejected) %>% nrow()
 
 p_a <- ggplot(data = arrange(df_a, rejected_by_annotated), aes(x = old_pvalue, y = new_pvalue,
              color = rejected_by_annotated, label = pair_number, shape = outlier_gene)) + geom_point(alpha = 0.9) +
@@ -183,7 +185,7 @@ p_d <- TF_enrichments %>% arrange(desc(method)) %>%
 p_e <- ggplot() + theme_bw() + ggtitle("eQTL and eRNA validation") + theme(plot.title = element_text(hjust = 0.5), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank())
 
 # combine plots: option 1
-left_col <- plot_grid(p_a, p_d, align = "v", labels = c("a", "d"), ncol = 1, rel_heights = c(2,1))
-right_col <- plot_grid(p_b, p_c, p_e, align = "vh", labels = c("b", "c", "e"), ncol = 1)
+left_col <- plot_grid(p_a, p_d, align = "v", labels = c("a", "e"), ncol = 1, rel_heights = c(2,1))
+right_col <- plot_grid(p_e, p_b, p_c, align = "vh", labels = c("b", "c", "d"), ncol = 1)
 final_plot <- plot_grid(left_col, right_col, align = "h", ncol = 2, rel_widths = c(1.1,1))
 ggsave(filename = paste0(fig4_dir, "/subfigures_a_thru_e.pdf"), plot = final_plot, device = "pdf", scale = 1, width = 8, height = 8)

### 2020-11-29 sceptre_paper/manuscript/manuscript_v3.tex
update manuscript

@@
-%% Template for a preprint Letter or Article for submission
-%% to the journal Nature.
-%% Written by Peter Czoschke, 26 February 2004
-%%
-
 \documentclass{nature}
-
-%% make sure you have the nature.cls and naturemag.bst files where
-%% LaTeX can find them
-
 \usepackage{graphicx}	
@@
 \newcommand\independent{\perp \!\!\! \perp}
+\usepackage{capt-of}
 % \usepackage{chngcntr}
@@
 
-%\usepackage[minnames = 3, maxnames=3, url = false, doi = false, isbn = false, sorting = none, backend=bibtex]{biblatex}
-%\addbibresource{/home/ekatsevi/Dropbox/Research/Literature/library.bib}
-
 \title{Conditional resampling improves calibration \\ in single cell CRISPR screen analysis}
-
-%% Notice placement of commas and superscripts and use of &
-%% in the author list
 
@@
 This methodology, designed to avoid calibration issues due to technical confounders and expression model misspecification, infers associations between perturbations and expression by resampling the former according to a working model for perturbation detection probability in each cell.
-Applying SCEPTRE to the CRISPR screen data yields excellent calibration on negative control perturbations and 200 regulatory relationships not found in the original study, many of which are supported by existing functional data.
+SCETPRE demonstrates excellent calibration, sensitivity, and specificity on the CRISPR screen data and yields 200 new regulatory relationships, many of which are supported by existing functional data.
 \end{abstract}
@@
 These pooled screens deliver a library of CRISPR perturbations, targeting either genes or non-coding regulatory elements, to a population of cells and then employ single cell RNA sequencing (scRNA-seq) to identify the perturbations each cell received and measure its transcriptional response.
-Single cell CRISPR screens have been scaled up through multiplexing to systematically map gene-enhancer relationships\cite{Xie2017,Gasperini2019,Xie2019}, applied using combinatorial perturbation libraries to study genetic interactions\cite{Norman2019}, and extended to single cell protein expression readout\cite{Mimitou2019,Frangieh2020}.
-Given their scale, resolution, and versatility, these assays are likely to be a major driver of discovery in functional genomics and its applications to the molecular understanding of human disease in the coming decade\cite{Gasperini2020}.
+Single cell CRISPR screens have been scaled up through multiplexing to systematically map gene-enhancer relationships,\cite{Xie2017,Gasperini2019,Xie2019} applied using combinatorial perturbation libraries to study genetic interactions,\cite{Norman2019} and extended to single cell protein expression readout.\cite{Mimitou2019,Frangieh2020} Given their scale, resolution, and versatility, these assays are likely to be a major driver of discovery in functional genomics and its applications to the molecular understanding of human disease in the coming decade.\cite{Gasperini2020}
 
@@
 In this work, we explore statistical challenges of single cell CRISPR screens and present a novel analysis methodology to address them.
-In addition to known scRNA-seq analysis challenges such as expression normalization and discreteness\cite{Townes2019,Hafemeister2019,Svensson2020}, we identify a key challenge that sets CRISPR screens apart from traditional differential expression experiments: the ``treatment"---in this case the presence of a CRISPR perturbation in a given cell---is subject to measurement error\cite{Dixit2016,Hill2018,Replogle2020}.
-In fact, underlying this measurement error are the same technical factors contributing to errors in the measurement of gene expression, including sequencing depth and batch effects.
+In addition to known scRNA-seq analysis challenges such as expression normalization and discreteness,\cite{Townes2019,Hafemeister2019,Svensson2020} we identify a key challenge that sets CRISPR screens apart from traditional differential expression experiments: the ``treatment"---in this case the presence of a CRISPR perturbation in a given cell---is subject to measurement error.\cite{Dixit2016,Hill2018,Replogle2020} In fact, underlying this measurement error are the same technical factors contributing to errors in the measurement of gene expression, including sequencing depth and batch effects.
 These technical factors therefore act as confounders, invalidating traditional nonparametric calibration approaches such as those based on ranks or permutations.
@@
 To address these challenges, we propose SCEPTRE (analysis of Single CEll PerTurbation screens via conditional REsampling; pronounced ``scepter'').
-This resampling methodology is based on the conditional randomization test\cite{CetL16}, a recently proposed statistical methodology which, like traditional nonparametric methods, does not rely on correct specification of the outcome (expression) model.
+This resampling methodology is based on the conditional randomization test,\cite{CetL16} a recently proposed statistical methodology which, like traditional nonparametric methods, does not rely on correct specification of the outcome (expression) model.
 Unlike traditional nonparametric methods, SCEPTRE employs a resampling scheme that explicitly accounts for heterogeneity in technical factors among cells.
-We analyze three datasets -- that of Gasperini et al.\cite{Gasperini2019}, that of Xie et al.\cite{Xie2019}, and a synthetic dataset -- to illustrate the aforementioned analysis challenges and the performance of the proposed methodology.
-SCEPTRE demonstrates excellent calibration and reveals many novel regulatory relationships supported by a variety of existing sources of functional evidence.
+We analyze data from two recent single cell pooled CRISPR screen experiments--one produced by Gasperini et al.\cite{Gasperini2019} and one produced by Xie et al.\cite{Xie2019}--to illustrate the aforementioned analysis challenges and the performance of the proposed methodology.
+SCEPTRE demonstrates excellent calibration, specificity, and sensitivity on the data and reveals many novel regulatory relationships supported by a variety of existing sources of functional evidence.
 
@@
 Gasperini et al.\ obtained $p$-values from a DESeq2\cite{Love2014}-inspired negative binomial regression analysis implemented in Monocle2.\cite{Qiu2017} Xie et al., in contrast, obtained $p$-values from Virtual FACS, a nonparametric rank-based method proposed by the authors.\cite{Xie2017} Gasperini et al.\ and Xie et al.\  observed miscalibration in their respective sets of $p$-values.
-Gasperini et al.\ examined the distribution of negative binomial $p$-values obtained from pairing each protein-coding gene with each of 50 non-targeting control gRNAs.
+Gasperini et al.\ examined the distribution of negative binomial $p$-values obtained from pairing each protein-coding gene with each of 50 non-targeting control (or negative control) gRNAs.
 These ``null'' $p$-values exhibited inflation, deviating substantially from the expected uniform distribution (Figure \ref{fig:analysis-challenges}a, top panel).
@@
 This confounding effect, if not explicitly accounted for, results in test miscalibration.
-Confounding is especially problematic for nonparametric methods like Virtual FACS, which implicitly (and incorrectly) assume that there are no technical factors that impact both gRNA detection and gene expression measurement.
+Confounding is especially problematic for traditional nonparametric methods like Virtual FACS, which implicitly (and incorrectly) assume that there are no technical factors that impact both gRNA detection and gene expression measurement.
 
@@
 
-Gasperini et al.\ and Xie et al.\ adjusted their analyses to remedy the observed calibration problems.
-Gasperini et al.\ calibrated the candidate enhancer $p$-values against the distribution of all gene-negative control $p$-values instead of the uniform distribution.
+Gasperini et al.\ and Xie et al.\ introduced ad hoc adjustments to their analyses to remedy the observed calibration issues.
+Gasperini et al.\ calibrated the candidate enhancer $p$-values against the distribution of all negative control gRNA $p$-values instead of the uniform distribution.
 The resulting ``empirical $p$-values'' were used for determining significance.
@@
 These analysis adjustments, unfortunately, did not satisfactorily improve calibration.
-The adjustment of Gasperini et al., for example, lead to overcorrection for some gene-enhancer pairs (false negatives) and undercorrection for others (false positives) (Figure 1s).
-Methods that avoid reliance on ad hoc calibration fixes might produce more consistently calibrated $p$-values across entire CRISPR screen datasets.
+The adjustment employed by Gasperini et al., for example, leads to overcorrection for some gene-enhancer pairs (false negatives) and undercorrection for others (false positives) (Figure \ref{fig:gasp_correction}).
+Methods that avoid reliance on ad hoc calibration fixes likely would produce more consistently calibrated $p$-values across entire CRISPR screen datasets.
 
@@
 These sets of $p$-values deviate from the expected uniform distribution, indicating test miscalibration.
-\textbf{b-d}, Sequencing depth (total mRNA UMIs per cell) impacts gRNA detection and observed gene expression levels in both Gasperini et al.\ (\textbf{b}) and Xie et al.\ (\textbf{c}) data.
-Thus, sequencing depth acts as a confounder (\textbf{d}).
+\textbf{b-d}, Sequencing depth (total mRNA UMIs per cell) impacts gRNA detection and observed gene expression levels in both Gasperini et al.\ (b) and Xie et al.\ (c) data.
+Thus, sequencing depth acts as a confounder (d).
 This analysis challenge distinguishes CRISPR screens from traditional differential expression applications.
@@
 This estimation procedure leads to miscalibration for high-dispersion genes.
-% \textbf{b-c}, Sequencing depth (total mRNA UMIs per cell) impacts gRNA detection (b) and observed expression levels, and therefore acts as a confounder (c).
-This analysis challenge distinguishes CRISPR screens from traditional differential expression applications.
-}
 }
@@
 We also applied the improved negative binomial approach to the Xie et al.\ ARL15-enh data.
-In the latter analysis we included sequencing depth, number of detected gRNA UMIs per cell, and sequencing batch as covariates.
-We found that the improved negative binomial model was better calibrated than the Monocle2 negative binomial model on the Gasperini data (Figure 3c; improved negative binomial denoted ``Hafemeister NB'').
-However, on both the Gasperini et al.\ data and Xie et al.\ data, the improved negative binomial model showed clear signs of p-value inflation (Figure 3c-d).
+We included sequencing depth, number of detected gRNA UMIs per cell, and sequencing batch as covariates in the analysis of the latter dataset.
+The improved negative binomial method exhibited better calibration than the Monocle negative binomial method on the Gasperini data (Figure \ref{fig:ground_truth}b).
+However, on both the Gasperini et al.\ data and Xie et al.\ data, the improved negative binomial method showed clear signs of p-value inflation (Figure \ref{fig:ground_truth}b-c).
 
@@
 
-% The difficulty of applying parametric models to single cell data highlights the appeal of nonparametric approaches for this purpose.
-\clearpage
 \paragraph{SCEPTRE: Analysis of single cell perturbation screens via conditional resampling.}
@@
 
-\clearpage
 \thispagestyle{empty}
@@
 \end{figure}
-\clearpage
 
@@
 We simulated expression data for a single gene in 1000 cells using four models selected from this class: the first with $\alpha = 1$, the second with $\alpha = 0.25$, the third with $\alpha = 5$, and the last with $\alpha = 1,$ but with 25$\%$ zero-inflation.
-We also simulated negative control gRNA data using a logistic regression model with the same covariates as in the gene expression model.
+We also simulated negative control gRNA data using a logistic regression model with the same covariates as the gene expression model.
 We assessed the calibration of three methods across the four simulated datasets: SCEPTRE, improved negative binomial regression, and scMAGeCK-LR,\cite{Yang2020} a recently-proposed, permutation-based nonparametric method.
@@
 We note that the empirical correction employed by Gasperini et al.\ limits the accuracy of $p$-values to about $10^{-6}$, explaining at least part of the difference.
-Finally, we compared the SCEPTRE results for ARL15-enh to the results of an arrayed CRISPR screen of ARL15-enh in which populations of perturbed cells were compared to populations of unperturbed cells via bulk RNA-seq.\cite{Xie2019} Both SCEPTRE and the bulk RNA-seq differential expression analysis rejected gene \textit{ARL15} (and only gene \textit{ARL15}) at an FDR of 0.1 after a Benjamini-Hochberg correction, increasing our confidence in sensitivity (and specificity) of SCEPTRE (Figure \ref{fig:ground_truth}e).
+Finally, we compared the SCEPTRE ARL15-enh results to the results of an arrayed CRISPR screen of ARL15-enh in which populations of perturbed cells were compared to populations of unperturbed cells via bulk RNA-seq.\cite{Xie2019} Both SCEPTRE and the bulk RNA-seq differential expression analysis rejected gene \textit{ARL15} (and only gene \textit{ARL15}) at an FDR of 0.1 after a Benjamini-Hochberg correction, increasing our confidence in sensitivity (and specificity) of SCEPTRE (Figure \ref{fig:ground_truth}e).
 
-\newpage
-\begin{figure}[h!]
+\clearpage
+\begin{figure*}[h!]
 	\includegraphics[width = \textwidth]{figures/Figure3/Figure3_cropped.pdf}
-	\caption{\textbf{SCEPTRE demonstrates excellent calibration, specificity, and sensitivity when the ground truth is known.}}
+ \end{figure*}
+\captionof{figure}[short caption here]{
 	\label{fig:ground_truth}
- \end{figure}
-\clearpage
-
-% \clearpage
-% \thispagestyle{empty} 
-% \addtocounter{page}{-1}
-%\begin{figure}[h!]
-%	\includegraphics[width = \textwidth]{figures/Figure3/Figure3.png}
-%	\caption{\textbf{SCEPTRE improves on the sensitivity and specificity of the parametric approach by correcting for expression model misspecification.} \textbf{a,} Numerical simulation comparing three ways of calibrating a negative binomial test statistic (parametric, permutation, conditional resampling), when the test statistic is based on the correct model or contains expression model misspecifications.
-Only the conditional resampling approach maintains calibration despite model misspecification.
-\textbf{b,c,} Application of SCEPTRE to Gasperini et al negative control gRNAs, comparing all gene-NTC pairs to the uniform distribution (b) or breaking down by NTC (c); details as in Figure~\ref{fig:calibration-comparison-all}a,c.
-The matching of SCEPTRE's colored region with the gray one in panel c, together with the overall near-uniformity in panel b, demonstrates its excellent calibration on real negative control data.
-\textbf{d,} SCEPTRE $p$-values for TSS-targeting controls generally more significant than the original empirical $p$-values; dashed line is the $45^\circ$ line.
-Note: $p$-values truncated for visualization, and empirical $p$-values in panel d at least 1/538,560 $\approx$ 2$\times 10^{-6}$ by construction.}
-%	\label{fig:calibration}
-% \end{figure}
-\clearpage
+	\textbf{SCEPTRE demonstrates excellent calibration, specificity, and sensitivity under known ground truth.} \textbf{a,} Numerical simulation comparing three methods -- SCEPTRE, improved negative binomial regression, and scMAGeCK-LR -- on four simulated datasets.
+Only SCEPTRE maintains calibration despite model misspecification and confounder presence.
+\textbf{b-c},  	Application of SCEPTRE, improved negative binomial regression, Monocle regression, and Virtual FACS to (b) all negative control gRNAs and all genes in the Gasperini et al.\ data, and (c) all ARL15-enh-targeting gRNAs and all genes in the Xie et al.\ data.
+Compared to the other methods, SCEPTRE shows excellent calibration.
+Gene \textit{ARL15} annotated in (c).
+\textbf{d,} SCEPTRE $p$-values for Gasperini et al.\ TSS-targeting controls are highly significant, and in general, more significant than the original empirical $p$-values.
+\textbf{e,} Comparison of $p$-values produced by SCEPTRE for ARL15-enh to $p$-values produced by an arrayed, bulk RNA-seq CRISPR screen of ARL15-enh.
+The results of the two analyses coincide exactly, with both aanlyses rejecting gene \textit{ARL15} (and only gene \textit{ARL15}) after a Benjamini-Hochberg correction.
+Dotted blue lines, rejection thresholds.}
+\noindent\hrulefill
 
@@
 Comparing the SCEPTRE $p$-values against the original empirical $p$-values (Figure~\ref{fig:application}a), we see that the two often diverge substantially.
-Our analysis found 200 gene-enhancer pairs that the original analysis did not, while 107 were found only by the original analysis.
-Many of the discoveries found only in the original analysis show signs of the $p$-value inflation observed in Figure~\ref{fig:analysis-challenges}.
+SCEPTRE found 200 gene-enhancer pairs that the original analysis did not, while 107 were found only by the original analysis.
+Many of the discoveries found only in the original analysis show signs of the $p$-value inflation observed in Figure~\ref{fig:gasp_correction}.
 
@@
 
+\clearpage
+\begin{figure}[h!]
+	\includegraphics[width = \textwidth]{figures/Figure4/Figure4_cropped.pdf}
+\end{figure}
+\captionof{figure}[short caption here]{
+	\label{fig:application}
+	\textbf{Application of SCEPTRE to Gasperini et al.\ data yields biologically plausible gene-enhancer links.} \textbf{a,} Comparison of the original empirical $p$-values to those obtained from SCEPTRE.
+The two analysis methods differ substantially, with 200 gene-enhancer links discovered only by SCEPTRE and 107 discovered only by the original.
+\textbf{b}, Five gene-enhancer pairs discovered by SCEPTRE but not the original analysis, each supported by a whole blood GTEx eQTL or FANTOM enhancer RNA correlation $p$-value.
+\textbf{c,} Distribution of distances from TSS to upstream paired enhancers.
+Compared to the original analysis, SCEPTRE pairs genes with nearer enhancers on average.
+\textbf{d,} For those gene-enhancer pairs falling in the same TAD, the cumulative distribution of the fractional rank of the HI-C interaction frequency compared to other distance-matched loci pairs within the same TAD.
+SCEPTRE shows similar enrichment despite finding 25\% more within-TAD pairs.
+Inset table shows gene-enhancer pairs falling in the same TAD.
+SCEPTRE finds 93 more total pairs, and a higher percentage of pairs fall in the same TAD.
+\textbf{e,} Enrichment of ChIP-seq signal from seven cell-type relevant transcription factors and one histone mark among paired enhancers.
+SCEPTRE exhibits greater enrichment across all ChIP-seq targets.}
+\noindent\hrulefill
+
 We also found that the total set of gene-enhancer pairs discovered was better enriched for regulatory biological signals, including HI-C and ChIP-seq.
@@
 We found similar levels of contact frequency enrichment, despite the fact that SCEPTRE discovered 85 more gene-enhancer pairs in the same TADs.
-Finally, we repeated the ChIP-seq enrichment analysis of Gasperini et al.
-for seven cell-type relevant transcription factors and H3K27ac.
+Finally, we repeated the ChIP-seq enrichment analysis of Gasperini et al.\ for seven cell-type relevant transcription factors and H3K27ac.
 We quantified enrichment of ChIP-seq signal in paired enhancers as the odds ratio that a candidate enhancer is paired to a gene, comparing those falling in the top quintile of candidate enhancers by ChIP-seq signal and those not overlapping a ChIP-seq peak at all.
@@
 
-\clearpage
-% \newgeometry{top=0.6in}
-% \thispagestyle{empty} 
-% \addtocounter{page}{-1}
-% \vspace{-2in}
-% \begin{figure}
-%	\includegraphics[width = \textwidth]{figures/Figure4/Figure4.png}
-%	\caption{\textbf{Application to Gasperini et al.
-data yields biologically plausible gene-enhancer links.}  \textbf{a,} Well-calibrated NTC $p$-values give confidence in SCEPTRE $p$-values for candidate enhancers (compare to Figure~\ref{fig:analysis-challenges}a).
-\textbf{b}, Comparing the original empirical $p$-values to those obtained from SCEPTRE.
-The two analysis methods differ substantially, with 217 gene-enhancer links discovered only by SCEPTRE and 102 discovered only by the original.
-\textbf{c}, Five gene-enhancer pairs discovered by SCEPTRE but not the original analysis, each supported by a whole blood GTEx eQTL or FANTOM enhancer RNA correlation $p$-value.
-\textbf{d,} Distribution of distances from TSS to upstream paired enhancers.
-Compared to the original analysis, SCEPTRE pairs genes with nearer enhancers on average.
-\textbf{e,} For those gene-enhancer pairs falling in the same TAD, the cumulative distribution of the fractional rank of the HI-C interaction frequency compared to other distance-matched loci pairs within the same TAD.
-SCEPTRE shows similar enrichment despite finding 32\% more within-TAD pairs.
-Inset table shows gene-enhancer pairs falling in the same TAD.
-SCEPTRE finds 115 more total pairs, and a higher percentage of pairs fall in the same TAD.
-\textbf{f,} Enrichment of ChIP-seq signal from seven cell-type relevant transcription factors and one histone mark among paired enhancers.
-SCEPTRE exhibits greater enrichment across all ChIP-seq targets.}
-%	\label{fig:application}
-% \end{figure}
-
-\begin{figure}
-\includegraphics[width = \textwidth]{figures/Figure4/Figure4_cropped.pdf}
-	\caption{\textbf{Application to Gasperini et al. data yields biologically plausible gene-enhancer links.} }
-	\label{fig:application}
-\end{figure}
-
-\clearpage
-% \restoregeometry
-
 \section*{Discussion}
 
-Here we illustrate a variety of statistical challenges arising in the analysis of single cell CRISPR screens, leaving existing methods (based on parametric expression models, permutations, or negative control data) vulnerable to calibration issues.
+In this paper we illustrate a variety of statistical challenges arising in the analysis of single cell CRISPR screens, leaving existing methods (based on parametric expression models, permutations, or negative control data) vulnerable to calibration issues.
 To address these challenges, we develop SCEPTRE, a novel resampling method based on modeling the probability a given gRNA will be observed in a given cell, based on that cell's technical factors.
-SCEPTRE exhibits excellent calibration despite the presence of confounding technical factors and despite misspecification of single cell gene expression models.
-We implement computational accelerations to bring the cost of our resampling-based methodology down to well within an order of magnitude of the traditional approach, making it quite feasible to apply for large-scale data.
-We use SCEPTRE to reanalyze the Gasperini et al.
-data and a subset of the Xie et al.
-data, yielding many new biologically plausible gene-enhancer relationships supported by evidence from eQTL, enhancer RNA co-expression, ChIP-seq, and HI-C data.
+SCEPTRE exhibits excellent calibration despite the presence of confounding technical factors and misspecification of single cell gene expression models.
+We implement computational accelerations to bring the cost of our resampling-based methodology down to well within an order of magnitude of the traditional negative binomial parametric approach, making it quite feasible to apply for large-scale data.
+We use SCEPTRE to reanalyze the Gasperini et al.\ data and a subset of the Xie et al.\ data, yielding many biologically plausible gene-enhancer relationships supported by evidence from eQTL, enhancer RNA co-expression, ChIP-seq, HI-C, and bulk CRISPR screen data.
 
 The main assumption behind SCEPTRE's validity is the accuracy of the model for gRNA observation.
-Our calibration results suggest that this is a reasonable assumption for the simulated and negative control data considered here.
+Our calibration results suggest that this is a reasonable assumption for the real and simulated data considered here.
 However, more work is necessary to assess the quality of gRNA measurement models across a broader range of contexts.
-Furthermore, we conjecture that misspecifications in the gRNA measurement model can be compensated by improved gene expression models, a statistical phenomenon referred to as ``double robustness"\cite{Robins2001,VanderLaan2003}.
-Better expression models can also improve the sensitivity of the conditional randomization test\cite{Katsevich2020a}.
+Furthermore, we conjecture that misspecifications in the gRNA measurement model can be compensated for by improved gene expression models, a statistical phenomenon referred to as ``double robustness."\cite{Robins2001,VanderLaan2003} Better expression models can also improve the sensitivity of the conditional randomization test.\cite{Katsevich2020a}
 
 There are several directions for improvement in the analysis of single cell CRISPR screens, which are not addressed by SCEPTRE.
-Many of these have been considered before, especially in the context of non-single-cell CRISPR screens.
-Such remaining challenges include variable effectiveness of gRNAs\cite{Li2015,Daley2018,Xie2019}, interactions among perturbed elements\cite{Xie2017,Zamanighomi2019,Norman2019}, and the limited resolution of CRISPR interference\cite{Hsu2018}.
-Additionally, single cell CRISPR screens would benefit from the adoption of techniques from the increasingly rich literature on scRNA-seq analysis.
+Many of these have been considered before, especially in the context of bulk CRISPR screens.
+Such remaining challenges include variable effectiveness of gRNAs,\cite{Li2015,Daley2018,Xie2019} interactions among perturbed elements,\cite{Xie2017,Zamanighomi2019,Norman2019} and the limited resolution of CRISPR interference.\cite{Hsu2018} Additionally, single cell CRISPR screens would benefit from the adoption of techniques from the increasingly rich literature on scRNA-seq analysis.
 
@@
 However, the capture rate can vary substantially across gRNAs.
-We look forward to adapting the SCEPTRE framework to the analysis of such new CRISPR screen technologies.
+The SCEPTRE framework could be extended to analyze such new CRISPR screen technologies.
 Finally, confounding technical factors like those pictured in Figure~\ref{fig:analysis-challenges}c are ubiquitous across genomics; we anticipate the conditional randomization test will complement traditional parametric and nonparametric methods in such contexts as a valuable analysis tool.
 
+\clearpage
 \bibliography{library}
-
-\clearpage
 
@@
 This work was supported, in part, by National Institute of Mental Health (NIMH) grants R01MH123184 and R37MH057881 as well as SFARI Grant 575547.
-Part of the data analysis used the Extreme Science and Engineering Discovery Environment (XSEDE)\cite{XSEDE}, which is supported by National Science Foundation grant number ACI-1548562.
-Specifically, it used the Bridges system\cite{Bridges}, which is supported by NSF award number ACI-1445606, at the Pittsburgh Supercomputing Center (PSC).
+Part of the data analysis used the Extreme Science and Engineering Discovery Environment (XSEDE),\cite{XSEDE} which is supported by National Science Foundation grant number ACI-1548562.
+Specifically, it used the Bridges system,\cite{Bridges} which is supported by NSF award number ACI-1445606, at the Pittsburgh Supercomputing Center (PSC).
 \item[Competing Interests] The authors declare that they have no
@@
 
+\clearpage
 \section*{Methods}
@@
 In practice, we must estimate the gRNA probabilities $\pi_i$ as well as the $p$-value $p_{\text{CRT}}$.
-This is because usually we do not know the distribution $X|Z$, and cannot compute the conditional probability in equation~\eqref{conditional-randomization} exactly.
+This is because usually we do not know the distribution $X|Z$ and cannot compute the conditional probability in equation~\eqref{conditional-randomization} exactly.
 We propose to estimate $\pi_i$ via logistic regression of $X$ on $Z$, and to estimate $p_{\text{CRT}}$ by resampling $\widetilde X$ a large number of times and then fitting a skew-$t$ distribution to the resampling null distribution $T(\widetilde X, Y, Z)|X,Y,Z$.
@@
 \end{enumerate}
-In our data analysis, we used $B = 500$ resamples.
-Following Gasperini et al.\, we based our analysis on the top two gRNAs targeting each enhancer.
+In our data analysis we used $B = 500$ resamples.
+Following Gasperini et al., we based our analysis on the top two gRNAs targeting each enhancer.
 Some enhancers were also targeted with two additional gRNAs, but we excluded these from the analysis.
@@
 We simulated one gene $Y_i,$ five gRNAs $X_{i1}, X_{i2}, \dots, X_{i5},$ and two confounders $Z_{i1}, Z_{i2}$ in $n = 1000$ cells.
-We generated the confounders $Z_{i1}$ and $Z_{i2}$ by sampling with replacement the sequence batch IDs and log-transformed sequencing depths, respectively, of the cells in the Gasperini dataset.
+We generated the confounders $Z_{i1}$ and $Z_{i2}$ by sampling with replacement the batch IDs and log-transformed sequencing depths of the cells in the Gasperini dataset.
 The batch ID confounder $Z_{i1}$ was a binary variable, as the Gasperni data included two batches.
@@
 We applied SCEPTRE, negative binomial regression, and scMAGeCK-LR\cite{Yang2020} to the four problem settings, each with $n_{sim} = 500$ repetitions.
-The negative binomial method was based on the $z$ statistic from the Hafemeister-inspired negative binomial model (\ref{negative-binomial}) with $\alpha = 1$.
+The negative binomial method, and in turn SCEPTRE, was based on the $z$ statistic from the Hafemeister-inspired negative binomial model (\ref{negative-binomial}) with $\alpha = 1$.
 scMAGeCK-LR differs from SCEPTRE and the negative binomial method in that scMAGeCK-LR computes $p$-values for all enhancers simultaneously.
@@
 Furthermore, the 664 total gene-enhancer pairs reported in the original analysis were the result of a Benjamini-Hochberg FDR correction that included not only the candidate enhancers but also hundreds of positive controls.
-While Bonferroni corrections can only become more conservative when including more hypotheses, BH corrections are known to become anticonservative when extra positive controls are included\cite{FR01}.
-To avoid this extra risk of false positives, we chose to use the ``high-confidence" set throughout.
+While Bonferroni corrections can only become more conservative when including more hypotheses, BH corrections are known to become anticonservative when extra positive controls are included.\cite{FR01} To avoid this extra risk of false positives, we chose to use the ``high-confidence" set throughout.
 
@@
 All analysis was performed on publicly available data.
-The CRISPR screen data\cite{Gasperini2019} is available at \url{https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE120861}.
-The ChIP-seq data are drawn from the ENCODE project\cite{Dunham2012} and are available at \url{https://www.encodeproject.org/}.
-The HI-C enrichment analysis is based on the data from Rao et al\cite{Rao2014}, available at \url{https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE63525}.
-The eQTL and eRNA co-expression $p$-values are taken from the GeneHancer database\cite{Fishilevich2017}, available as part of GeneCards (\url{https://www.genecards.org/}).
+The Gasperini et al.\ CRISPR screen data\cite{Gasperini2019} are available at \url{www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE120861}.
+The Xie et al.\ single-cell and bulk CRISPR screen data are available at \url{www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE129837}.
+The ChIP-seq data are taken from the ENCODE project\cite{Dunham2012} and are available at \url{www.encodeproject.org/}.
+The HI-C enrichment analysis is based on the data from Rao et al.,\cite{Rao2014} available at \url{www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE63525}.
+The eQTL and eRNA co-expression $p$-values are taken from the GeneHancer, database\cite{Fishilevich2017} available as part of GeneCards (\url{www.genecards.org/}).
 
@@
 
-Code to reproduce all data analysis is available on Github at \url{https://github.com/ekatsevi/SCEPTRE}.
+The \texttt{sceptre} R package and data analysis scripts for this paper are available on Github at \url{https://github.com/Timothy-Barry/SCEPTRE}.
+
+\clearpage 
+\appendix
+
+\section*{Supplementary Figures}
+\thispagestyle{empty}
+\setcounter{figure}{0}    
+\renewcommand{\thefigure}{S\arabic{figure}}
+
+\begin{figure}[h!]
+	\centering
+	\includegraphics[width = \textwidth]{figures/FigureS1/FigureS1.png}
+	\caption{\textbf{Gasperini et al.'s empirical correction is insufficient to correct for miscalibration}.
+\textbf{a,} Dispersion estimation procedure employed leads to miscalibration for high-dispersion genes, which the empirical correction does not adequately correct for, as measured by KS test applied to empirical $p$-values per gene (point colors).
+\textbf{b,} Raw $p$-values already well-calibrated for \textit{LUC7L2} gene, so empirical correction unnecessarily shrinks the significance of the association with TSS-targeting gRNA, depicted by horizontal lines, by three orders of magnitude.
+\textbf{c,} Empirical correction not strong enough for \textit{HIST1H1D}, which is among circled genes in panel a, which have an NTC-based miscalibration $p$-value smaller than the Bonferroni threshold.
+\textbf{d}, Under-correction leads to two potential false discoveries for \textit{HIST1H1D}.
+Dashed horizontal line represents the multiple testing threshold.}
+	\label{fig:gasp_correction}
+\end{figure}
 
@@
 
-\appendix
-%\newgeometry{top=1in}
-% \thispagestyle{empty} 
-% \addtocounter{page}{-1}
-
-% \counterwithin{figure}{section}
-\section{Supplementary Figures}
-\renewcommand{\thesection}{S}
-
-\vspace{0.25in}
-
-%\clearpage
-
-%\makeatletter 
-%\renewcommand{\thefigure}{S\@arabic\c@figure}
-%\makeatother
-
-%\section*{Supplementary Figures}
 \thispagestyle{empty} 
-%\vspace{-0.1in}
-%\setcounter{figure}{0}    
-%\renewcommand{\thefigure}{S\arabic{figure}}
-
-%\begin{figure}[h!]
-%	\centering
-%	\includegraphics[width = \textwidth]{figures/FigureS1/FigureS1.png}
-%	\caption{Gasperini et al.'s empirical correction is insufficient to correct for miscalibration.
-\textbf{a,} Dispersion estimation procedure employed leads to miscalibration for high-dispersion genes, which the empirical correction does not adequately correct for, as measured by KS test applied to empirical $p$-values per gene (point colors).
-\textbf{b,} Raw $p$-values already well-calibrated for LUC7L2 gene, so empirical correction unnecessarily shrinks the significance of the association with TSS-targeting gRNA, depicted by horizontal lines, by three orders of magnitude.
-\textbf{c,} Empirical correction not strong enough for HIST1H1D, which is among circled genes in panel a, which have an NTC-based miscalibration $p$-value smaller than the Bonferroni threshold.
-\textbf{d}, Under-correction leads to two potential false discoveries for HIST1H1D.
-Dashed horizontal line represents the multiple testing threshold.}
-%	\label{fig:mosaicseq}
-% \end{figure}
+\begin{figure}[h!]
+	\includegraphics[width = \textwidth]{figures/FigureS2/FigureS2.png}
+	\caption{\textbf{Comparison of negative binomial and conditional resampling $p$-values based on the same test statistic.} \textbf{a,} The standard parametric negative binomial $p$-value versus that obtained from the same test statistic by conditional resampling, for each gene / candidate enhancer pair (both truncated at $10^{-10}$ for visualization).
+The two can diverge fairly substantially.
+\textbf{b-d,} Parametric and conditional resampling null distributions for the negative binomial $z$-value in three cases: the conditional resampling $p$-value is more significant (b), the parametric $p$-value is more significant (c), the two $p$-values are about the same (d).}
+	\label{fig:NB-vs-CP}
+ \end{figure}
 
  \clearpage
-
-% \thispagestyle{empty} 
-% \begin{figure}[h!]
-%	\includegraphics[width = \textwidth]{figures/FigureS2/FigureS2.png}
-%	\caption{\textbf{Comparing negative binomial and conditional resampling $p$-values based on the same test statistic.} \textbf{a,} The standard parametric negative binomial $p$-value versus that obtained from the same test statistic by conditional resampling, for each gene / candidate enhancer pair (both truncated at $10^{-10}$ for visualization).
-The two can diverge fairly substantially.
-\textbf{b-d,} Parametric and conditional resampling null distributions for the negative binomial $z$-value in three cases: the parametric $p$-value is more significant (b), the conditional resampling $p$-value is more significant (c), the two $p$-values are about the same (d).}
-%	\label{fig:NB-vs-CP}
-% \end{figure}
-
-%\restoregeometry
-
-% \clearpage
-% \thispagestyle{empty} 
-% \begin{figure}[h!]
-%	\includegraphics[width = \textwidth]{figures/FigureS3/FigureS3.pdf}
-%	\caption{\textbf{Details on ChIP-seq enrichment analysis.} Fraction of candidate enhancers paired with a gene, broken down by quintile of ChIP-seq signal (0 means the candidate enhancer did not overlap a ChIP-seq peak), based on which the odds ratios in Figure~\ref{fig:application}f were computed.
+ \thispagestyle{empty} 
+\begin{figure}[h!]
+	\includegraphics[width = \textwidth]{figures/FigureS3/FigureS3.pdf}
+	\caption{\textbf{Details on ChIP-seq enrichment analysis.} Fraction of candidate enhancers paired with a gene, broken down by quintile of ChIP-seq signal (0 means the candidate enhancer did not overlap a ChIP-seq peak), based on which the odds ratios in Figure~\ref{fig:application}f were computed.
 Both methods generally pair candidate enhancers in higher ChIP-seq quintiles more frequently, but this enrichment is more pronounced in SCEPTRE across all eight ChIP-seq targets.}
-%	\label{fig:chipseq-enrichment-details}
-% \end{figure}
+	\label{fig:chipseq-enrichment-details}
+ \end{figure}
 

### 2020-11-29 sceptre_paper/plotting/Figure3.R

@@ -170,7 +170,7 @@ p_e <- ggplot(data = to_plot, mapping = aes(x = bulk_pval, y = sceptre_pval, col
   annotate(geom = "text", x = 1e-12, y = 1e-14, label = "ARL15", col = "firebrick3")
 
 # Legend
-ex <- tibble(x = 1:5, y = 1:5, Method = c("SCEPTRE", "Monocle NB", "Hafemeister NB", "Virtual FACS", "scMAGeCK")) %>% mutate(Method = factor(x = Method, levels = c("SCEPTRE", "Monocle NB", "Hafemeister NB", "Virtual FACS", "scMAGeCK"), labels = c("SCEPTRE", "Monocle NB", "Hafemeister NB", "Virtual FACS", "scMAGeCK")))
+ex <- tibble(x = 1:5, y = 1:5, Method = c("SCEPTRE", "Monocle NB", "Improved NB", "Virtual FACS", "scMAGeCK")) %>% mutate(Method = factor(x = Method, levels = c("SCEPTRE", "Monocle NB", "Improved NB", "Virtual FACS", "scMAGeCK"), labels = c("SCEPTRE", "Monocle NB", "Improved NB", "Virtual FACS", "scMAGeCK")))
 p_vert <- ggplot(data = ex, mapping = aes(x = x, y = y, col = Method)) + geom_point() + scale_colour_manual(values = setNames(plot_colors[c("sceptre", "gasperini_nb", "hf_nb", "hypergeometric", "scMAGeCK")], NULL), name = "Method") + theme_bw() + theme(legend.title = element_blank())
 legend <- get_legend(plot = p_vert)
 ggsave(plot = legend, filename = paste0(manuscript_figure_dir, "/Figure3/vert_legend.pdf"), width = 1.5, height = 1.5)

### 2020-12-03 sceptre_paper/manuscript/manuscript_v3.tex
remove reliance on defunct ravel package.

@@
 We included sequencing depth, number of detected gRNA UMIs per cell, and sequencing batch as covariates in the analysis of the latter dataset.
-The improved negative binomial method exhibited better calibration than the Monocle negative binomial method on the Gasperini data (Figure \ref{fig:ground_truth}b).
+The improved negative binomial method exhibited better calibration than the Monocle negative binomial method on the Gasperini et al.\ data (Figure \ref{fig:ground_truth}b).
 However, on both the Gasperini et al.\ data and Xie et al.\ data, the improved negative binomial method showed clear signs of p-value inflation (Figure \ref{fig:ground_truth}b-c).
@@
 
-\paragraph{SCEPTRE produces correct results when the ground truth is known.}
+\paragraph{SCEPTRE shows good calibration, sensitivity, and specificity on real and simulated data.}
 
-We assessed the calibration, sensitivity, and specificity of SCEPTRE in several settings in which the ground truth was known.
+%We assessed the calibration, sensitivity, and specificity of SCEPTRE in several settings in which the ground truth was known.
 First, we demonstrated the calibration of SCEPTRE in a small, proof-of-concept simulation study (Figure \ref{fig:ground_truth}a).
 We considered a class of negative binomial regression models with dispersion $\alpha$ and two technical covariates (sequencing depth and batch).
-We simulated expression data for a single gene in 1000 cells using four models selected from this class: the first with $\alpha = 1$, the second with $\alpha = 0.25$, the third with $\alpha = 5$, and the last with $\alpha = 1,$ but with 25$\%$ zero-inflation.
+We simulated expression data for a single gene in 1000 cells using four models selected from this class: the first with $\alpha = 1$, the second with $\alpha = 0.2$, the third with $\alpha = 5$, and the last with $\alpha = 1,$ but with 25$\%$ zero-inflation.
 We also simulated negative control gRNA data using a logistic regression model with the same covariates as the gene expression model.
@@
 	\label{fig:ground_truth}
-	\textbf{SCEPTRE demonstrates excellent calibration, specificity, and sensitivity under known ground truth.} \textbf{a,} Numerical simulation comparing three methods -- SCEPTRE, improved negative binomial regression, and scMAGeCK-LR -- on four simulated datasets.
+	\textbf{SCEPTRE demonstrates good calibration, specificity, and sensitivity under known ground truth.} \textbf{a,} Numerical simulation comparing three methods -- SCEPTRE, improved negative binomial regression, and scMAGeCK-LR -- on four simulated datasets.
 Only SCEPTRE maintains calibration despite model misspecification and confounder presence.
@@
 
+\clearpage
 \section*{Data availability}

### 2020-12-03 sceptre_paper/analysis_drivers/analysis_drivers_xie/check_packages_0.R

@@ -3,10 +3,10 @@ args <- commandArgs(trailingOnly = TRUE)
 code_dir <- if (is.na(args[1])) "/Users/timbarry/Box/SCEPTRE/sceptre_paper/" else args[1]
 source(paste0(code_dir, "/sceptre_paper/utilities/verify_all_packages_available.R"))
 
-packages <- c("furrr", "Seurat", "R.matlab", "R.utils", "fst", "sn", "MASS", "VGAM", "tidyverse", "bigstatsr", "openxlsx", "rhdf5", "ravel", "sceptre", "katsevich2020")
-locs <- c(rep("CRAN", 11), "Bioc", rep("github", 3))
-github_repo <- c(rep(NA, 12), "Timothy-Barry/ravel", "Timothy-Barry/sceptre_paper", "Timothy-Barry/sceptre_paper")
-github_repo_subdir <- c(rep(NA, 13), "sceptre", "katsevich2020")
+packages <- c("furrr", "Seurat", "R.matlab", "R.utils", "fst", "sn", "MASS", "VGAM", "tidyverse", "bigstatsr", "openxlsx", "rhdf5", "sceptre", "katsevich2020")
+locs <- c(rep("CRAN", 11), "Bioc", rep("github", 2))
+github_repo <- c(rep(NA, 12), "Timothy-Barry/sceptre_paper", "Timothy-Barry/sceptre_paper")
+github_repo_subdir <- c(rep(NA, 12), "sceptre", "katsevich2020")
 df <- data.frame(package = packages, loc = locs, github_repo = github_repo, github_repo_subdir = github_repo_subdir)
 
 verify_all_packages_available(df)

### 2020-12-03 sceptre_paper/analysis_drivers/analysis_drivers_xie/pre_process_data_3.R

@@ -2,7 +2,7 @@
 args <- commandArgs(trailingOnly = TRUE)
 code_dir <- if (is.na(args[1])) "/Users/timbarry/Box/SCEPTRE/sceptre_paper/" else args[1]
 source(paste0(code_dir, "/sceptre_paper/analysis_drivers/analysis_drivers_xie/paths_to_dirs.R"))
-packs <- c("future", "furrr", "Matrix", "rhdf5", "stringr", "openxlsx", "ravel")
+packs <- c("future", "furrr", "Matrix", "rhdf5", "stringr", "openxlsx", "katsevich2020")
 for (pack in packs) suppressPackageStartupMessages(library(pack, character.only = TRUE))
 
 ###################
@@ -153,7 +153,7 @@ write.fst(cell_covariate_matrix, paste0(processed_dir, "/cell_covariate_matrix.f
 ##############
 bulk_info <- read.xlsx(xlsxFile = paste0(raw_data_dir, "/bulk_rna_info.xlsx"), sheet = 3) %>% select(library_name = Library.Name, gRNA = sgRNA, region = Region, biological_duplicate = Biological.Duplicate)
 bulk_info_arl15_enh <- slice(bulk_info, 1:25)
-bulk_info_myb_enh3 <- slice(bulk_info, 26:49) 
+bulk_info_myb_enh3 <- slice(bulk_info, 26:49)
 
 bulk_df <- suppressWarnings(read_tsv(file = paste0(raw_data_dir, "/GSE129825_Libraries.FeatureCounts.ARL15_enhancer.txt"), col_types = "ccccciiiiiiiiiiiiiiiiiiiiiiiiii")) %>% rename("PZ788" = "PZ778", "PZ778" = "PZ778_1")
 bulk_df_arl15_enh <- filter(bulk_df, Geneid %in% all_protein_coding_genes)

### 2020-12-03 sceptre_paper/katsevich2020/R/processing_functions.R

@@ -1,0 +1,69 @@
+# Processing functions
+
+#' Extract vector from compressed matrix
+#'
+#' A function to extract a vector(s) from a matrix stored in compressed column or row format
+#'
+#' @param vector_nos indexes of the vectors to extract
+#' @param x sparse-matrix data
+#' @param i sparse-matrix indexes
+#' @param p sparse-matrix pointers
+#' @param vector_length length of the vectors to extract
+#' @param zero_based_idx (boolean; default TRUE) do i and p use zero-based indexing?
+#' @param vector_subset (optional) integers indicating a subset of the vectors to return
+#'
+#' @return the extracted vectors
+#' @export
+extract_vector_from_compressed_matrix <- function(vector_nos, x, i, p, vector_length, zero_based_idx = TRUE, vector_subset = NULL) {
+  if (zero_based_idx) {
+    i <- i + 1
+    p <- p + 1
+  }
+  f <- function(no) {
+    curr_ptr_range <- p[no:(no + 1)] + c(0,-1)
+    out <- rep(0, vector_length)
+    if (curr_ptr_range[1] <= curr_ptr_range[2]) {
+      x_out <- x[curr_ptr_range[1]:curr_ptr_range[2]]
+      i_out <- i[curr_ptr_range[1]:curr_ptr_range[2]]
+      out[i_out] <- x_out
+    }
+    if (!is.null(vector_subset)) out <- out[vector_subset]
+    return(out)
+  }
+  ret_val <- if (length(vector_nos) == 1) f(vector_nos) else sapply(X = vector_nos, FUN = f)
+  return(ret_val)
+}
+
+
+#' Extract column from compressed sparse column matrix
+#'
+#' @param col_nos idxs of the columns to extract
+#' @param csc_mat a list representing a matrix stored in csc form containing the entries p, i, x, and dim.
+#' @param zero_based_idx (boolean) do i and p use zero-based indexing
+#' @param row_subset (optional) subset the extracted columns by row index
+#'
+#' @return a matrix containing the extracted columns
+#' @export
+extract_column_from_csc_matrix <- function(col_nos, csc_mat, zero_based_idx = TRUE, row_subset = NULL) {
+  p <- csc_mat$p
+  i <- csc_mat$i
+  x <- csc_mat$x
+  n_rows <- csc_mat$dim[1]
[…]

### 2020-12-04 sceptre_paper/analysis_drivers/analysis_drivers_gasp/check_directory_structure_1.R
update bash scripts

@@ -12,8 +12,4 @@ sub_dirs <- c(create_parent_directories("data/gasperini/raw"), create_parent_dir
   "figures", create_parent_directories("logs/gasperini")) %>% unique()
 
 dirs_to_create <- paste0(offsite_dir, "/", sub_dirs)
-for (directory in dirs_to_create) {
-  if (!dir.exists(directory)) {
-    dir.create(directory)
-  }
-}
+check_directories(dirs_to_create)

### 2020-12-04 sceptre_paper/analysis_drivers/analysis_drivers_xie/check_directory_structure_1.R

@@ -10,8 +10,4 @@ sub_dirs <- c(create_parent_directories("data/xie/raw"), create_parent_directori
               create_parent_directories("results/xie/sceptre"), "results/xie/negative_binomial", "results/xie/bulk_rna_seq", "figures", create_parent_directories("logs/xie")) %>% unique()
 
 dirs_to_create <- paste0(offsite_dir, "/", sub_dirs)
-for (directory in dirs_to_create) {
-  if (!dir.exists(directory)) {
-    dir.create(directory)
-  }
-}
+check_directories(dirs_to_create)

### 2020-12-04 sceptre_paper/analysis_drivers/analysis_drivers_xie/download_data_2.R

@@ -43,7 +43,7 @@ download.file(url = "https://ars.els-cdn.com/content/image/1-s2.0-S2211124719313
 dest <- paste0(raw_data_dir, "/bulk_rna_info.xlsx")
 download.file(url = "https://ars.els-cdn.com/content/image/1-s2.0-S2211124719313956-mmc3.xlsx", destfile = dest)
 
-# Download the Genes.xlsx file from this url: https://osf.io/edjzv/ (source: Human protein-coding genes and gene feature statistics in 2019 by Piovesan et al in BMC Research Notes)
+# Download the Genes.xlsx file from this url: https://osf.io/edjzv/ (source: Human protein-coding genes and gene feature statistics in 2019 by Piovesan et al in BMC Research Notes). 
 
 ###############################
 # 4. Results from other authors

### 2020-12-04 sceptre_paper/plotting/Tims_plotting/arl15_bulk_rnaseq_validation_2.R

@@ -1,28 +1,0 @@
-args <- commandArgs(trailingOnly = TRUE)
-code_dir <- if (is.na(args[1])) "/Users/timbarry/Box/SCEPTRE/sceptre_paper/" else args[1]
-library(katsevich2020)
-source(paste0(code_dir, "/sceptre_paper/analysis_drivers/analysis_drivers_xie/paths_to_dirs.R"))
-source(paste0(code_dir, "/sceptre_paper/plotting/aux_objects.R"))
-
-p_vals_sceptre <- paste0(results_dir, "/all_results_with_names.fst") %>% read.fst() %>% as_tibble() %>% filter(enh_names == "ARL15-enh") %>% select(p_value, gene_names) %>% mutate(p_value_adj = p.adjust(p_value, method = "BH"))
-p_vals_bulk <- paste0(offsite_dir, "/results/xie/bulk_rna_seq/pvals_arl15_enh.rds") %>% readRDS() %>% as_tibble() %>% rename(gene_names = gene_id)
-to_plot <- inner_join(p_vals_bulk, p_vals_sceptre, by = "gene_names") %>% rename(bulk_pval_adj = p_value_adj.x, bulk_pval = p_value.x, sceptre_pval_adj = p_value_adj.y, sceptre_pval = p_value.y) %>% mutate(is_arl15 = (gene_names == "ARL15"))
-
-p_vals_sceptre %>% arrange(p_value_adj) %>% head()
-p_vals_bulk %>% arrange(p_value_adj) %>% head() 
-
-p <- ggplot(data = to_plot, mapping = aes(x = bulk_pval, y = sceptre_pval, col = is_arl15)) +
-  geom_point(alpha = 0.5) +
-  scale_colour_manual(values = c("grey60", "firebrick3")) +
-  scale_x_continuous(trans = revlog_trans(base = 10)) +
-  scale_y_continuous(trans = revlog_trans(base = 10)) +
-  theme_bw() + theme(text = element_text(size = 12),
-                     legend.background = element_rect(fill = "transparent", color = NA),
-                     panel.grid = element_blank(),
-                     panel.border = element_blank(),
-                     axis.line = element_line(), legend.position = "none") +
-                       xlab("Bulk RNA-seq p-value") + ylab("SCEPTRE p-value") +
[…]

### 2020-12-04 sceptre_paper/plotting/Tims_plotting/arl15_enh_qqplots_1.R

@@ -1,53 +1,0 @@
-args <- commandArgs(trailingOnly = TRUE)
-code_dir <- if (is.na(args[1])) "/Users/timbarry/Box/SCEPTRE/sceptre_paper/" else args[1]
-require(katsevich2020)
-source(paste0(code_dir, "/sceptre_paper/analysis_drivers/analysis_drivers_xie/paths_to_dirs.R"))
-source(paste0(code_dir, "/sceptre_paper/plotting/aux_objects.R"))
-
-p_vals_sceptre <- paste0(results_dir, "/all_results_with_names.fst") %>% read.fst() %>% filter(enh_names == "ARL15-enh")
-p_vals_nb <- paste0(results_dir_negative_binomial, "/all_results.fst") %>% read.fst() %>% filter(gRNA_id == as.character(p_vals_sceptre$gRNA_id[1]))
-p_vals_hypergeo <- paste0(processed_dir, "/xie_p_values.rds") %>% readRDS() %>% pluck("arl15_enh")
-p_vals_hypergeo <- p_vals_hypergeo[names(p_vals_hypergeo) %in% p_vals_sceptre$gene_id]
-to_plot <- tibble(method = rep(x = c("SCEPTRE", "NB regression", "Hypergeometric"), each = length(p_vals_hypergeo)) %>% factor(), p_value = c(p_vals_sceptre$p_value, p_vals_nb$p_value, set_names(p_vals_hypergeo, NULL) ), gene = c(as.character(p_vals_sceptre$gene_id), as.character(p_vals_nb$gene_id), names(p_vals_hypergeo)))
-
-truncate_thresh <- 1e-9
-qq_data <- to_plot %>%
-  rename(pvalue = p_value) %>%
-  group_by(method) %>%
-  mutate(r = rank(pvalue), expected = ppoints(n())[r],
-         clower = qbeta(p=(1-ci)/2, shape1 = r, shape2 = n()+1-r),
-         cupper = qbeta(p=(1+ci)/2, shape1 = r, shape2 = n()+1-r)) %>% ungroup()
-
-min_expected_p <- min(qq_data$expected)
-sorted_methods_rightmost_pnt <- filter(qq_data, expected == min_expected_p) %>% arrange(pvalue) %>% pull(method) %>% as.character()
-qq_data <- mutate(qq_data, pvalue = ifelse(pvalue <= truncate_thresh, truncate_thresh, pvalue))
[…]

### 2020-12-04 sceptre_paper/plotting/Tims_plotting/aux_objects.R

@@ -1,1 +1,0 @@
-plot_colors <- c(original_nb = "red", improved_nb = "violet", sceptre = "blue", hypergeometric = "turquoise", scMAGeCK = "magenta4")

### 2020-12-04 sceptre_paper/plotting/Tims_plotting/gasperini_qqplots_3.R

@@ -1,102 +1,0 @@
-# Load the gasperini, (HF) negative binomial, and sceptre results
-args <- commandArgs(trailingOnly = TRUE)
-code_dir <- if (is.na(args[1])) "/Users/timbarry/Box/SCEPTRE/sceptre_paper/" else args[1]
-require(katsevich2020)
-source(paste0(code_dir, "/sceptre_paper/analysis_drivers/analysis_drivers_gasp/file_paths_to_dirs.R"))
-source(paste0(code_dir, "/sceptre_paper/plotting/aux_objects.R"))
-
-gasperini_results <- paste0(processed_dir, "/original_results.fst") %>% read.fst()
-sceptre_results <- paste0(results_dir, "/resampling_results.fst") %>% read.fst()
-negbin_results <- paste0(results_dir_negative_binomial, "/likelihood_results.fst") %>% read.fst()
-
-df_NTC <- rbind(select(gasperini_results, gene_id, grna_group, pvalue = pvalue.raw, site_type) %>% mutate(method = "Original NB"),
-                select(sceptre_results, gene_id, grna_group, pvalue = p_value, site_type) %>% mutate(method = "SCEPTRE"),
-                select(negbin_results,  gene_id, grna_group, pvalue, site_type) %>% mutate(method = "Regularized NB")) %>% filter(site_type == "NTC") %>% mutate_at(.vars = c("gene_id", "grna_group", "site_type", "method"), .funs = factor)
-
-# compute information for overall QQ plot
-ci <- 0.95
-subsampling_factor <- 250
-
-df1 <- df_NTC %>%
-  group_by(method) %>%
-  mutate(r = rank(pvalue), expected = ppoints(n())[r],
-         clower = qbeta(p=(1-ci)/2, shape1 = r, shape2 = n()+1-r),
-         cupper = qbeta(p=(1+ci)/2, shape1 = r, shape2 = n()+1-r)) %>%
-  ungroup()
-
-# compute information for QQ plot by gRNA
-df2 <- df_NTC %>%
-  group_by(grna_group, method) %>%
-  mutate(r = rank(pvalue), expected = ppoints(n())[r],
-         clower = qbeta(p=(1-ci)/2, shape1 = r, shape2 = n()+1-r),
-         cupper = qbeta(p=(1+ci)/2, shape1 = r, shape2 = n()+1-r)) %>%
-  ungroup() %>%
[…]

### 2020-12-04 sceptre_paper/plotting/Tims_plotting/simulation_qqplot_4.R

@@ -1,40 +1,0 @@
-args <- commandArgs(trailingOnly = TRUE)
-code_dir <- if (is.na(args[1])) "/Users/timbarry/Box/SCEPTRE/sceptre_paper/" else args[1]
-offsite_dir <- if (is.na(args[2])) "/Volumes/tims_new_drive/research/sceptre_files" else args[2]
-require(katsevich2020)
-source(paste0(code_dir, "/sceptre_paper/plotting/aux_objects.R"))
-
-truncate_thresh <- 1e-7
-qq_data <- simulation_results %>%
-  group_by(method, dataset_id) %>%
-  mutate(r = rank(p_value), expected = ppoints(n())[r],
-         clower = qbeta(p=(1-ci)/2, shape1 = r, shape2 = n()+1-r),
-         cupper = qbeta(p=(1+ci)/2, shape1 = r, shape2 = n()+1-r)) %>%
-  ungroup() %>%
-  mutate(pvalue = ifelse(p_value < truncate_thresh, truncate_thresh, p_value),
-         facet_label = factor(x = as.character(dataset_id), levels = c("2", "1", "3", "4"), labels = c("Correct model", "Dispersion too large", "Dispersion too small", "Zero inflation")),
-         method = factor(x = as.character(method), levels = c("sceptre", "negative_binomial", "scMAGeCK"), labels = c("SCEPTRE", "Fixed dispersion NB", "scMAGeCK")))
-
-p <- qq_data %>%
-  ggplot(aes(x = expected, col = method, y = pvalue, ymin = clower, ymax = cupper)) +
-  geom_ribbon(alpha = 0.2, color = NA) +
-  geom_abline(intercept = 0, slope = 1) +
-  geom_point(size = 1, alpha = 0.5) +
-  xlab("Expected null p-value") +
-  ylab("Observed p-value") +
-  scale_colour_manual(values = setNames(plot_colors[c("sceptre", "fixed_dispersion_nb", "scMAGeCK")], NULL), name = "Method") +
-  guides(color = guide_legend(override.aes = list(alpha = 1))) +
-  scale_x_continuous(trans = revlog_trans(base = 10)) +
-  scale_y_continuous(trans = revlog_trans(base = 10)) +
-  facet_wrap(.~facet_label, nrow = 2) +
-  theme_bw() + theme(
-    panel.spacing.x = unit(1.25, "lines"),
-    plot.title = element_text(hjust = 0.5),
-    strip.background = element_blank(),
[…]

### 2020-12-04 sceptre_paper/plotting/check_packages_and_dirs_0.R

@@ -1,0 +1,19 @@
+args <- commandArgs(trailingOnly = TRUE)
+code_dir <- if (is.na(args[1])) "/Users/timbarry/Box/SCEPTRE/sceptre_paper/" else args[1]
+source(paste0(code_dir, "/sceptre_paper/utilities/verify_all_packages_available.R"))
+require(tidyverse)
+
+packages <- c("ggpubr", "cowplot")
+locs <- c("CRAN", "CRAN")
+github_repo <- c(NA, NA)
+github_repo_subdir <- c(NA, NA)
+
+# Verify package availability
+df <- data.frame(package = packages, loc = locs, github_repo = github_repo, github_repo_subdir = github_repo_subdir)
+verify_all_packages_available(df)
+
+# Create subdirectories
+fig_dir <- "sceptre_paper/manuscript/figures"
+fig_subdirs <- paste0("Figure",  c(as.character(1:4), "S1", "S2", "S3"))
+dirs_to_create <- c(paste0(code_dir, create_parent_directories(fig_dir)), paste0(code_dir, fig_dir, "/", fig_subdirs))
+check_directories(dirs_to_create)

### 2020-12-04 sceptre_paper/simulations/check_directories_0.R

@@ -1,11 +1,8 @@
 # Check directory structure
 args <- commandArgs(trailingOnly = TRUE)
-offsite_dir <- if (is.na(args[1])) "/Volumes/tims_new_drive/research/sceptre_files" else args[1]
-
+code_dir <- if (is.na(args[1])) "/Users/timbarry/Box/SCEPTRE/sceptre_paper/" else args[1]
+offsite_dir <- if (is.na(args[2])) "/Volumes/tims_new_drive/research/sceptre_files" else args[2]
+source(paste0(code_dir, "/sceptre_paper/utilities/verify_all_packages_available.R"))
 sub_dirs <- c("data/simulations", "results/simulations", "logs/simulations")
 dirs_to_create <- paste0(offsite_dir, "/", sub_dirs)
-for (directory in dirs_to_create) {
-  if (!dir.exists(directory)) {
-    dir.create(directory)
-  }
-}
+check_directories(dirs_to_create)

### 2020-12-04 sceptre_paper/utilities/verify_all_packages_available.R

@@ -31,3 +31,11 @@ create_parent_directories <- function(s) {
   out <- map(.x = 1:length(dirs), .f = function(i) paste0(dirs[1:i], collapse = "/"))
   return(unlist(out))
 }
+
+check_directories <- function(dirs_to_create) {
+  for (directory in dirs_to_create) {
+    if (!dir.exists(directory)) {
+      dir.create(directory)
+    }
+  }
+}

### 2020-12-15 sceptre_package/sceptre/data-raw/generating_example_data.R
new example vignette, data

@@ -1,0 +1,18 @@
+# This script generates example data for the SCEPTRE package consisting of one gRNA, one gene, and the covariate matrix. The data are pulled from Gasperini.
+
+args <- commandArgs(trailingOnly = TRUE)
+code_dir <- if (is.na(args[1])) "/Users/timbarry/Box/SCEPTRE/SCEPTRE/" else args[1]
+source(paste0(code_dir, "/sceptre_paper/analysis_drivers/analysis_drivers_gasp/file_paths_to_dirs.R"))
+source(paste0(code_dir, "sceptre_paper/analysis_drivers/analysis_drivers_gasp/sceptre_function_args.R"))
+
+gene_id <- "ENSG00000140400"
+gRNA_id <- "chr15.2442_top_two"
+# expression data
+expressions <- cell_gene_expression_matrix[,which(ordered_gene_ids == gene_id)]
+gRNA_indics <- read_fst(path = gRNA_indicator_matrix_fp, gRNA_id) %>% pull() %>% as.integer()
+# subset according to cell id
+expressions <- expressions[cell_subset]
+gRNA_indicators <- gRNA_indics[cell_subset]
+covariate_matrix <- covariate_matrix[cell_subset,]
+# Save the data
+usethis::use_data(expressions, gRNA_indicators, covariate_matrix)

### 2020-12-15 sceptre_package/sceptre/R/plotting_functions.R
include plot function in sceptre; update data to use in vignette; update vignette; update readme by fixing pacakge location

@@ -1,0 +1,39 @@
+#' Plot skew-t
+#'
+#' @param resampled_zvalues the set of resampled z-values
+#' @param original_zvalue the z value of the original negative binomial fit
+#' @param dpv the skew-t fit MLE
+#'
+#' @return a ggplot object containing the plot
+#' @export
+plot_skew_t <- function(resampled_zvalues, original_zvalue, dp, interval = c(-4,4)) {
+  z = seq(interval[1], interval[2], length.out = 1000)
+  df_curves = tibble(z = z, fitted = dst(x = z, dp = dp), gaussian = dnorm(z)) %>%
+    gather("curve", "y", fitted, gaussian) %>%
+    mutate(curve = factor(curve, levels = c("fitted","gaussian"), labels = c("Conditional\nrandomization","Negative\nbinomial")))
+  df_ribbon = df_curves %>%
+    filter(z <= original_zvalue, curve == "Conditional\nrandomization") %>%
+    mutate(lower = 0, upper = y) %>%
+    select(z, lower, upper)
+  p <- ggplot() +
+    geom_histogram(aes(x = z, y = ..density..),
+                   data = tibble(z = resampled_zvalues),
+                   boundary = 0, colour = "black", fill = "lightgray", bins = 15) +
+    geom_line(aes(x = z, y = y, group = curve, colour = curve, linetype = curve), data = df_curves) +
+    geom_vline(xintercept = original_zvalue, colour = "firebrick3", linetype = "solid") +
+    geom_ribbon(aes(x = z, ymin = lower, ymax = upper), fill = "darkorchid2", alpha = 0.5, data = df_ribbon) +
+    scale_colour_manual(values = c("darkorchid2", "black"), name = "Null distribution") +
+    scale_linetype_manual(values = c("solid", "dashed"), name = "Null distribution") +
+    scale_y_continuous(expand = c(0,0)) +
+    xlab("Negative binomial z-value") + theme_bw() +
+    theme(legend.position = c(0.85, 0.8),
+          legend.background = element_rect(fill = "transparent", colour = NA),
+          plot.title = element_text(hjust = 0.5, size = 11),
+          panel.grid = element_blank(),
+          panel.border = element_blank(),
+          axis.line.x = element_line(),
+          axis.ticks.y = element_blank(),
+          axis.title.y = element_blank(),
+          axis.text.y = element_blank())
+  return(p)
+}

### 2020-12-15 sceptre_package/sceptre/R/resampling_method_functions.R

@@ -127,14 +127,14 @@ run_sceptre_using_precomp <- function(expressions, gRNA_indicators, gRNA_precomp
 #' covariate_matrix <- if (nrow(covariate_matrix) == 106666) covariate_matrix else covariate_matrix[cell_subset,]
 #' run_sceptre_gRNA_gene_pair(expressions, gRNA_indicators, covariate_matrix)
 #'
-run_sceptre_gRNA_gene_pair <- function(expressions, gRNA_indicators, covariate_matrix, gene_precomp_size = NULL, B = 500, seed = NULL) {
+run_sceptre_gRNA_gene_pair <- function(expressions, gRNA_indicators, covariate_matrix, gene_precomp_size = NULL, B = 500, seed = NULL, reduced_output = TRUE) {
   cat(paste0("Running gRNA precomputation.\n"))
   gRNA_precomp <- run_gRNA_precomputation(gRNA_indicators, covariate_matrix)
 
   cat(paste0("Running gene precomputation.\n"))
   gene_precomp <- run_gene_precomputation(expressions, covariate_matrix, gene_precomp_size)
 
-  out <- run_sceptre_using_precomp(expressions = expressions, gRNA_indicators = gRNA_indicators, gRNA_precomp = gRNA_precomp, gene_precomp_size = gene_precomp$gene_precomp_size, gene_precomp_offsets = gene_precomp$gene_precomp_offsets, B = B, seed = seed)
+  out <- run_sceptre_using_precomp(expressions = expressions, gRNA_indicators = gRNA_indicators, gRNA_precomp = gRNA_precomp, gene_precomp_size = gene_precomp$gene_precomp_size, gene_precomp_offsets = gene_precomp$gene_precomp_offsets, B = B, seed = seed, reduced_output = reduced_output)
   return(out)
 }
 

### 2020-12-15 sceptre_package/sceptre/R/sctransform_functions.R

@@ -1,3 +1,5 @@
+#### NOTE: These functions are borrowed from the sctransform package. Credit goes to the authors of sctransform, Christoph Hafemeister and Rahul Satija. We reworked the functions slightly to improve their scalability.
+
 #' Regularize thetas
 #'
 #' @param genes_log_gmean the log-transformed vector of the geometric mean of the gene expressions

### 2020-12-15 sceptre_package/sceptre/data-raw/generating_example_data.R

@@ -5,8 +5,8 @@ code_dir <- if (is.na(args[1])) "/Users/timbarry/Box/SCEPTRE/SCEPTRE/" else args
 source(paste0(code_dir, "/sceptre_paper/analysis_drivers/analysis_drivers_gasp/file_paths_to_dirs.R"))
 source(paste0(code_dir, "sceptre_paper/analysis_drivers/analysis_drivers_gasp/sceptre_function_args.R"))
 
-gene_id <- "ENSG00000140400"
-gRNA_id <- "chr15.2442_top_two"
+gene_id <- "ENSG00000135390"
+gRNA_id <- "chr12.1893_top_two"
 # expression data
 expressions <- cell_gene_expression_matrix[,which(ordered_gene_ids == gene_id)]
 gRNA_indics <- read_fst(path = gRNA_indicator_matrix_fp, gRNA_id) %>% pull() %>% as.integer()

### 2020-12-15 sceptre_package/sceptre/data-raw/generating_example_data.R
include vignette

@@ -5,8 +5,8 @@ code_dir <- if (is.na(args[1])) "/Users/timbarry/Box/SCEPTRE/SCEPTRE/" else args
 source(paste0(code_dir, "/sceptre_paper/analysis_drivers/analysis_drivers_gasp/file_paths_to_dirs.R"))
 source(paste0(code_dir, "sceptre_paper/analysis_drivers/analysis_drivers_gasp/sceptre_function_args.R"))
 
-gene_id <- "ENSG00000135390"
-gRNA_id <- "chr12.1893_top_two"
+gene_id <- "ENSG00000164713"
+gRNA_id <- "chr7.3255_top_two"
 # expression data
 expressions <- cell_gene_expression_matrix[,which(ordered_gene_ids == gene_id)]
 gRNA_indics <- read_fst(path = gRNA_indicator_matrix_fp, gRNA_id) %>% pull() %>% as.integer()

### 2020-12-16 sceptre_paper/utilities/get_package_versions.R
add more info to readme

@@ -1,0 +1,7 @@
+require(tidyverse)
+packages <- c("Seurat", "R.utils", "fst", "sn", "MASS", "VGAM", "tidyverse", "bigstatsr", "monocle", "sceptre", "katsevich2020", "furrr", "Seurat", "R.matlab", "R.utils", "fst", "sn", "MASS", "VGAM", "tidyverse", "bigstatsr", "openxlsx", "rhdf5", "sceptre", "katsevich2020", "ggpubr", "cowplot")
+packages <- packages %>% unique %>% sort
+for (pack in packages) {
+  library(pack, character.only = TRUE)
+}
+map(set_names(packages, packages), function(i) packageVersion(i) %>% as.character) %>% unlist()

### 2021-11-04 gb_final/manuscript.tex
final gb submission

\documentclass{article}
\usepackage{graphicx}	
\usepackage{amssymb, amsmath, bbm}
\usepackage{geometry}
\usepackage[colorlinks=true,citecolor=blue,linkcolor=blue]{hyperref}
\newcommand\independent{\perp \!\!\! \perp}
\usepackage{capt-of}
%\usepackage{ulem}
\bibliographystyle{naturemag}

\title{SCEPTRE improves calibration and sensitivity in single-cell CRISPR screen analysis}

\author{Timothy Barry$^{1}$, Xuran Wang$^1$, John A. Morris$^{2,3}$, Kathryn Roeder$^{1,4}$, and Eugene Katsevich$^{5,*}$}

\begin{document}

\maketitle

\addtocounter{page}{-1}
\thispagestyle{empty}

\begin{enumerate}
 \item Department of Statistics and Data Science, Carnegie Mellon University
 \item New York Genome Center, New York USA
 \item Department of Biology, New York University
 \item Computational Biology Department, Carnegie Mellon University
 \item Department of Statistics and Data Science, Wharton School, University of Pennsylvania \\
\textrm{*}Corresponding author
\end{enumerate}

\begin{abstract}
Single-cell CRISPR screens are a promising biotechnology for mapping regulatory elements to target genes at genome-wide scale. However, technical factors like sequencing depth impact not only expression measurement but also perturbation detection, creating a confounding effect. We demonstrate on two single-cell CRISPR screens how these challenges cause calibration issues. We propose SCEPTRE: analysis of single-cell perturbation screens via conditional resampling, which infers associations between perturbations and expression by resampling the former according to a working model for perturbation detection probability in each cell. SCEPTRE demonstrates excellent calibration and sensitivity on CRISPR screen data, yielding hundreds of new regulatory relationships supported by orthogonal biological evidence.
\end{abstract}

\clearpage
\section*{Background}
The noncoding genome plays a crucial role in human development and homeostasis: over 90\% of loci implicated by GWAS in diseases lie in regions outside protein-coding exons\cite{Gallagher2018}. Enhancers and silencers, segments of DNA that modulate the expression of a gene or genes in \textit{cis}, harbor many or most of these noncoding trait loci. While millions of \textit{cis}-regulatory elements (CREs) have been nominated through biochemical annotations, the functional role of these CREs, including the genes that they target, remain essentially unknown\cite{Gasperini2020}. A central challenge over the coming decade, therefore, is to unravel the \textit{cis}-regulatory landscape of the genome across various cell types and diseases.

Single-cell CRISPR screens (implemented by Perturb-seq\cite{Dixit2016, Adamson2016}, CROP-seq\cite{Datlinger2017}, ECCITE-seq\cite{Mimitou2019}, and other protocols) are the most promising technology for mapping CREs to their target genes at genome-wide scale. Single-cell CRISPR screens pair CRISPR perturbations with single-cell sequencing to survey the effects of perturbations on cellular phenotypes, including the transcriptome. High multiplicity of infection (MOI) screens deliver dozens perturbations to each cell\cite{Xie2017, Xie2019, Gasperini2019}, enabling the interrogation of hundreds or thousands of CREs in a single experiment. Single-cell screens overcome the limitations of previous technologies for mapping CREs\cite{Gasperini2019}: unlike eQTLs, single-cell screens are high-resolution and can target rare variants, and unlike bulk screens, single-cell screens measure the impact of perturbations on the entire transcriptome.

Despite their promise, high-MOI single cell CRISPR screens pose significant statistical challenges. In particular, researchers have encountered substantial difficulties in calibrating tests of association between a CRISPR perturbation and the expression of a gene. Gasperini et al.\cite{Gasperini2019} found considerable inflation in their negative binomial regression based $p$-values for negative control perturbations. Similarly, Xie et al.\cite{Xie2019} found an excess of false positive hits in their rank-based Virtual FACS analysis. Finally, Yang et al.\cite{Yang2020} found that their permutation-based scMAGeCK-RRA method deems almost all gene-enhancer pairs significant in a reanalysis of the Gasperini et al.\ data. These works propose ad hoc fixes to improve calibration, but we argue that these adjustments are insufficient to address the issue. Miscalibrated $p$-values can adversely impact the reliability of data analysis conclusions by creating excesses of false positive and false negative discoveries.

In this work we make two contributions. We (i) elucidate core statistical challenges at play in high-MOI single-cell CRISPR screens and (ii) present a novel analysis methodology to address them. We identify a key challenge that sets single-cell CRISPR screens apart from traditional differential expression experiments: the ``treatment"---in this case the presence of a CRISPR perturbation in a given cell---is subject to measurement error\cite{Dixit2016,Hill2018,Replogle2020}. In fact, underlying this measurement error are the same technical factors contributing to errors in the measurement of gene expression, including sequencing depth and batch effects. These technical factors therefore act as confounders, invalidating traditional nonparametric calibration approaches. On the other hand, parametric modeling of single-cell expression data is also fraught with unresolved difficulties.

To address these challenges, we propose SCEPTRE (analysis of Single-CEll PerTurbation screens via conditional REsampling; pronounced ``scepter''). SCEPTRE is based on the conditional randomization test\cite{CetL16}, a powerful and intuitive statistical methodology that, like parametric methods, enables simple confounder adjusment, and like nonparametric methods, is robust to expression model misspecification. We used SCEPTRE to analyze two recent, large-scale, high-MOI single-cell CRISPR screen experiments. SCEPTRE demonstrated excellent calibration and sensitivity on the data and revealed hundreds of new regulatory relationships, validated using a variety of orthogonal functional assays. In the Discussion we describe an independent work conducted in parallel to the current study in which we leveraged biobank-scale GWAS data, single-cell CRISPR screens, and SCEPTRE to dissect the \textit{cis} and \textit{trans} effects of noncoding variants associated with blood diseases\cite{Morris2021}. This work highlights what we see as a primary application of SCEPTRE: dissecting regulatory mechanisms underlying GWAS associations.

\section*{Results}

\paragraph{Analysis challenges.}

We examined two recent single-cell CRISPR screen datasets -- one produced by Gasperini et al.\cite{Gasperini2019} and the other by Xie et al.\cite{Xie2019} -- that exemplify several of the analysis challenges in high-MOI single-cell CRISPR screens. Gasperini et al.\ and Xie et al.\ used CRISPRi to perturb putative enhancers at high MOI in K562 cells. They sequenced polyadenylated gRNAs alongside the whole transcriptome and assigned perturbation identities to cells by thresholding the resulting gRNA UMI counts.

Both Gasperini et al.\ and Xie et al.\ encountered substantial difficulties in calibrating tests of association between candidate enhancers and genes. Gasperini et al.\ computed $p$-values using a DESeq2\cite{Love2014}-inspired negative binomial regression analysis implemented in Monocle2\cite{Qiu2017}, and Xie et al.\ computed $p$-values using Virtual FACS, a nonparametric method proposed by these authors. Gasperini et al.\ assessed calibration by pairing each of 50 non-targeting (or negative) control gRNAs with each protein-coding gene. These ``null'' $p$-values exhibited inflation, deviating substantially from the expected uniform distribution (Figure \ref{fig:analysis-challenges}a, red). To assess the calibration of Virtual FACS in a similar manner, we constructed a set of \textit{in silico} negative control pairs of genes and gRNAs on the Xie et al.\ data (see Methods). The resulting $p$-values were likewise miscalibrated, with some pairs exhibiting strong conservative bias and others strong liberal bias (Figure \ref{fig:analysis-challenges}a, gray-green).

A core challenge in the analysis of single-cell CRISPR screens is the presence of confounders, technical factors that impact both gRNA detection probability and gene expression. The total number of gRNAs detected in a cell increases with the total number of mRNA UMIs detected in a cell ($\rho = 0.35, p < 10^{-15}$ in Gasperini et al.\ data; $\rho = 0.25, p < 10^{-15}$ in Xie et al.\ data; Figures \ref{fig:analysis-challenges}b-c). Techical covariates, such as sequencing depth and batch, induce a correlation between gRNA detection probability and gene expression, even in the absence of a regulatory relationship (Figure \ref{fig:analysis-challenges}d). This confounding effect can lead to severe test miscalibration and is especially problematic for traditional nonparametric approaches, which implicitly (and incorrectly) treat cells symmetrically with respect to confounders.

Parametric regression approaches, like negative binomial regression, are the most straightforward way to adjust for confounders. However, parametric methods rely heavily on correct model specification, a challenge in single-cell analysis given the heterogeneity and complexity of the count data. We hypothesized that inaccurate estimation of the negative binomial dispersion parameter was (in part) responsible for the $p$-value inflation observed by Gasperini et al. Monocle2 estimates a raw dispersion for each gene, fits a parametric mean-dispersion relationship across genes, and finally collapses raw dispersion estimates onto this fitted line (Figure \ref{fig:analysis-challenges}e). We computed the deviation from uniformity of the negative control $p$-values for each gene using the Kolmogorov-Smirnov (KS) test, represented by the color of each point in Figure \ref{fig:analysis-challenges}e. Circled genes had significantly miscalibrated $p$-values based on a Bonferroni correction at level $\alpha = 0.05$. Genes significantly above the curve showed marked signs of $p$-value inflation, suggesting model misspecification. Analysis challenges are summarized in Table \ref{summary_tab}.

Gasperini et al.\ and Xie et al.\ incorporated ad hoc adjustments into their analyses to remedy the observed calibration issues. On closer inspection, however, these efforts were not satisfactory to ensure reliability of the results. Gasperini et al.\ attempted to calibrate $p$-values against the distribution negative control $p$-values instead of the more standard uniform distribution. This adjustment lead to overcorrection for some gene-enhancer pairs (false negatives) and undercorrection for others (false positives) (Figure \ref{fig:gasp_correction}).  Along similar lines Xie et al.\ compared their Virtual FACS $p$-values to gene-specific simulated null $p$-values to produce ``significance scores'' that were used to determine significance. These significance scores were challenging to interpret and could not be subjected to multiple hypothesis testing correction procedures, as they are not $p$-values.

\begin{table}
	\begin{tabular}{ | p{6em} p{7em} p{10em} p{8em} | } 
		\hline
		\textbf{Method class} & \textbf{Example}  & \textbf{Robust to expression model misspecification} & \textbf{Able to adjust for confounders} \\ 
		\hline
		Parametric & Monocle\cite{Qiu2017} & No & Yes \\ 
		\hline
		Nonparametric & Virtual FACS\cite{Xie2019} & Yes & No \\ 
		\hline
		Conditional resampling & SCEPTRE & Yes & Yes \\
		\hline
	\end{tabular}
	\caption{\label{summary_tab} Statistical methods employed in single-cell CRISPR screen analysis. Parametric methods are non-robust to misspecified gene expression distributions, and classical nonparametric methods cannot adjust for confounders. Conditional resampling (implemented in this work as SCEPTRE) addresses both challenges.}
\end{table}

\paragraph{Improvements to the negative binomial approach.}

We attempted to alleviate the miscalibration within the negative binomial regression framework by following the recommendations of Hafemeister and Satija, who recently proposed a strategy for parametric modeling of single-cell RNA-seq data\cite{Hafemeister2019}. First, we abandoned the DESeq2-style size factors of Monocle2 and instead corrected for sequencing depth by including it as a covariate in the negative binomial regression model. Second, we adopted a more flexible dispersion estimation procedure: we (i) computed raw dispersion estimates for each gene, (ii) regressed the raw dispersion estimates onto the mean gene expressions via kernel regression, and (iii) projected the raw dispersion estimates onto the fitted nonparametric regression curve.

We reanalyzed the Gasperini et al.\ and Xie et al.\ negative control data using the improved negative binomial regression approach. In addition to sequencing depth, we included as covariates in the regression model the total number of expressed genes per cell and the technical factors accounted for in the original analysis (total number of gRNAs detected per cell, percentage of transcripts mapped to mitochondrial genes, and sequencing batch). Improved negative binomial regression exhibited better calibration than Monocle regression on both Gasperini et al.\ and Xie et al.\ datasets. Still, improved negative binomial regression demonstrated clear $p$-value inflation. We concluded that parametric count models likely are challenging to calibrate to high-MOI single-cell CRISPR screen data.

\paragraph{SCEPTRE: Analysis of single-cell perturbation screens via conditional resampling.}

To address the challenges identified above, we propose SCEPTRE, a methodology for single-cell CRISPR screen analysis based on the simple and powerful conditional randomization test\cite{CetL16} (Figure \ref{fig:method}). To test the association between a given gRNA and gene, we first fit the improved negative binomial statistic described above. This yields a $z$-value, which typically would be compared to a standard normal null distribution based on the parametric negative binomial model. Instead, we build a null distribution for this statistic via conditional resampling. First, we estimate the probability that the gRNA will be detected in a given cell based on the cell's technical factors, such as sequencing depth and batch. Next, we resample a large number of  ``null'' datasets, holding gene expression and technical factors constant while redrawing gRNA assignment independently for each cell based on its fitted probability. We compute a negative binomial $z$-value for each resampled dataset, resulting in an empirical null distribution (gray histogram in Figure~\ref{fig:method}). Finally, we compute a left-, right-, or two-tailed probability of the original $z$-value under the empirical null distribution, yielding a well-calibrated $p$-value. This $p$-value can deviate substantially from that obtained based on the standard normal (Figure~\ref{fig:method}, Figure~\ref{fig:NB-vs-CP}). While we used a negative binomial regression test statistic for this work, SCEPTRE in principle is compatible with any test statistic that reasonably tracks the expression data, including, for example, statistics based on machine learning algorithms.

We leverage several computational accelerations to enable SCEPTRE to scale to large single-cell CRISPR screen datasets. First, we approximate the null histogram of the resampled test statistics using a skew-$t$ distribution to obtain precise $p$-values based on a limited number of resamples (500 in the current implementation). Second, we employ statistical shortcuts that reduce the cost of each resample by a factor of about 100 (see Methods). Finally, we implement the method so that it can run in parallel on hundreds or thousands of processors on a computer cluster. (We used this approach in our independent study of noncoding blood trait GWAS loci\cite{Morris2021}.) We estimate that SCEPTRE can analyze 2.5 million gene-gRNA pairs on a dataset of 200,000 cells in a single day using 500 processors.

\paragraph{SCEPTRE demonstrates good calibration and sensitivity on real and simulated data.}

First, we investigated the calibration of SCEPTRE in a small, proof-of-concept simulation study (Figure \ref{fig:ground_truth}a). We considered a class of negative binomial regression models with fixed dispersion and two technical covariates (sequencing depth and batch). We simulated expression data for a single gene in 1000 cells using four models selected from this class: the first with  dispersion $= 1$, the second with dispersion $= 0.2$, the third with dispersion $= 5$, and the last with dispersion $= 1,$ but with 25$\%$ zero-inflation. We also simulated negative control gRNA data using a logistic regression model with the same covariates as the gene expression model. We assessed the calibration of SCEPTRE and negative binomial regression across the four simulated datasets. To explore the impact of model misspecification on SCEPTRE and the negative binomial method (on which SCEPTRE relies), we fixed the dispersion of the negative binomial method to 1. The negative binomial method worked as expected when the model was correctly specified. However, negative binomial regression broke down in all three cases of model misspecification. SCEPTRE demonstrated good calibration in all settings.

Next, to assess the calibration of SCEPTRE on real data, we applied SCEPTRE to test the association between negative control gRNAs and genes in the Gasperini et al.\ data (Figure \ref{fig:ground_truth}b) and Xie et al.\ data (Figure \ref{fig:ground_truth}c). We compared SCEPTRE to Monocle regression and the improved negative binomial method. For the Xie et al.\ data, we also compared to Virtual FACS, the method originally applied to the data. SCEPTRE showed good calibration on both datasets; by contrast, Monocle regression and improved negative binomial regression demonstrated signs of severe $p$-value inflation, while Virtual FACS exhibited a bimodal $p$-value distribution peaked at 0 and 1.

SCEPTRE demonstrated modestly better calibration on the Gasperini et al.\ data than on the Xie et al.\ data. This likely is because the Gasperini et al.\ negative control pairs -- which consisted of real, non-targeting gRNAs -- were higher-quality than the Xie et al.\ negative control pairs  -- which were constructed \textit{in-silico} using enhancer-targeting gRNAs (see Methods). We reasoned that the Xie et al.\ negative controls carried mild regulatory signal, resulting in slight inflation of the SCEPTRE $p$-values on these data relative to the Gasperini et al.\ data.

To assess the sensitivity of SCEPTRE, we applied SCEPTRE to test the 381 positive control pairs of genes and TSS-targeting gRNAs assayed by Gasperini et al.\ (Figure~\ref{fig:ground_truth}d). Allowing for the fact that the empirical correction employed by Gasperini et al.\ limited the accuracy of $p$-values to about $10^{-6}$, the SCEPTRE $p$-values for the positive controls were highly significant, and in particular, almost always more significant than the original empirical $p$-values, indicating greater sensitivity. Finally, we assessed the sensitivity of SCEPTRE on the Xie et al.\ data. Xie et al.\ conducted an arrayed CRISPR screen with bulk RNA-seq readout of ARL15-enh, a putative enhancer of gene \textit{ARL15}. Both SCEPTRE and the bulk RNA-seq differential expression analysis rejected \textit{ARL15} at an FDR of 0.1 after a Benjamini-Hochberg (BH) correction, increasing our confidence in the calibration and sensitivity of SCEPTRE (Figure \ref{fig:ground_truth}e).

\paragraph{Analysis of candidate \textit{cis}-regulatory pairs.}

We applied SCEPTRE to test all candidate \textit{cis}-regulatory pairs in the Gasperini et al.\ ($n=84,595$) and Xie et al.\ ($n=5,209$) data. A given gene and gRNA were considered a ``candidate pair'' if the gRNA targeted a site within one Mb the gene's TSS. SCEPTRE discovered $563$ and 139 gene-enhancer links at an FDR of 0.1 on the Gasperini et al.\ and Xie et al.\ data, respectively. We used several orthogonal assays to quantify the enrichment of SCEPTRE's discovery set for regulatory biological signals, and we compared the SCEPTRE results to those of other methods.

SCEPTRE's discovery set on the Gasperini et al.\ data was highly biologically plausible, and in particular, more enriched for biological signals of regulation than the original discovery set. Gasperini et al.\ discovered 470 gene-gRNA pairs at a reported FDR of $0.1$. The SCEPTRE $p$-values and original empirical $p$-values diverged substantially: of the 670 gene-enhancer pairs discovered by either method, SCEPTRE and the original method agreed on only 363, or 54\% (Figure~\ref{fig:application}a). Gene-enhancer pairs discovered by SCEPTRE were physically closer (mean distance $=65$ kb) to each other than those discovered by the original method (mean distance $=81$ kb; Figure \ref{fig:application}b). Furthermore, SCEPTRE's gene-enhancer pairs fell within the same topologically associating domain (TAD) at a higher frequency ($74\%$) than the original pairs ($71\%$). Pairs within the same TAD showed similar levels of HI-C interaction frequency across methods, despite the fact that SCEPTRE discovered 85 more same-TAD pairs (Figure \ref{fig:application}c). Finally, enhancers discovered by SCEPTRE showed improved enrichment across all eight cell-type relevant ChIP-seq targets reported by Gasperini et al.\ (Figure~\ref{fig:application}d, Figure~\ref{fig:chipseq-enrichment-details}a).

When we compared discoveries unique to SCEPTRE ($n = 200$) against those unique to the original method ($n = 107$), the disparities became more extreme (Figure \ref{fig:symmetric-difference-gasp}). For example, only 57\% of pairs unique to the original method fell within the same TAD, compared to 73\% unique to SCEPTRE. We concluded that many pairs in the Gasperini et al.\ discovery set likely were false positives. Finally, when we compared SCEPTRE to the improved negative binomial method ($n$ = 824 discoveries), we observed even greater differences in discovery set quality in favor of SCEPTRE (Figures \ref{fig:application}b - \ref{fig:application}d).

We highlight several especially interesting gene-enhancer pairs discovered by SCEPTRE. Five discoveries (Figure \ref{fig:application}a labels 1-5; Figure \ref{fig:application}e) were nominated as probable gene-enhancer links by eQTL\cite{GTExPilot} and eRNA\cite{Andersson2014} $p$-values in relevant tissue types. The SCEPTRE $p$-values for these pairs were 1-2 orders of magnitude smaller than the original empirical $p$-values, hinting at SCEPTRE's greater sensitivity. Additionally, six pairs (Figure~\ref{fig:application}a, blue triangles) were discovered by SCEPTRE but discarded as outliers by the original analysis, underscoring SCEPTRE's ability to handle genes with arbitrary expression distributions.

We repeated the same orthogonal analyses for the SCEPTRE discoveries on the Xie et al.\ data, comparing SCEPTRE's results to those of Xie et al. Xie et al.'s analysis method, Virtual FACS, outputted ``significance scores'' rather than $p$-values (see section ``Analysis challenges''). Because significance scores cannot be subjected to multiple hypothesis testing correction procedures (like BH), we compared the top 139 Virtual FACS pairs (ranked by significance score) against the set of 139 (FDR $= 0.1$) SCEPTRE discoveries (Figure \ref{fig:application-xie}a; see Methods). Of the 195 pairs in either set, SCEPTRE and Virtual FACS agreed on only 83, or 43\%. The SCEPTRE discoveries were more biologically plausible: compared to the Virtual FACS pairs, the SCEPTRE pairs were (i) physically closer (Figure \ref{fig:application-xie}b), (ii) more likely to fall within the same TAD (Figure \ref{fig:application-xie}c), (iii) more likely to interact when in the same TAD (Figure \ref{fig:application-xie}c), and (iv) more enriched for all eight cell-type relevant ChIP-seq targets (Figure~\ref{fig:application-xie}d, Figure~\ref{fig:chipseq-enrichment-details}b). When we examined the symmetric difference of the discovery sets, these differences became more pronounced (Figure \ref{fig:symmetric-difference-xie}).

We additionally compared SCEPTRE to Monocle regression ($n = 180$ discoveries) and improved negative binomial regression ($n = 156$ discoveries) on the Xie et al.\ data. SCEPTRE uniformly dominated Monocle: SCEPTRE pairs were physically closer to one another (median distance = $44$kb versus $110$kb; Figure \ref{fig:application-xie}b); SCEPTRE pairs interacted more frequently and were more likely to fall within the same TAD (68\% versus 61\%; Figure \ref{fig:application-xie}c); and SCEPTRE pairs were more enriched for seven of eight cell type-relevant ChIP-seq targets (one target, DP22, was a tie; Figure \ref{fig:application-xie}d). Improved negative binomial regression was more competitive than Monocle across metrics (Figure \ref{fig:application-xie}b-d). However, as noted earlier, improved negative binomial regression exhibited severe miscalibration on the negative control pairs (Figure \ref{fig:ground_truth}d), rendering its discovery set less reliable than that of SCEPTRE.

\paragraph{Gene expression level and sensitivity.}
To better understand when SCEPTRE works best, we investigated the impact of gene expression level on the sensitivity of SCEPTRE. We binned candidate gene-enhancer pairs into non-overlapping categories based on mean expression level of the gene. On both the Gasperini et al.\ and Xie et al.\ data, we found that candidate pairs containing a highly-expressed gene were more likely to be rejected than candidate pairs containing a lowly-expressed gene (Table \ref{binning_tab_gasp},\ref{binning_tab_xie}), indicating SCEPTRE's greater sensitivity for highly expressed genes. We observed similar trends for the other methods (not shown), consistent with the intuition that highly-expressed genes carry more information.

\section*{Discussion}

In this work we illustrated a variety of statistical challenges arising in the analysis of high-MOI single-cell CRISPR screens, leaving existing methods (based on parametric expression models, permutations, or negative control data) vulnerable to miscalibration. To address these challenges, we developed SCEPTRE, a resampling method based on modeling the probability a given gRNA will be detected in a given cell, based on that cell's technical factors. We found that SCEPTRE exhibited very good calibration despite the presence of confounding technical factors and misspecification of single-cell gene expression models. We implemented computational accelerations to bring the cost of the resampling-based methodology down to well within an order of magnitude of the traditional negative binomial parametric approach, making it quite feasible to apply for large-scale data. We used SCEPTRE to reanalyze the Gasperini et al.\ and Xie et al.\ data.  While our analysis replicated many of their findings, it also clarified other relationships, removing a large set ($>20\%$ for Gasperini) of pairs that exhibited a weak relationship and adding an even larger set ($>40\%$ for Gasperini) of new, biologically plausible gene-enhancer relationships. These links were supported by orthogonal evidence from eQTL, enhancer RNA co-expression, ChIP-seq, and HI-C data.

As an example application of SCEPTRE, we highlight \textit{STING-seq}, a platform that we developed in parallel to the current work in an independent study\cite{Morris2021}. STING-seq leverages biobank-scale GWAS data and single-cell CRISPR screens to map noncoding, disease-associated variants at scale. First, we used statistical fine-mapping to identify a set of 88 putatively causal variants from 56 loci associated with quantitative blood traits. We perturbed the selected variants at high MOI in K562 cells using an improved CRISPRi platform and sequenced gRNAs and transcriptomes in individual cells using ECCITE-seq\cite{Mimitou2019}, a protocol that enables the profiling of multiple modalities and the direct capture of gRNAs.

We then applied SCEPTRE to quantify associations between perturbations and changes in gene expression in \textit{cis} (within 500 kb) and \textit{trans}. SCEPTRE confidently mapped 37 noncoding variants to their \textit{cis} target genes, in some cases identifying a causal variant among a set of candidate variants in strong LD. Nine variants were found to regulate a gene other than the closest gene, and four variants were found to regulate multiple genes, an apparent example of pleiotropy. Several perturbations lead to widespread changes in gene expression, illuminating \textit{trans}-effects networks. For example, two variants that were found to regulate the transcription factor \textit{GFI1B} in \textit{cis} altered the expression of hundreds of genes in \textit{trans} upon perturbation; these differentially expressed genes were strongly enriched for GFI1B binding sites and blood disease GWAS hits. We concluded on the basis of this study SCEPTRE can power the systematic dissection regulatory networks underlying GWAS associations.

Despite these exciting results, key challenges remain in the analysis of single-cell CRISPR screens. Currently, SCEPTRE does not estimate the effect sizes of perturbations, disentangle interactions among perturbed regulatory elements\cite{Zamanighomi2019,Norman2019}, or leverage information from off-targeting gRNAs to improve power. Such extensions could be implemented by harnessing more sophisticated, multivariate models of gRNA detection or applying methods for estimating variable importance in the presence of possibly misspecified models\cite{Zhang2020}. The statistical challenges that we identified in this study -- specifying an accurate expression model, accounting for technical factors -- and the solutions that we proposed -- conditional resampling, massively parallel association testing -- will help guide the development of future versions of SCEPTRE.

\section*{Conclusions}

Single-cell CRISPR screens will play a key role in unraveling the regulatory architecture of the noncoding genome\cite{Przybyla2021}. Technological improvements and methodological innovations will increase the scope, scale, and variety of theses screens over the coming years. For example, screens of candidate CREs could be extended to different, disease-relevant cell types and tissues (although this remains a challenge); new combinatorial indexing strategies, such as scifi-RNA-seq, could enable the scaling-up of such screens to millions of cells\cite{Datlinger2021}; different CRISPR technologies, such as CRISPRa, could enable the activation, rather than repression, of candidate CREs, yielding new insights; and information-rich, multimodal single-cell readouts could strengthen conclusions drawn about regulatory relationships\cite{Pierce2020}. SCEPTRE is a flexible, robust, and efficient method: it has now successfully been applied to three single-cell CRISPR screen datasets, across two technologies (CROP-seq and ECCITE-seq), to map regulatory relationships both in \textit{cis} and in \textit{trans}. We expect SCEPTRE to facilitate the analysis of future single-cell screens of the noncoding genome, advancing understading of CREs and enabling the detailed interpretation of GWAS results.

\section*{Methods}

\paragraph{Gasperini et al.\ and Xie et al.\ data.}
Gasperini et al.\ used CROP-seq\cite{Datlinger2017,Hill2018} to transduce a library of CRISPR guide RNAs (gRNAs) into a population of 207,324 K562 cells expressing the Cas9-KRAB repressive complex at a high multiplicity of infection. Each cell received an average of 28 perturbations. The gRNA library targeted 5,779 candidate enhancers, 50 negative controls, and 381 TSS-targeting positive controls. Xie et al.\ used Mosaic-seq\cite{Xie2019,Xie2017} to perturb at a high multiplicity of infection 518 putative enhancers in a population of 106,670 Cas9-KRAB-expressing K562 cells. Each putative enhancer was perturbed in an average of 1,276 cells.
\paragraph{\textit{Cis} and \textit{in silico} negative control pairs for Xie et al.\ data.} We generated the set of candidate \textit{cis} gene-enhancer relationships on the Xie et al.\ data by pairing each protein-coding gene with each gRNA targeting a site within 1 Mb of the TSS of the gene. This procedure resulted in 3,553 candidate \textit{cis} gene-enhancer links that we tested using SCEPTRE and Virtual FACS.

To generate the set of \textit{in silico} negative control pairs for calibration assessment, we (i) identified gRNAs that targeted sites far ($>1$ Mb) from the TSSs of known transcription factor genes and (ii) paired these gRNAs with genes located on other chromosomes. We excluded all pairs containing genes known to be transcription factors, and we downsampled the pairs so that each gRNA was matched to 500 genes. The final \textit{in silico} negative control set consisted of 84,500 pairs, the elements of which were not expected to exhibit a regulatory relationship.

\paragraph{Conditional randomization test.}
Consider a particular gene/gRNA pair. For each cell $i = 1, \dots, n$, let $X_i \in \{0,1\}$ indicate whether the gRNA was present in the cell, let $Y_i \in \{0,1,2,\dots\}$ be the gene expression in the cell, defined as the number of unique molecular identifiers (UMIs) from this gene, and let $Z_i \in \mathbb R^d$ be a list of cell-level technical factors. Letting $(X, Y, Z) = \{(X_i, Y_i, Z_i)\}_{i = 1}^{n}$, consider any test statistic $T(X,Y,Z)$ measuring the effect of the gRNA on the expression of the gene. The conditional randomization test\cite{CetL16} is based on resampling the gRNA indicators independently for each cell. Letting $\pi_i = \mathbb P[X_i = 1 | Z_i]$, define random variables
\begin{equation}
\widetilde X_i \overset{\text{ind}}\sim \text{Ber}(\pi_i). 
\label{X-tilde-dist}
\end{equation}
Then, the CRT $p$-value is given by
\begin{equation}
p_{\text{CRT}} = \mathbb P[T(\widetilde X, Y, Z) \geq T(X,Y,Z)\mid X,Y,Z].
\label{conditional-randomization}
\end{equation}
This translates to repeatedly sampling $\widetilde X$ from the distribution~\eqref{X-tilde-dist}, recomputing the test statistic with $X$ replaced by $\widetilde X$, and defining the $p$-value as the probability the resampled test statistic exceeds the original. Under the null hypothesis that the gRNA perturbation does not impact the cell (adjusting for technical factors), i.e. $Y \independent X \mid Z$, we obtain a valid $p$-value~\eqref{conditional-randomization}, \textit{regardless of the expression distribution $Y|X,Z$ and regardless of the test statistic $T$}. We choose as a test statistic $T$ the $z$-score of $X_i$ obtained from a negative binomial regression of $Y_i$ on $X_i$ and $Z_i$:

\begin{equation}
Y_i \overset{\text{ind}} \sim \text{NegBin}(\mu_i, \alpha); \quad \log(\mu_i) = \beta_0 + X_i \beta + Z_i^T \gamma
\label{negative-binomial},
\end{equation}
where $\alpha$ is the dispersion. Following Hafemeister and Satjia\cite{Hafemeister2019}, we estimate $\alpha$ by pooling dispersion information across genes, and we include sequencing depth as an entry in the vector of technical factors  $Z_i$ (see section \textit{Improvements to the negative binomial approach}).

\paragraph{Accelerations to the conditional randomization test.}
We implemented computational accelerations to the conditional randomization test. First, we employed the recently proposed\cite{Liu2020} \textit{distillation} technique to accelerate the recomputation of the negative binomial regression for each resample. The idea is to use a slightly modified test statistic, consisting of two steps:
\begin{enumerate}
	\item Fit $(\widehat \beta_0, \widehat \gamma)$ from the negative binomial regression~\eqref{negative-binomial} except without the gRNA term:
	\begin{equation}
	Y_i \overset{\text{ind}}\sim \textrm{NegBin}(\mu_i, \alpha); \quad \log(\mu_i) = \beta_0 + Z_i^T \gamma.
	\label{negative-binomial-distillation}
	\end{equation}
	\item Fit $\widehat \beta$ from a negative binomial regression with the estimated contributions of $Z_i$ from step 1 as offsets:
	\begin{equation}
	Y_i \overset{\text{ind}}\sim \textrm{NegBin}(\mu_i, \alpha); \quad \log(\mu_i) = X_i \beta + \widehat \beta_0 + Z_i^T \widehat \gamma.
	\label{distilled-negative-binomial}
	\end{equation}
\end{enumerate}
Conditional randomization testing with this two step test statistic, which is nearly identical to the full negative binomial regression~\eqref{negative-binomial}, is much faster. Indeed, since the first step is not a function of $X_i$, it remains the same for each resampled triple $(\widetilde X, Y, Z)$. Therefore, only the second step must be recomputed with each resample, and this step is faster because it involves only a univariate regression.

Next, we accelerated the second step above using the sparsity of the binary vector $(X_1, \dots, X_n)$ (or a resample of it). To do so, we wrote the log-likelihood of the reduced negative binomial regression~\eqref{distilled-negative-binomial} as follows, denoting by $\ell(Y_i, \log(\mu_i))$ the negative binomial log-likelihood:
\begin{equation*}
\begin{split}
\sum_{i = 1}^n \ell(Y_i, X_i \beta + \widehat \beta_0 + Z_i^T \widehat \gamma) &= \sum_{i: X_i = 0} \ell(Y_i, \widehat \beta_0 + Z_i^T \widehat \gamma) + \sum_{i: X_i = 1} \ell(Y_i, \beta + \widehat \beta_0 + Z_i^T \widehat \gamma) \\
&= C + \sum_{i: X_i = 1} \ell(Y_i, \beta + \widehat \beta_0 + Z_i^T \widehat \gamma).
\end{split}
\end{equation*}
This simple calculation shows that, up to a constant that does not depend on $\beta$, the negative binomial log-likelihood corresponding to the model~\eqref{distilled-negative-binomial} is the same as that corresponding to the model with only intercept and offset term for those cells with a gRNA:
\begin{equation}
Y_i \overset{\text{ind}}\sim \textrm{NegBin}(\mu_i, \alpha); \quad \log(\mu_i) = \beta + \widehat \beta_0 + Z_i^T \widehat \gamma, \quad \text{for } i \text{ such that } X_i = 1.
\label{distilled-negative-binomial-accelerated}
\end{equation}
The above negative binomial regression is therefore equivalent to equation~\eqref{distilled-negative-binomial}, but much faster to compute, because it involves much fewer cells. For example, in the Gasperini et al. data, each gRNA is observed in only about 1000 of the 200,000 total cells.

\paragraph{SCEPTRE methodology.}

In practice, we must estimate the gRNA probabilities $\pi_i$ as well as the $p$-value $p_{\text{CRT}}$. This is because usually we do not know the distribution $X|Z$ and cannot compute the conditional probability in equation~\eqref{conditional-randomization} exactly. We propose to estimate $\pi_i$ via logistic regression of $X$ on $Z$, and to estimate $p_{\text{CRT}}$ by resampling $\widetilde X$ a large number of times and then fitting a skew-$t$ distribution to the resampling null distribution $T(\widetilde X, Y, Z)|X,Y,Z$. We outline SCEPTRE below: 
\begin{enumerate}
	\item Fit technical factor effects $(\widehat \beta_0,\widehat \gamma)$ on gene expression using the negative binomial regression~\eqref{negative-binomial-distillation}.
	\item Extract a $z$-score $z(X,Y,Z)$ from the reduced negative binomial regression~\eqref{distilled-negative-binomial-accelerated}.
	\item Assume that 
	\begin{equation}
	X_i \overset{\text{ind}}\sim \text{Ber}(\pi_i); \quad \log\left(\frac{\pi_i}{1-\pi_i}\right) = \tau_0 + Z_i^T \tau
	\label{logistic-regression}
	\end{equation}
	for $\tau_0 \in \mathbb R$ and $\tau \in \mathbb R^d$, and fit $(\widehat \tau_0, \widehat \tau)$ via logistic regression of $X$ on $Z$. Then, extract the fitted probabilities $\widehat \pi_i = (1+\exp(-(\widehat \tau_0 + Z_i^T \widehat \tau)))^{-1}$.	
	\item For $b = 1, \dots, B$,
	\begin{itemize}
		\item Resample the gRNA assignments based on the probabilities $\widehat \pi_i$ to obtain $\widetilde X^b$~\eqref{X-tilde-dist}.
		\item Extract a $z$-score $z(\widetilde X^b, Y, Z)$ from the reduced negative binomial regression~\eqref{distilled-negative-binomial-accelerated}.
	\end{itemize}
	\item Fit a skew-$t$ distribution $\widehat F_{\text{null}}$ to the resampled $z$-scores $\{z(\widetilde X^b, Y, Z)\}_{b = 1}^B$. 
	\item Return the $p$-value $\widehat p_{\text{SCEPTRE}} = \mathbb P[\widehat F_{\text{null}} \leq z(X,Y,Z)]$.
\end{enumerate}
In our data analysis we used $B = 500$ resamples.

\paragraph{Numerical simulation to assess calibration.}

We simulated one gene $Y_i,$ one gRNA $X_{i}$, and two confounders $Z_{i1}, Z_{i2}$ in $n = 1000$ cells. We generated the confounders $Z_{i1}$ and $Z_{i2}$ by sampling with replacement the batch IDs and log-transformed sequencing depths of the cells in the Gasperini dataset. The batch ID confounder $Z_{i1}$ was a binary variable, as the Gasperni data included two batches. Next, we drew the gRNA indicators $X_{i}$ i.i.d.\ from the logistic regression model (7), with $\tau_0 = -7, \tau_1 = -2,$ and $\tau_2 = 0.5$. We selected these parameters to make the probability of gRNA occurrence about 0.04 across cells. Finally, we drew the gene expression $Y_i$ from the following zero-inflated negative binomial model:
$$ Y_i \sim \lambda \delta_0 + (1 - \lambda) \textrm{NegBin}(\mu_i, \alpha), \textrm{ } \log(\mu_i) = \beta_0 + Z_i^T \beta.$$ Note that gRNA presence or absence does not impact gene expression in this model. 
We set $\beta_0 = -2.5, \beta_1 = -2, \beta_2 = 0.5$ to make the average gene expression about $4$ across cells. We generated the four datasets shown in Figure \ref{fig:ground_truth}a by setting the dispersion parameter $\alpha$ and the zero inflation rate parameter $\lambda$ equal to the following values:
$$ (\lambda_1, \alpha_1) = (0,1) ; \textrm{    } (\lambda_2, \alpha_2) = (0, 5); \textrm{     } (\lambda_3, \alpha_3) = (0, 0.2); \textrm{     } (\lambda_4, \alpha_4) = (0.25, 1).$$ For the first, the negative binomial model is correctly specified. For the second and third, the dispersion estimate of 1 is too small and too large, respectively. The last setting exhibits zero inflation. We applied SCEPTRE and negative binomial regression to the four problem settings, each with $n_{sim} = 500$ repetitions. The negative binomial method, and in turn SCEPTRE, was based on the $z$ statistic from the Hafemeister-inspired negative binomial model (\ref{negative-binomial}) with $\alpha = 1$. We used $B = 500$ resamples for SCEPTRE, the default choice.

\paragraph{scMAGeCK.} scMAGeCK-LR\cite{Yang2020} is a method for high MOI single-cell CRISPR screen analysis. (A complimentary method, scMAGeCK-RRA, is designed for the low-MOI setting) scMAGeCK-LR (henceforth scMAGeCK) uses a permutation test with ridge regression test statistic to compute $p$-values for pairs of genes and gRNAs. Unfortunately, we were unable to apply scMAGeCK to the real data. First, we were unable to understand the documentation of the scMAGeCK software well enough to confidently apply the method. Second, scMAGeCK is prohibitively expensive to apply at-scale. The authors of the original scMAGeCK study applied their method only to a small subset of pairs in the Gasperini et al.\ data. We could not meaningfully compare scMAGeCK to SCEPTRE on calibration and sensitivity metrics without applying scMAGeCK to the full set of gRNA-gene pairs, which, to our knowledge, never has been done (and likely is infeasible).

To enable a simple comparison to scMAGeCK on the simulated data, we implemented a custom, in-house version of scMAGeCK based on a careful examination of the scMAGeCK codebase and a close reading of the original paper. We view this custom implementation as a faithful interpretation of the method in the specialized one-gene to one-NTC setting. We applied our implementation of scMAGeCK to the simulated data, using $B = 1,000$ permutations, the default option. To reduce confusion, we reported the results of the scMAGeCK simulation study in the supplementary materials (Figure \ref{fig:scmageck}) rather than the Results section. We could not apply our custom implementation of scMAGeCK to the real data, because the real data are significantly more complex than the simulated data. For example, the real data consist of many genes and gRNAs, and the gRNAs are differently typed (e.g., negative control, positive control, enhancer-targeting, etc.), complicating the analysis considerably.

\paragraph{Definition of Gasperini et al.\ discovery set.}

Gasperini et al.\ reported a total of 664 gene-enhancer pairs, identifying 470 of these as ``high-confidence." We chose to use the latter set, rather than the former, for all our comparisons. Gasperini et al.\ carried out their ChIP-seq and HI-C enrichment analyses only on the high-confidence discoveries, so for those comparisons we did the same. Furthermore, the 664 total gene-enhancer pairs reported in the original analysis were the result of a BH FDR correction that included not only the candidate enhancers but also hundreds of positive controls. While Bonferroni corrections can only become more conservative when including more hypotheses, BH corrections are known to become anticonservative when extra positive controls are included\cite{FR01}. To avoid this extra risk of false positives, we chose to use the ``high-confidence" set throughout.

\paragraph{Xie et al.\ significance scores and  discovery set.} Xie et al.\ reported a local (or \textit{cis}) discovery set, which consisted of gene-gRNA pairs with a significance score of greater than zero (see original manuscript for definition of ``significance score''\cite{Xie2019}; cutoff of zero arbitrary). This discovery set was not directly comparable to the SCEPTRE discovery set. First, the candidate set of \textit{cis} gene-gRNA pairs tested by Xie et al.\ consisted of gRNAs within \textit{two} Mb of a protein-coding gene \textit{or} long-noncoding RNA. Our candidate \textit{cis} set, by contrast, consisted of gRNAs within \textit{one} Mb of a protein-coding gene. We defined our candidate \textit{cis} set differently than Xie et al.\ to mantain consistency with Gasperini et al. Second, Xie et al.\ appear to have used a significantly more conservative threshold than Gasperini et al.\ in defining their discovery set, but this was challenging to ascertain given the impossibility of FDR correction on the significance scores. To enable a meaningful comparison between Virtual FACS and SCEPTRE, we therefore ranked the Virtual FACS pairs by their significance score and selected the top $n$ pairs, where $n$ was the size of the SCEPTRE discovery set at FDR 0.1.

\paragraph{ChIP-seq, HI-C enrichment analyses.}

ChIP-seq and HI-C enrichment analyses analyses on the Gasperini et al.\ data (see Figures~\ref{fig:application}e-f and \ref{fig:chipseq-enrichment-details}a) were carried out almost exactly following Gasperini et al. The only change we made is in our quantification of the ChIP-seq enrichment (Figure~\ref{fig:application}f). We used the odds ratio of a candidate enhancer being paired to a gene, comparing the top and bottom ChIP-seq quintiles. On the Xie et al.\ data, we binned the candidate enhancers into two (rather than five) quantiles due to the fewer number of candidate \textit{cis} pairs. We computed odds ratios by comparing enhancers in the upper quantile to those that did not intersect a ChIP-seq peak at all (Figure \ref{fig:chipseq-enrichment-details}b).

\bibliography{library}

\section*{Figure legends}
\begin{enumerate}
\item \textbf{CRISPR screen analysis challenges can lead to false positives and false negatives}. \textbf{a}, QQ-plot of negative control $p$-values produced by Gasperini et al.\ (red; downsampled for visualization) and Xie et al.\ (gray-green). These $p$-values deviate substantially from the expected uniform distribution, indicating test miscalibration. \textbf{b-d}, Technical factors, such as sequencing depth and batch, impact gRNA detection probability and observed gene expression levels in both Gasperini et al.\ (b) and Xie et al.\ (c) data. Thus, technical factors act as confounders (d), differentiating CRISPR screens from traditional differential expression applications. \textbf{e}, Monocle2 estimates the dispersion of each gene by projecting each gene's raw dispersion estimate onto the fitted raw dispersion-mean expression curve. This estimation procedure leads to miscalibration for high-dispersion genes.
\item \textbf{SCEPTRE: Analysis of single-cell perturbation screens via conditional resampling.} A schematic and outline of the SCEPTRE methodology for one gene and one gRNA. SCEPTRE estimates the probability of gRNA detection in each cell based on its technical factors. It then builds a null distribution for the negative binomial $z$-value by independently resampling gRNA presence or absence for each cell according to these probabilities to form ``negative control" datasets. A skew-$t$ distribution is fit to the resulting histogram to obtain precise $p$-values based on a limited number of resamples, against which the original NB $z$-value is compared. The dashed line shows the standard normal distribution, against which the NB $z$-value typically would be compared.
\item \textbf{SCEPTRE demonstrates good calibration and sensitivity under known ground truth.} \textbf{a,} Numerical simulation comparing SCEPTRE and improved negative binomial regression on four simulated datasets. The negative binomial model was incorrectly specified on three of the four datasets. SCEPTRE maintained good calibration across all four datasets, despite model misspecification and confounder presence. \textbf{b-c}, Application of SCEPTRE, improved negative binomial regression, Monocle regression, and Virtual FACS to pairs of negative control gRNAs and genes in (b) the Gasperini et al.\ data, and (c) the Xie et al.\ data. Compared to the other methods, SCEPTRE showed good calibration. \textbf{d,} SCEPTRE $p$-values for Gasperini et al.\ TSS-targeting controls were highly significant, and in general, more significant than the original empirical $p$-values. \textbf{e,} Comparison of $p$-values produced by SCEPTRE for ARL15-enh to $p$-values produced by an arrayed, bulk RNA-seq CRISPR screen of ARL15-enh. The results of the two analyses coincided almost exactly, with both analyses rejecting gene \textit{ARL15} with high confidence after a BH correction. Dotted blue lines, rejection thresholds.
\item \textbf{Application of SCEPTRE to Gasperini et al.\ data yields biologically plausible gene-enhancer links.} \textbf{a,} Comparison of the original empirical $p$-values to those obtained from SCEPTRE. The two analysis methods differed substantially, with 200 gene-enhancer links discovered only by SCEPTRE and 107 discovered only by the original analysis. Annotations correspond to pairs in panel (e). \textbf{b,} Distribution of distances from TSS to upstream paired enhancers. Compared to Monocle NB (original) and improved NB analyses, SCEPTRE paired genes with nearer enhancers on average. \textbf{c,} For those gene-enhancer pairs falling in the same TAD, the cumulative distribution of the fractional rank of the HI-C interaction frequency compared to other distance-matched loci pairs within the same TAD. SCEPTRE showed similar enrichment despite finding 25\% more within-TAD pairs compared to the original analysis. Inset table shows gene-enhancer pairs falling in the same TAD. SCEPTRE found 93 more total pairs compared to the original analysis, and a higher percentage of pairs fell within the same TAD. \textbf{d,} Enrichment of ChIP-seq signal from seven cell-type relevant transcription factors and one histone mark (H3K27ac) among paired enhancers. SCEPTRE showed stronger enrichment across all ChIP-seq targets. \textbf{e}, Five gene-enhancer pairs discovered by SCEPTRE but not the original analysis, each supported by a whole blood GTEx eQTL or FANTOM enhancer RNA correlation $p$-value.
\item \textbf{SCEPTRE discovers biologically plausible gene-enhancer links on Xie et al.\ data.} \textbf{a,} Comparison of SCEPTRE $p$-values to Virtual FACS significance scores. Significant SCEPTRE $p$-values ($n = 139$) are colored in blue and purple, and the top 139 Virtual FACS pairs, as ranked by significance score, are colored in gray-green and purple. The two sets diverged substantially, with only 43\% of pairs in shared across sets. \textbf{b-d,} These panels are similar to the corresponding panels in Figure \ref{fig:application}. SCEPTRE pairs showed strong enrichment for biological signals associated with enhancer activity on (b) physical distance, (c) HIC interaction, and (d) ChIP-seq metrics relative to other methods.
\end{enumerate}

\section*{Acknowledgments}
 We are indebted to Molly Gasperini, Jacob Tome, and Andrew Hill for clarifying several aspects of their data analysis\cite{Gasperini2019} and to the Shendure lab for providing extensive feedback on an earlier draft of this paper. We thank Shiqi Xie for providing guidance on using the Xie et al.\ 2019 data and Wei Li for a helpful discussion on scMAGeCK. Finally, we thank Tom Norman, Atray Dixit, and Wesley Tansey for useful discussions on single-cell CRISPR screens. This work was supported, in part, by National Institute of Mental Health (NIMH) grant R01MH123184 as well as SFARI Grant 575547. Part of the data analysis used the Extreme Science and Engineering Discovery Environment (XSEDE)\cite{XSEDE}, which is supported by National Science Foundation grant number ACI-1548562. Specifically, it used the Bridges system\cite{Bridges}, which is supported by NSF award number ACI-1445606, at the Pittsburgh Supercomputing Center (PSC).

\section*{Competing Interests} The authors declare that they have no
competing financial interests.
\section*{Correspondence}
Correspondence and requests for materials
should be addressed to E.K. \\ (email: ekatsevi@wharton.upenn.edu).
\section*{Author contributions}
GK and KR identified the problem, conceived the research, and provided supervision. GK developed the method with input from TB and KR. TB and GK implemented the method with assistance from JM. TB, XW, and GK performed the analyses. JM provided guidance on interpretation and communication of results. TB and GK wrote the manuscript with input from all authors.
\section*{Availability of data and materials} Analysis results are available online at \\
\url{https://upenn.box.com/v/sceptre-files-v8}. All analysis was performed on publicly available data. The Gasperini et al.\ CRISPR screen data\cite{Gasperini2019} are available at \url{www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE120861}. The Xie et al.\ single-cell and bulk CRISPR screen data are available at \url{www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE129837}. The ChIP-seq data are taken from the ENCODE project\cite{Dunham2012} and are available at \url{www.encodeproject.org/}. The HI-C enrichment analysis is based on the data from Rao et al.\cite{Rao2014}, available at \url{www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE63525}. The eQTL and eRNA co-expression $p$-values are taken from the GeneHancer, database\cite{Fishilevich2017} available as part of GeneCards (\url{www.genecards.org/}).

The \texttt{sceptre} R package is available at \url{github.com/Katsevich-Lab/sceptre}. Vignettes and tutorials are available at \url{katsevich-lab.github.io/sceptre/}. The scripts used to run the analyses reported in this paper are available at \\ \url{github.com/Katsevich-Lab/sceptre-manuscript}. (permanent version deposited at Zenodo; DOI 10.5281/zenodo.5643541 \cite{barrycode}). All software is released under an MIT license.
\section*{Ethical approval}
Ethical approval not applicable.

\clearpage
\appendix
\clearpage
\section*{Main text figures}
% 1
\begin{figure}[h]
\includegraphics[width = \textwidth]{Figure1.jpg}
\caption{}
\label{fig:analysis-challenges}
\end{figure}
\clearpage
% 2
\begin{figure}[h]
	\includegraphics[width = \textwidth]{Figure2.png}
	\caption{}
	\label{fig:method}
\end{figure}
\clearpage
% 3
\begin{figure}[h]
	\includegraphics[width = \textwidth]{Figure3.png}
	\label{fig:ground_truth}
	\caption{}
\end{figure}
\clearpage
% 4
\begin{figure}[h]
	\includegraphics[width = \textwidth]{Figure4.png}
	\caption{}
	\label{fig:application}
\end{figure}
\clearpage
% 5
\clearpage
\begin{figure}[h]
	\includegraphics[width = \textwidth]{Figure5.png}
	\caption{}
	\label{fig:application-xie}
\end{figure}

\clearpage
\section*{Supplementary figures and tables}
\setcounter{figure}{0}    
\renewcommand{\thefigure}{S\arabic{figure}}

\begin{figure}[h]
	\centering
	\includegraphics[width = \textwidth]{FigureS1.png}
	\caption{\textbf{Gasperini et al.'s empirical correction is insufficient to correct for miscalibration}. \textbf{a,} Dispersion estimation procedure employed leads to miscalibration for high-dispersion genes, which the empirical correction does not adequately correct for, as measured by KS test applied to empirical $p$-values per gene (point colors). \textbf{b,} Raw $p$-values already well-calibrated for \textit{LUC7L2} gene, so empirical correction unnecessarily shrinks the significance of the association with TSS-targeting gRNA, depicted by horizontal lines, by three orders of magnitude. \textbf{c,} Empirical correction not strong enough for \textit{HIST1H1D}, which is among circled genes in panel a, which have an NTC-based miscalibration $p$-value smaller than the Bonferroni threshold. \textbf{d}, Under-correction leads to two potential false discoveries for \textit{HIST1H1D}. Dashed horizontal line represents the multiple testing threshold.}
	\label{fig:gasp_correction}
\end{figure}

\clearpage

\thispagestyle{empty} 
\begin{figure}
	\includegraphics[width = \textwidth]{FigureS2.png}
	\caption{\textbf{Comparison of negative binomial and conditional resampling $p$-values based on the same test statistic.} \textbf{a,} The standard parametric negative binomial $p$-value versus that obtained from the same test statistic by conditional resampling, for each gene / candidate enhancer pair (both truncated at $10^{-10}$ for visualization). The two can diverge fairly substantially. \textbf{b-d,} Parametric and conditional resampling null distributions for the negative binomial $z$-value in three cases: the conditional resampling $p$-value is more significant (b), the parametric $p$-value is more significant (c), the two $p$-values are about the same (d).}
	\label{fig:NB-vs-CP}
 \end{figure}

 \clearpage
 \thispagestyle{empty}
\begin{figure}
    \centering
	\includegraphics[width = 0.5\textwidth]{FigureS3.pdf}
	\caption{\textbf{Discoveries unique to SCEPTRE on the Gasperini data exhibited greater enrichment for biological signal than those unique to the original method.} This figure shows the 200 discoveries unique to SCEPTRE and the 107 discoveries unique to the original method (in contrast to Figure \ref{fig:application}, which shows the entire discovery set of both methods). \textbf{a,} On average, enhancers discovered by SCEPTRE were less than half the distance to their target genes than those discovered by Gasperini et al. \textbf{b,} $73\%$ of the gene-enhancer pairs discovered by SCEPTRE fell within the same TAD, in contrast to $57\%$ of those discovered by the original method. HI-C interaction frequency was similar across methods (though slightly higher for the original), despite the fact that SCEPTRE found 85 more same-TAD pairs. \textbf{c,} Enhancers returned by SCEPTRE showed significantly greater enrichment across all ChIP-seq targets, especially H3K27ac.}
	\label{fig:symmetric-difference-gasp}
 \end{figure}

  \clearpage
 \thispagestyle{empty} 
\begin{figure}
    \centering
	\includegraphics[width = 0.5\textwidth]{FigureS4.png}
	\caption{\textbf{SCEPTRE-specific discoveries on the Xie et al.\ data were more enriched for biological signals of regulation than those specific to the original method.} This Figure is analogous to Figure \ref{fig:symmetric-difference-gasp} but corresponds to the Xie et al.\ data; only the discoveries unique to SCEPTRE ($n = 56$) and Virtual FACS ($n = 56$) are depicted. \textbf{a,} Gene-enhancer links discovered by SCEPTRE were physically closer (median = $116$ kb) to one another than those discovered by Virtual FACS (median = $366$ kb). \textbf{b,} SCEPTRE pairs exhibited higher HI-C interaction frequency and were more likely to fall within the same TAD (52\%) than Virtual FACS pairs (29\%). \textbf{c,} SCEPTRE pairs showed greater enrichment across all eight cell type-relevant ChIP-seq targets. Virtual FACS odds ratios were exactly equal to zero for six of eight targets.}
	\label{fig:symmetric-difference-xie}
 \end{figure}

 \clearpage
 \thispagestyle{empty}
\begin{figure}
	\includegraphics[width =\textwidth]{FigureS5.png}
	\caption{\textbf{Details on ChIP-seq enrichment analysis.} Fraction of candidate enhancers linked to gene, broken down by quantile of ChIP-seq signal for (a) Gasperini et al.\ data (five quantiles used) and (b) Xie et al.\ data (two quantiles used). ``0'' indicates that the candidate enhancer did not overlap a ChIP-seq peak. \textbf{a}, On the Gasperini et al.\ data, methods generally paired candidate enhancers in higher ChIP-seq quantiles more frequently. This enrichment was most pronounced for SCEPTRE across all eight ChIP-seq targets. \textbf{b}, Trends were less monotonic on the Xie et al.\ data, possibly due to the fact that Xie et al. used a different strategy for selecting candidate enhancers.\cite{Xie2019}
 	}
	\label{fig:chipseq-enrichment-details}
 \end{figure}

  \clearpage
 \thispagestyle{empty} 
\begin{figure}
	\includegraphics[width = \textwidth]{FigureS6.png}
	\caption{\textbf{scMAGeCK simulation study.} Application of an in-house, specialized version of scMAGeCK to the simulated data. Compared to SCEPTRE (Figure \ref{fig:ground_truth}), scMAGeCK produced inflated $p$-values.}
	\label{fig:scmageck}
 \end{figure}

\clearpage
\thispagestyle{empty}
\setcounter{table}{0}    
\renewcommand{\thetable}{S\arabic{table}}
\begin{table}
\begin{center}
	\begin{tabular}{|lllll|} 
		\hline
		\textbf{Quintile} & \textbf{\textit{N} genes}  & \textbf{\textit{N} pairs} & \textbf{Mean expression} & \textbf{Percent rejected} \\ 
		\hline
		1 & 2108 & 16253 & 0.33 & 0.13\% \\ 
		\hline
			2 & 2108 & 16702 & 1.5 & 0.33\% \\ 
		\hline
			3 & 2108 & 16452 & 3.8 & 0.62\% \\
		\hline
	 	4 & 2108 & 16692 & 8.9 & 0.88\% \\
		\hline
		5 & 2108 & 18496 & 105.0 & 1.30\% \\
		\hline
	\end{tabular}
	\caption{\label{binning_tab_gasp} To investigate the impact of gene expression level on the sensitivity of SCEPTRE, we binned candidate \textit{cis} genes into five groups based on their mean expression (i.e., mean number of UMIs per cell). Gene-enhancer pairs consisting of genes in higher expression quintiles were more likely to be rejected than pairs consisting of genes in lower expression quintiles, suggesting that SCEPTRE was better able to detect gene-enhancer links for highly-expressed genes. Results reported for Gasperini et al.\ data.}
\end{center}
\end{table}

\begin{table}
\begin{center}
	\begin{tabular}{|lllll|} 
		\hline
		\textbf{Tertile} & \textbf{\textit{N} genes}  & \textbf{\textit{N} pairs} & \textbf{Mean expression} & \textbf{Percent rejected} \\ 
		\hline
		1 & 249 & 1867 & 0.001 & 1.66\% \\ 
		\hline
			2 & 249 & 1868 & 0.16 & 3.32\% \\ 
		\hline
			3 & 248 & 1474 & 68.2 & 3.12\% \\
		\hline
	\end{tabular}
	\caption{\label{binning_tab_xie} We replicated the analysis reported in table \ref{binning_tab_gasp} on the Xie et al.\ data, binning genes into tertiles rather than quintiles. We observed a similar pattern: genes in the second and third tertile were more likely to be rejected than genes in the first tertile.}
\end{center}
\end{table}

\end{document}

### 2021-11-04 sceptre_paper/plotting/FigureS4.R

@@ -14,8 +14,6 @@ original_results <- left_join(original_results, resampling_results[, c('gene_id'
                                                    'target_site.mid')], 
           by = c("gene_id", "gRNA_id"))
 
-
-
 # subfigure a (distances)
 df = original_results %>% 
   dplyr::rename(rejected_old = rejected, old_pvalue = ss.down) %>%

### 2021-11-04 sceptre_paper/plotting/load_data_for_plotting.R

@@ -14,8 +14,8 @@ original_results_xie = readRDS(sprintf("%s/data/xie/processed/raw_pval_xie.rds",
 resampling_results_xie_with_names = read.fst(sprintf("%s/results/xie/sceptre/all_results_annotated.fst", offsite_dir)) %>% as_tibble()
 likelihood_results_xie = read.fst(sprintf("%s/results/xie/negative_binomial/all_results.fst", offsite_dir)) %>% as_tibble()
 
-p_vals_bulk <- paste0(offsite_dir, "/results/xie/bulk_rna_seq/pvals_arl15_enh.rds") %>% readRDS() %>% as_tibble() %>% rename(gene_names = gene_id)
-p_vals_bulk_myb3_enh3 <- paste0(offsite_dir, "/results/xie/bulk_rna_seq/pvals_myb_enh3.rds") %>% readRDS() %>% as_tibble() %>% rename(gene_names = gene_id)
+# p_vals_bulk <- paste0(offsite_dir, "/results/xie/bulk_rna_seq/pvals_arl15_enh.rds") %>% readRDS() %>% as_tibble() %>% rename(gene_names = gene_id)
+# p_vals_bulk_myb3_enh3 <- paste0(offsite_dir, "/results/xie/bulk_rna_seq/pvals_myb_enh3.rds") %>% readRDS() %>% as_tibble() %>% rename(gene_names = gene_id)
 covariates_xie = read.fst(sprintf("%s/data/xie/analysis_ready/covariate_model_matrix.fst", offsite_dir)) %>% as_tibble()
 grna_indicator_matrix_xie = read.fst(sprintf("%s/data/xie/analysis_ready/gRNA_indicator_matrix.fst", offsite_dir)) %>% as_tibble()
 ss_xie_cis = readRDS(sprintf("%s/data/xie/processed/ss_xie_cis.rds", offsite_dir)) %>% as_tibble()

### 2022-07-15 sceptre_paper/katsevich2020/R/plotting_functs.R
finalize proofs

@@ -3,20 +3,17 @@
 #' @param base base of the log (default e)
 #' @export
 revlog_trans <- function(base = exp(1)) {
-  ## Define the desired transformation.
   trans <- function(x) {
     -log(x, base)
   }
-  ## Define the reverse of the desired transformation
   inv <- function(x){
     base^(-x)
   }
-  ## Creates the transformation
-  trans_new(paste("revlog-", base, sep = ""),
-            trans, ## The transformation function (can be defined using anonymous functions)
-            inv,  ## The reverse of the transformation
-            log_breaks(base = base), ## default way to define the scale breaks
-            domain = c(1e-100, Inf) ## The domain over which the transformation is valued
+  scales::trans_new(paste("revlog-", base, sep = ""),
+            trans,
+            inv,
+            log_breaks(base = base),
+            domain = c(1e-100, Inf)
   )
 }
 
